###
### name: דוד בורג
### ID  : [מספר תעודת זהות]
### course: אפידמיולוגיה (HIT.ac.il)
### lesson: סטטיסטיקה תאורית וניתות בסיסי 02-04   
### date  : 15/06/2025
###  
###  נלמד לבצע ניתוח סטטיסטי פשוט
###


## Reset memory
```{r setup-packages, include false}
rm(list = ls(all.names = TRUE))
lapply(paste('package:',names(sessionInfo()$otherPkgs),sep=""),detach,character.only=TRUE,unload=TRUE,force=TRUE)
gc()
```

## Load required packages
```{r load-packages, include false}  
options("install.lock"=FALSE)
install.packages("pacman")
library(pacman)
pacman::p_load(
  roi,          # File import
  here,         # File locator
  skimr,        # get overview of data
  tidyverse,    # data management + ggplot2 graphics, 
  gtsummary,    # summary statistics and tests
  rstatix,      # statistics
  nortest,      # Anderson-Darling test for normality
  moments,      # skewness and kurtosis
  corrr,        # correlation analayis for numeric variables
  janitor,      # adding totals and percents to tables
  flextable,    # converting tables to HTML
  dplyr,        # For data manipulation
  ggplot2,      # For creating plots
  DescTools,    # For statistical functions like skewness and kurtosis
  car,          # For Levene's test and ANOVA
  vctrs         # For data type handling
            )
```

## Access the data
```{r read-data, echo=FALSE}
## read the data in a file online
# file <- "linelist_cleaned.rds"     # use this if the file is on your computer
file <- "https://github.com/appliedepi/epirhandbook_eng/raw/master/data/case_linelists/linelist_cleaned.rds"
df1 <- import(file, trust  = TRUE)
summary(df1)
glimpse(df1)
```

## Replace bmi with calculation
```{r read-data, echo=FALSE}
## read the data in a file online
df1 <- df1 %<%
  mutate(bmi = wt_kg / (ht_cm / 100)^2)
summary(df1)
glimpse(df1)
```

## Select numerical variable for analysis
```{r descriptives, echo=TRUE}
## ---= Central tendency and dispersion =---
variable <- "age"
```

## Desctiptive statistics for a numeric variable
```{r descriptives, echo=TRUE}
## ---= Central tendency and dispersion =---
mean1 <- mean(df1[[variable]], na.rm = TRUE)        # Mean
mode1 <- as.vector(Mode(df1[[variable]]))           # Mode
median1 <- median(df1[[variable]], na.rm = TRUE)    # Median
sd1 <- sd(df1[[variable]], na.rm = TRUE)            # Standard Deviation
iqr1 <- IQR(df1[[variable]], na.rm = TRUE)          # Interquartile Range
min1 <- min(df1[[variable]], na.rm = TRUE)          # Minimum
max1 <- max(df1[[variable]], na.rm = TRUE)          # Maximum

# Create a data frame to neatly display the statistics
stats_df <- data.frame(
  Statistic = c("Mean", "Mode", "Median", "Standard Deviation", "Interquartile Range", "Minimum", "Maximum"),
  Value = c(mean1, mode1, median1, sd1, iqr1, min1, max1)
)
# print a table with knitr
knitr::kable(stats_df, col.names = c("Statistic", "Value"), caption = "Summary Statistics")
```

## Normality, skewness and kurtosis
```{r read-data, echo=FALSE}
## Perform a Anderson-Darling test for normality
norm1 <- ad.test(df1[[variable]][!is.na(df1[[variable]])])$p.value
## Calculate skewness
skew1 <- skewness(df1[[variable]], na.rm = TRUE)
## Calculate kurtosis
kurt1 <- kurtosis(df1[[variable]], na.rm = TRUE)

# Create a data frame to neatly display the statistics
stats_df <- data.frame(
  Statistic = c("Mean", "Mode", "Median", "Standard Deviation", "Interquartile Range", "Minimum", "Maximum", "Skewnees", "Kurtosis", "Normality (p<)"),
  Value = c(mean1, mode1, median1, sd1, iqr1, min1, max1, skew1, kurt1, norm1)
)
# print a table with knitr
knitr::kable(stats_df, col.names = c("Statistic", "Value"), caption = "Summary Statistics")
```

## Nice summanry statistis table (with confidence intervals)
```{r nice-table, echo=TRUE}
## Use gtsummary
summary_stats <- tbl_summary(
  data = df1,
  include = bmi,
  statistic = list(all_continuous() ~
             c("{min}",
               "{max}",
               "{mean}",
               "{sd}",
               "{median}",
               "[{p25}, {p75}]", 
               #"Skewness: {skew1}",
               #"Kurtosis: {kurt1}"
               "{n_missing}"
              )),
  digits = all_continuous() ~ 1,
    type = list(bmi = "continuous2"),
  missing = "no"
) %>%
  modify_header(label = "**Variable**") %>%
  modify_caption("**Summary of BMI**")

# Add skewness and kurtosis manually to the table body
summary_stats$table_body <- summary_stats$table_body %>%
  bind_rows(
    tibble(
      variable = "bmi",
      label = "Skewness",
      stat_0 = as.character(round(skew1, 1))
    ),
    tibble(
      variable = "bmi",
      label = "Kurtosis",
      stat_0 = as.character(round(kurt1, 1))
    ),
    tibble(
      variable = "bmi",
      label = "Normality (p<)",
      stat_0 = as.character(round(norm1, 3))
    )
  )  

# Print the table
summary_stats
```


## Pie Chart 
```{r pie, echo=TRUE}
pie_var <- "age_cat"  # Change this to the categorial variable
ggplot(df1, aes(x = "", fill = .data[[pie_var]])) +
  geom_bar(width = 1, stat = "count") +
  coord_polar(theta = "y") +
  theme_void() +
  #scale_fill_manual(values = c("lightblue", "lightgreen", "lightcoral", "lightgoldenrodyellow", "lightpink")) +
  labs(
    title = "Chart of age_cat",
  )
```

## Histogram
```{r histogram, echo=TRUE}
hist_var <- "age"   # Change this to the numerical variable
ggplot(df1, aes(x = .data[[hist_var]])) +
  geom_histogram(binwidth = 5, fill = "grey", color = "black") +
  theme_minimal() +
  labs(
    title = "Histogram of Age",
    x = "Age (years)",
    y = "Count"
  )
```


## Boxplot
```{r boxplot, echo=TRUE}
x_var <- "cough"  # Change this to the categorial variable
y_var <- "temp"   # Change this to the numerical variable
ggplot(df1, aes(x = .data[[x_var]], y = .data[[y_var]], fill = .data[[x_var]])) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.2, alpha = 0.25, size = 0.25, color = "grey") +
  theme_minimal() +
  scale_fill_manual(values = c("yes" = "lightblue", "no" = "lightgreen", "unknown" = "lightgray")) +
  labs(
    title = "Boxplot of Temperature by Cough Status",
    x = "Cough",
    y = "°C",
    fill = "Cough"
  )
```





## T test for independent samples
```{r t-test, echo=FALSE}
# Test for variance homogeneity
levene_result <- levene(ht_cm ~ gender, data = df1, var.equal = TRUE)
(levene_result)

t_test_result <- t.test(ht_cm ~ gender, data = df1, var.equal = FALSE)
t_test_result
```

# Paired T test
```{r paired-t-test, echo=FALSE}
paired_t_test_result <- t.test(data$weight0, data$weight1, paired = TRUE)
print(paired_t_test_result)
```

## Two-way ANOVA
```{r ANOVA, echo=FALSE}
# Perform Levene's Test for homogeneity of variances

# Do the ANOVA
anova_result <- aov(temp ~ cough * fever, data = df1)
summary(anova_result)

# Tukey HSD test
tukey_result <- TukeyHSD(anova_result)
tukey_result
```

## Chi-Square Test 
```{r chi-square, echo=FALSE}


```









```{r question1}

library(ggplot2)


ggplot(data, aes(x = height0, y = weight0)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Adding the trend line with linear model
  labs(title = "Scatter Plot of Weight vs Height",
       x = " Height (cm)",
       y = "Weight (kg)")



##pearson
test_result <- cor.test(data$weight0, data$height0, method = "pearson", use = "complete.obs")
print(test_result)



#spearman
spearman_test_result <- cor.test(data$weight0, data$height0, method = "spearman", use = "complete.obs")

# Print the test result
print(spearman_test_result)
```
##chi square
```{r pressure, echo=FALSE}

 
data$pacifier_dich <- factor(data$pacifier_dich, levels = c(0, 1), labels = c("no", "yes"))
data$newborn_gender <- factor(data$newborn_gender, levels = c(0, 1), labels = c("male", "female"))
# Create a contingency table
table_data <- table(data$pacifier_dich, data$newborn_gender)
table_data

# Calculate and print the percentage table
percent_table <- prop.table(table_data, margin = 1) * 100  # Row percentages
print(percent_table)
print(column_percentages)
column_percentages <- prop.table(table_data, margin = 2) * 100 #column percentages
print(column_percentages)

# Perform the chi-square test
chi_test <- chisq.test(table_data)

# Print the results of the chi-square test
print(chi_test)


expected_counts <- chi_test$expected
expected_counts
count_violations <- sum(expected_counts < 5)
total_counts <- length(expected_counts)
percentage_violations <- (count_violations / total_counts) * 100
percentage_violations


cat("Percentage of expected counts less than 5:", percentage_violations, "%\n")
if (percentage_violations > 20) {
  cat("Warning: More than 20% of expected counts are less than 5, which violates chi-square test assumptions.\n")
}

```


## OR and RR for 2*2
```{r question1}
install.packages("epitools")
library(epitools)

# Create a contingency table
contingency_table <- table(data$newborn_gender ,data$pacifier_dich)

# Compute Odds Ratio (OR) and Relative Risk (RR)
OR_result <- epitab(contingency_table, method = "oddsratio")
print(OR_result)

RR_result <- epitab(contingency_table, method = "riskratio")
print(RR_result)
```


##fisher
```{r pressure, echo=FALSE}
contingency_table <- table(data$newborn_gender, data$pacifier_dich)

# Perform Fisher's exact test
fisher_result <- fisher.test(contingency_table)

# Print the result
print(fisher_result)
```
