###
### name: דוד בורג
### ID  : [מספר תעודת זהות]
### course: אפידמיולוגיה (HIT.ac.il)
### lesson: סטטיסטיקה תאורית וניתות בסיסי 02-04   
### date  : 15/06/2025
###  
###  נלמד לבצע ניתוח סטטיסטי פשוט
###
### source: https://epirhandbook.com/en/


## Reset memory
```{r setup-packages, include false}
rm(list = ls(all.names = TRUE))
lapply(paste('package:',names(sessionInfo()$otherPkgs),sep=""),detach,character.only=TRUE,unload=TRUE,force=TRUE)
gc()
dev.off()
```

## Load required packages
```{r load-packages, include false}  
options("install.lock"=FALSE)
if (!require("pacman")) install.packages("pacman")
library(pacman)
pacman::p_load(
  rio,          # File import
  tidyverse,    # data management + ggplot2 graphics, 
  gt,           # Nice beuatiful tables
  nortest,      # Anderson-Darling test for normality
  ggplot2,      # For creating plots
  DescTools    # For statistical functions like skewness and kurtosis
            )
```

## Access the data
```{r read-data, echo=FALSE}
## read the data in a file online
# file <- "./EpiData/linelist_cleaned.rds"     # use this if the file is on your computer
file <- "https://github.com/appliedepi/epirhandbook_eng/raw/master/data/case_linelists/linelist_cleaned.rds"
df1 <- import(file, trust  = TRUE)
summary(df1)
glimpse(df1)
```

## Replace bmi with calculation
```{r read-data, echo=FALSE}
## read the data in a file online
df1 <- df1 %<%
  mutate(bmi = wt_kg / (ht_cm / 100)^2)
summary(df1)
glimpse(df1)
```

## Select numerical variable for analysis
```{r descriptives, echo=TRUE}
variable <- "age"
```

## Desctiptive statistics for a numeric variable
```{r descriptives, echo=TRUE}

# Summary statistics with Desctools
desc_result <- Desc(df1[variable], plotit = TRUE)
desc_result

mean1 <- Mean(df1[[variable]], na.rm = TRUE)
sd1 <- sd(df1[[variable]], na.rm = TRUE)
CI1 <- MeanCI(df1[[variable]], na.rm = TRUE)
  lower_meanci_1 <- as.numeric((CI1["lwr.ci"]))
  upper_meanci_1 <- as.numeric((CI1["upr.ci"]))
mode1 <- as.numeric(Mode(df1[[variable]], na.rm = TRUE))
median1 <- Median(df1[[variable]], na.rm = TRUE)
q25_1 <- as.numeric(Quantile(df1[[variable]], probs = 0.25, na.rm = TRUE))
q75_1 <- as.numeric(Quantile(df1[[variable]], probs = 0.75, na.rm = TRUE))
IQR1 <- IQR(df1[[variable]], na.rm = TRUE)
min1 <- min(df1[[variable]], na.rm = TRUE)
max1 <- max(df1[[variable]], na.rm = TRUE)
skew1 <- Skew(df1[[variable]], na.rm = TRUE)                            # Skewness
kurt1 <- Kurt(df1[[variable]], na.rm = TRUE)                            # Kurtosis
norm1 <- ad.test(df1[[variable]][!is.na(df1[[variable]])])$p.value      # Anderson-Darling test for normality
```

## Nice summanry statistis table
```{r nice-table, echo=TRUE}
# Create a data frame to neatly display the statistics
stats_df <- data.frame(
  Statistic = c("Mean", "SD", "⠀⠀CI₉₅ low", "⠀⠀CI₉₅ high", "Mode", "Median", "⠀⠀Q1", "⠀⠀Q3", "⠀⠀IQR", "Minimum", "Maximum", "Skewness", "Kurtosis", "Normality test")
  , Value = c(mean1, sd1, lower_meanci_1, upper_meanci_1, mode1, median1, q25_1, q75_1, IQR1, min1, max1, skew1, kurt1, norm1)
  , סטטיסטי = c("ממוצע", "ס'ת", "⠀⠀רווח סמך", "⠀⠀רווח סמך", "שכיח", "חציון", "⠀⠀רבעון 1", "⠀⠀רבעון 3", "⠀⠀מרווח בין רבועני", "מינימום", "מקסימום", "צידוד", "גבנוניות", "נורמליות")
)

# Clean, simple table
summary_stats <- stats_df %>% 
  gt() %>%
  tab_header(title = "Summary Statistics⠀⠀⠀⠀סטטיסטקה תאורית") %>%
  fmt_number(columns = Value, decimals = 1)

summary_stats              # Show the table
```


## Pie Chart 
```{r pie, echo=TRUE}
pie_var <- "age_cat"  # Change this to the categorial variable
ggplot(df1, aes(x = "", fill = .data[[pie_var]])) +
  geom_bar(width = 1, stat = "count") +
  coord_polar(theta = "y") +
  theme_void() +
  #scale_fill_manual(values = c("lightblue", "lightgreen", "lightcoral", "lightgoldenrodyellow", "lightpink")) +
  labs(
    title = "Chart of age_cat",
  )
```

## Histogram
```{r histogram, echo=TRUE}
hist_var <- variable 
ggplot(df1, aes(x = get(hist_var))) +
  geom_histogram(binwidth = 5, fill = "grey", color = "black") +
  theme_minimal() +
  labs(
    title = "Histogram of Age",
    x = "Age (years)",
    y = "Count"
  )
```


## Boxplot
```{r boxplot, echo=TRUE}
x_var <- "cough"  # Change this to a categorial variable
y_var <- "temp"   # Change this to a numerical variable
ggplot(df1, aes(x = get(x_var), y = get(y_var), fill = get(x_var))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.2, alpha = 0.25, size = 0.25, color = "grey") +
  theme_minimal() +
  scale_fill_manual(values = c("yes" = "lightblue", "no" = "lightgreen", "unknown" = "lightgray")) +
  labs(
    title = "Boxplot of Temperature by Cough Status",
    x = "Cough",
    y = "°C",
    fill = "Cough"
  )
```
