###
### name: דוד בורג
### ID  : [מספר תעודת זהות]
### course: אפידמיולוגיה (HIT.ac.il)
### lesson: סטטיסטיקה תאורית וניתות בסיסי 02-04   
### date  : 15/06/2025
###  
###  נלמד לבצע ניתוח סטטיסטי פשוט
###


## Reset memory
```{r setup-packages, include false}
rm(list = ls(all.names = TRUE))
lapply(paste('package:',names(sessionInfo()$otherPkgs),sep=""),detach,character.only=TRUE,unload=TRUE,force=TRUE)
gc()
```

## Load required packages
```{r load-packages, include false}  
options("install.lock"=FALSE)
if (!require("pacman")) install.packages("pacman")
library(pacman)
pacman::p_load(
  rio,          # File import
  here,         # File locator
  skimr,        # get overview of data
  tidyverse,    # data management + ggplot2 graphics, 
  gtsummary,    # summary statistics and tests
  rstatix,      # statistics
  nortest,      # Anderson-Darling test for normality
  moments,      # skewness and kurtosis
  corrr,        # correlation analayis for numeric variables
  janitor,      # adding totals and percents to tables
  flextable,    # converting tables to HTML
  dplyr,        # For data manipulation
  ggplot2,      # For creating plots
  DescTools,    # For statistical functions like skewness and kurtosis
  car,          # For Levene's test and ANOVA
  vctrs         # For data type handling
            )
```

## Access the data
```{r read-data, echo=FALSE}
## read the data in a file online
# file <- "./EpiData/linelist_cleaned.rds"     # use this if the file is on your computer
file <- "https://github.com/appliedepi/epirhandbook_eng/raw/master/data/case_linelists/linelist_cleaned.rds"
df1 <- import(file, trust  = TRUE)
summary(df1)
glimpse(df1)
```

## Replace bmi with calculation
```{r read-data, echo=FALSE}
## read the data in a file online
df1 <- df1 %<%
  mutate(bmi = wt_kg / (ht_cm / 100)^2)
summary(df1)
glimpse(df1)
```

## Select numerical variable for analysis
```{r descriptives, echo=TRUE}
## ---= Central tendency and dispersion =---
variable <- "age"
```

## Desctiptive statistics for a numeric variable
```{r descriptives, echo=TRUE}
## ---= Central tendency and dispersion =---
mean1 <- mean(df1[[variable]], na.rm = TRUE)        # Mean
mode1 <- as.vector(Mode(df1[[variable]]))           # Mode
median1 <- median(df1[[variable]], na.rm = TRUE)    # Median
sd1 <- sd(df1[[variable]], na.rm = TRUE)            # Standard Deviation
iqr1 <- IQR(df1[[variable]], na.rm = TRUE)          # Interquartile Range
min1 <- min(df1[[variable]], na.rm = TRUE)          # Minimum
max1 <- max(df1[[variable]], na.rm = TRUE)          # Maximum

# Create a data frame to neatly display the statistics
stats_df <- data.frame(
  Statistic = c("Mean", "Mode", "Median", "Standard Deviation", "Interquartile Range", "Minimum", "Maximum"),
  Value = c(mean1, mode1, median1, sd1, iqr1, min1, max1)
)
# print a table with knitr
knitr::kable(stats_df, col.names = c("Statistic", "Value"), caption = "Summary Statistics")
```

## Normality, skewness and kurtosis
```{r read-data, echo=FALSE}
## Perform a Anderson-Darling test for normality
norm1 <- ad.test(df1[[variable]][!is.na(df1[[variable]])])$p.value
## Calculate skewness
skew1 <- skewness(df1[[variable]], na.rm = TRUE)
## Calculate kurtosis
kurt1 <- kurtosis(df1[[variable]], na.rm = TRUE)

# Create a data frame to neatly display the statistics
stats_df <- data.frame(
  Statistic = c("Mean", "Mode", "Median", "Standard Deviation", "Interquartile Range", "Minimum", "Maximum", "Skewnees", "Kurtosis", "Normality (p<)"),
  Value = c(mean1, mode1, median1, sd1, iqr1, min1, max1, skew1, kurt1, norm1)
)
# print a table with knitr
knitr::kable(stats_df, col.names = c("Statistic", "Value"), caption = "Summary Statistics")
```

## Nice summanry statistis table (with confidence intervals)
```{r nice-table, echo=TRUE}
## Use gtsummary
summary_stats <- tbl_summary(
  data = df1,
  include = bmi,
  statistic = list(all_continuous() ~
             c("{min}",
               "{max}",
               "{mean}",
               "{sd}",
               "{median}",
               "[{p25}, {p75}]", 
               #"Skewness: {skew1}",
               #"Kurtosis: {kurt1}"
               "{n_missing}"
              )),
  digits = all_continuous() ~ 1,
    type = list(bmi = "continuous2"),
  missing = "no"
) %>%
  modify_header(label = "**Variable**") %>%
  modify_caption("**Summary of BMI**")

# Add skewness and kurtosis manually to the table body
summary_stats$table_body <- summary_stats$table_body %>%
  bind_rows(
    tibble(
      variable = "bmi",
      label = "Skewness",
      stat_0 = as.character(round(skew1, 1))
    ),
    tibble(
      variable = "bmi",
      label = "Kurtosis",
      stat_0 = as.character(round(kurt1, 1))
    ),
    tibble(
      variable = "bmi",
      label = "Normality (p<)",
      stat_0 = as.character(round(norm1, 3))
    )
  )  

# Print the table
summary_stats
```


## Pie Chart 
```{r pie, echo=TRUE}
pie_var <- "age_cat"  # Change this to the categorial variable
ggplot(df1, aes(x = "", fill = .data[[pie_var]])) +
  geom_bar(width = 1, stat = "count") +
  coord_polar(theta = "y") +
  theme_void() +
  #scale_fill_manual(values = c("lightblue", "lightgreen", "lightcoral", "lightgoldenrodyellow", "lightpink")) +
  labs(
    title = "Chart of age_cat",
  )
```

## Histogram
```{r histogram, echo=TRUE}
hist_var <- "age"   # Change this to the numerical variable
ggplot(df1, aes(x = .data[[hist_var]])) +
  geom_histogram(binwidth = 5, fill = "grey", color = "black") +
  theme_minimal() +
  labs(
    title = "Histogram of Age",
    x = "Age (years)",
    y = "Count"
  )
```


## Boxplot
```{r boxplot, echo=TRUE}
x_var <- "cough"  # Change this to the categorial variable
y_var <- "temp"   # Change this to the numerical variable
ggplot(df1, aes(x = .data[[x_var]], y = .data[[y_var]], fill = .data[[x_var]])) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.2, alpha = 0.25, size = 0.25, color = "grey") +
  theme_minimal() +
  scale_fill_manual(values = c("yes" = "lightblue", "no" = "lightgreen", "unknown" = "lightgray")) +
  labs(
    title = "Boxplot of Temperature by Cough Status",
    x = "Cough",
    y = "°C",
    fill = "Cough"
  )
```
