###
### name  :  专
### ID    : [住驻专 转注转 转]
### course: 驻 (HIT.ac.il)
### lesson: 住住拽 转专转 转 住住 02-03   
### date  : 13/07/2025
###  
###   爪注 转 住住 驻砖


## Reset memory
```{r setup-packages, include = FALSE}
rm(list = ls(all.names = TRUE))       #  Clear all data objects         Clear packages
lapply(paste("package:", names(sessionInfo()$otherPkgs), sep = ""), detach, character.only = TRUE, unload = TRUE, force = TRUE)           # Remove packages from memory
gc()                                  # Clear unused data from memory ("garbage cllection")
dev.off()                             # Reset the display output of R
q("no")                               # This will completely shutdown R terminal
```

## Load required packages
```{r load-packages, include = FALSE}  
options("install.lock" = FALSE)
if (!require("pacman")) install.packages("pacman") ; library(pacman)
pacman::p_load(
  Rcpp, devtools, gh, remotes,   # For installing packages/files online (like from GitHub)
  rio, here,              # File import and path support
  labelled,               # deal with labels
  tidyverse,              # Data management
  broom,                  # For tidying model outputs
  dplyr,                  # For data manipulation
  gt,                     # Nice beuatiful tables
  gtsummary,              # summary statistics and tests
  nortest,                # Anderson-Darling test for normality
  ggplot2,                # For creating plots
  DescTools,              # For statistical functions like skewness and kurtosis
  summarytools            # to get a table of all summary statistics
              )
```

## Access the data
```{r read-data, echo = FALSE}
# file <- "./EpiData/01 ebola.rds"     #  use this if the file is on your computer
file <- "https://raw.githubusercontent.com/Model-Lab-Net/Courses/refs/heads/main/Epi/Data/01%20ebola.rds"
df1 <- import(file, trust = TRUE)
# show variable translations to Hebrew in lables
labels <- data.frame(Variable = names(df1), Label = unlist(var_label(df1)))   ;   gt(labels)
glimpse(df1)               # get an overview of the data
```

## Select numerical variable for analysis
```{r descriptives, echo = TRUE}
# Select variables from the df to calculate summary statistics
variables <- c("age", "bmi", "temp")

# Calculate summary statistics
summary_stats <- descr(df1[, variables], round.digits = 2, headings = TRUE)
summary_stats
```

## Send to nice table
```{r descriptives, echo = TRUE}
# Switch around the table
summary_stats <- as.data.frame((summary_stats))
summary_stats <- cbind(Statistic = rownames(summary_stats), summary_stats)   # need this to keep

# Convert to nice table
summary_stats_tbl <- summary_stats %>%
  gt() %>%
  tab_header(title = "Descriptive Statistics住住拽 转专转") %>%
  fmt_number(
    columns = where(is.numeric),
    decimals = 2
  )
summary_stats_tbl # Display the table
```

## Test normality (Ho: data are normal dist)
```{r descriptives-old, echo = TRUE}
variable <- c("age")
ad <- ad.test(df1[[variable]])$p.value      # Anderson-Darling test for normality
ad <- paste0("p < ", formatC(ad, format = "f", digits = 4))
ad
```

## The old way to do descriptives
```{r descriptives-old, echo = TRUE}
# Select variable from the df to calculate summary statistics
variable <- c("temp")

# Get summary statistics
desc_result <- Desc(df1[variable], plotit = TRUE)
desc_result

# Collect all the values
mean_var1 <- mean(df$var.1, na.rm = TRUE)
median_var1 <- median(df$var.1, na.rm = TRUE)

sd_var1 <- mean(df$var.1, na.rm = TRUE)
iqr_var1 <- median(df$var.1, na.rm = TRUE)

mean1 <- Mean(df1[[variable]], na.rm = TRUE)
sd1 <- sd(df1[[variable]], na.rm = TRUE)
CI1 <- MeanCI(df1[[variable]], na.rm = TRUE)
   lower_meanci_1 <- as.numeric((CI1["lwr.ci"]))
   upper_meanci_1 <- as.numeric((CI1["upr.ci"]))
mode1 <- as.numeric(Mode(df1[[variable]], na.rm = TRUE))
median1 <- Median(df1[[variable]], na.rm = TRUE)
q25_1 <- as.numeric(Quantile(df1[[variable]], probs = 0.25, na.rm = TRUE))
q75_1 <- as.numeric(Quantile(df1[[variable]], probs = 0.75, na.rm = TRUE))
IQR1 <- IQR(df1[[variable]], na.rm = TRUE)
min1 <- min(df1[[variable]], na.rm = TRUE)
max1 <- max(df1[[variable]], na.rm = TRUE)
skew1 <- Skew(df1[[variable]], na.rm = TRUE) # Skewness
kurt1 <- Kurt(df1[[variable]], na.rm = TRUE) # Kurtosis
norm1 <- ad.test(df1[[variable]][!is.na(df1[[variable]])])$p.value # Anderson-Darling test for normality

# Create a data frame to neatly display the statistics
stats_df <- data.frame(
  Statistic = c("Mean", "SD", "CI low", "CI high", "Mode", "Median", "Q1", "Q3", "IQR", "Minimum", "Maximum", "Skewness", "Kurtosis", "Normality test"),
  Value = c(mean1, sd1, lower_meanci_1, upper_meanci_1, mode1, median1, q25_1, q75_1, IQR1, min1, max1, skew1, kurt1, norm1),
  住住 = c("爪注", "住'转", "专 住", "专 住", "砖", "爪", "专注 1", "专注 3", "专  专注", "", "拽住", "爪", "转", "专转")
)

# Clean, simple table
summary_stats <- stats_df %>%
  gt() %>%
  tab_header(title = "Summary Statistics住住拽 转专转") %>%
  fmt_number(columns = Value, decimals = 1)

summary_stats # Show the table
```

## Pie Chart 
```{r pie, echo = TRUE}
pie_var <- "age.cat"                                   # Change this to the categorial variable
ggplot(df1, aes(x = "", fill = .data[[pie_var]])) +
  geom_bar(width = 1, stat = "count") +
  coord_polar(theta = "y") +
  theme_void() +
  # scale_fill_manual(values = c("lightblue", "lightgreen", "lightcoral", "lightgoldenrodyellow", "lightpink")) +
  labs(
    title = "Chart of age.cat",
  )
```

## Histogram
```{r histogram, echo = TRUE}
hist_var <- variable
ggplot(df1, aes(x = get(hist_var))) +
  geom_histogram(bins = 10, fill = "grey", color = "black") +
  theme_minimal() +
  labs(
    title = "Histogram of Age",
    x = "Age (years)",
    y = "Count"
  )
```

## Boxplot
```{r boxplot, echo = TRUE}
x_var <- "cough" # Change this to a categorial variable
y_var <- "temp" # Change this to a numerical variable
ggplot(df1, aes(x = get(x_var), y = get(y_var), fill = get(x_var))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.2, alpha = 0.25, size = 0.25, color = "#000000") +
  theme_minimal() +
  scale_fill_manual(values = c("yes" = "lightblue",
                   "no" = "lightgreen", "unknown" = "lightgray")
                   ) +
  labs(
    title = "Boxplot of Temperature by Cough Status",
    x = "Cough",
    y = "掳C",
    fill = "Cough"
  )
```

## Plot the histograms and density functions of the data rom above
```{r boxplot, echo = TRUE}
desc_result <- Desc(df1[variables], plotit = TRUE)
desc_result
```

## Crosstables / Contigengcy tables
```{r chi-square, echo = TRUE}
glimpse(df1)
# Create contigency table
contingency_table <- table(df1$fever, df1$outcome)
contingency_table

# Convert to probabilities
percent_table <- prop.table(contingency_table, margin = 1)
percent_table

percent_table <- round(prop.table(contingency_table), 3)      # round to 3 digits
percent_table

# Chi-Square test
chisq.test(x = df1$fever, y = df1$outcome)

# Chi-Fisher Exact test
fisher.test(x = df1$fever, y = df1$outcome)

# Create cross-tabulation with chi-square test
tbl_cross(data = df1_clean, row = fever, col = outcome,
          missing = "no",                                               # Ignore missing values
          percent = "row"                                               # Row percentages (use "column" or "cell" if needed)
         ) %>%
         add_p(test = "chisq.test", test.args = list(correct = TRUE),   # Add chi-square test. Replace with fisher.test if needed.
         pvalue_fun = ~formatC(., digits = 3, format = "f")
         )
```



### -------------------------------------------------
## 转专-注爪

## 砖 1
```{r question-1, echo = TRUE}
# 

```

## 砖 2
```{r question-2, echo = TRUE}
# 

```

## 砖 3
```{r question-3, echo = TRUE}
# 

```










### source: https://epirhandbook.com/en/new_pages/data_used.html
###         https://github.com/appliedepi/epirhandbook_eng

## Clean data
```{r clean-data, echo = FALSE}
# Load the data
file <- "https://github.com/appliedepi/epirhandbook_eng/raw/master/data/case_linelists/linelist_cleaned.rds"
df1 <- import(file, trust = TRUE)
write.csv(df1, here("EpiData", "01 ebola.csv"))

# Clean up variable names
df1 <- clean_names(df1)
names(df1) <- gsub("_", ".", names(df1))

# Convert chr variables to fct
df1 <- df1 %>% 
  mutate_if(is.character, as.factor)

# Convert cough NAs to "no"
df1 <- df1 %>%
  mutate(cough = replace_na(cough, "no"))

# Calculate BMI
df1 <- df1 %>%
  mutate(bmi = wt_kg / (ht_cm / 100)^2)

summary(df1)
glimpse(df1)

# Save to R data file
saveRDS(df1, here("EpiData", "01 ebola.rds"))
```

## Give variable explanations in labels
```{r labels, echo = TRUE}
attr(df1$case.id, "label") <- "拽  砖 驻"
attr(df1$generation, "label") <- "专"
attr(df1$date.infection, "label") <- "转专 拽"
attr(df1$date.onset, "label") <- "转专 转转 转住"
attr(df1$date.hospitalisation, "label") <- "转专 砖驻"
attr(df1$date.outcome, "label") <- "转专 住"
attr(df1$gender, "label") <- "专"
attr(df1$outcome, "label") <- "( / 驻专) 住"
attr(df1$age, "label") <- ""
attr(df1$age.unit, "label") <- " (转)"
attr(df1$age.years , "label") <- " 砖"
attr(df1$age.cat, "label") <- " 拽专"
attr(df1$age.cat5, "label") <- " 拽专 5"
attr(df1$hospital, "label") <- "转 "
attr(df1$lon, "label") <- "拽 拽 专"
attr(df1$lat, "label") <- "拽 拽 专"
attr(df1$infector, "label") <- "拽  砖 拽"
attr(df1$source, "label") <- "拽专"
attr(df1$wt.kg, "label") <- "砖拽 -拽'"
attr(df1$ht.cm, "label") <- " -住'"
attr(df1$ct.blood, "label") <- "转 祝 "
attr(df1$fever, "label") <- ""
attr(df1$chills, "label") <- "专注转"
attr(df1$cough, "label") <- "砖注"
attr(df1$aches, "label") <- ""
attr(df1$vomit, "label") <- "拽转"
attr(df1$temp, "label") <- "驻专专转 祝"
attr(df1$time.admission, "label") <- " 砖驻"
attr(df1$bmi, "label") <- "BMI"
attr(df1$days.onset.hosp, "label") <- "住驻专   驻注转 转住 注 砖驻"

labels <- NULL
labels <- data.frame(Variable = names(df1), Label = unlist(var_label(df1)))
gt(labels)
```
