###
### name: דוד בורג
### course: אפידמיולוגיה (HIT.ac.il)
### lesson: 2
### date: 19/05/2025
###  
###  כאן נלמד לטפל בנתונים
###

# Reset memory
```{r setup-packages, include false}
rm(list = ls(all.names = TRUE))
lapply(paste('package:',names(sessionInfo()$otherPkgs),sep=""),detach,character.only=TRUE,unload=TRUE,force=TRUE)
gc()
```

# Load required packages
```{r setup-packages, include false}
library(tidyverse)
library(kableExtra)
library(tibble)
library(readxl)
library(lubridate)
library(plotly)
```

# Get data
```{r load data, include false}
path <- "./Data/stroke_data.csv"
stroke <- read.csv(path)

path <- "./Data/peptic_ulcer.xlsx"
pep <- read_excel(path)
```

# Check out the variables
```{r load data, include false}
glimpse(stroke)
print(stroke, n=20)


kable_table_stroke <- stroke %>%
  slice_head(n = 20) %>%  # Show first 20 rows
  kable("html", align = 'c', caption = "Tibble for Stroke data (First 20 Rows)") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    font_size = 14
  ) %>%
  scroll_box(height = "400px")

# Display in console/viewer (R Markdown or viewer needed for full effect)
print(kable_table_stroke)


glimpse(pep)
print(pep, n=20)

kable_table_pep <- pep %>%
  slice_head(n = 20) %>%  # Show first 20 rows
  kable("html", align = 'c', caption = "Tibble for Peptic Ulcer data (First 20 Rows)") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    font_size = 14
  ) %>%
  scroll_box(height = "400px")

# Display in console/viewer (R Markdown or viewer needed for full effect)
print(kable_table_pep)
```

# sort a variable
```{r sort_data, include false}
# This take the ten variables we want and put them is a new dataframe
stroke %>% 
  arrange(doa)

kable_table_stroke <- stroke %>%
  slice_head(n = 20) %>%  # Show first 20 rows
  kable("html", align = 'c', caption = "Tibble for Peptic Ulcer smaller data (First 20 Rows)") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    font_size = 14
  ) %>%
  scroll_box(height = "400px")

# Display in console/viewer (R Markdown or viewer needed for full effect)
print(kable_table_stroke)
```


# Select observation using filter()
```{r select_data, include false}
# filter() will select males 7+ years old and send
# the data to a new dataframe
stroke_m_7 <- stroke %>% 
  filter(sex == 'male', gcs >= 7)
stroke_m_7

kable_table_stroke_m_7 <- stroke_m_7 %>%
  slice_head(n = 20) %>%  # Show first 20 rows
  kable("html", align = 'c', caption = "Tibble for Peptic Ulcer smaller data (First 20 Rows)") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    font_size = 14
  ) %>%
  scroll_box(height = "400px")

# Display in console/viewer (R Markdown or viewer needed for full effect)
print(kable_table_stroke_m_7)
```

# Convert numerical variables to binary categorical variables 
```{r um-to-cat, include false}
# This will take the systolic **and** diastolic blood pressure
# and create a new variable called high_bp ("High" or "Not High")
stroke <- stroke %>% 
  mutate(high_bp = if_else(sbp >= 130 | dbp >= 90, 
                           "High", "Not High"))
stroke %>% count(high_bp)

kable_table_stroke <- stroke %>%
  slice_head(n = 20) %>%  # Show first 20 rows
  kable("html", align = 'c', caption = "Tibble for Stroke data (First 20 Rows)") %>%
  kable_styling(
    #bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    #full_width = FALSE,
    font_size = 14
  ) %>%
  scroll_box(height = "400px")

# Display in console/viewer (R Markdown or viewer needed for full effect)
print(kable_table_stroke)
```

# Convert numerical variable to multi- categorical variable
```{r num-to-cat, include false}
# Take the systolic blood pressure and create a new variable
# called cat_sbp with three categories.
# DOESNT WORK YET
stroke <- stroke %>%
  mutate(sbp_cat = case_when(
    sbp < 120 ~ "low",
    sbp >= 120 & sbp <= 130 ~ "normal",
    sbp > 130 ~ "high"
  ))

kable_table_stroke <- stroke %>%
  slice_head(n = 20) %>%  # Show first 20 rows
  kable("html", align = 'c', caption = "Tibble for Peptic Ulcer smaller data (First 20 Rows)") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    font_size = 14
  ) %>%
  scroll_box(height = "400px")

# Display in console/viewer (R Markdown or viewer needed for full effect)
print(kable_table_stroke)
```


# Convert categorical variable to numerical (dummy) variable
```{r cat-to-dummy, include false}
# Replace in-place sex variable (male/female) with 0/1
stroke <- stroke %>%
  mutate(sex = ifelse(sex == "male", 0, 1))

kable_table_stroke <- stroke %>%
  slice_head(n = 20) %>%  # Show first 20 rows
  kable("html", align = 'c', caption = "Tibble for Peptic Ulcer smaller data (First 20 Rows)") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    font_size = 14
  ) %>%
  scroll_box(height = "400px")

# Display in console/viewer (R Markdown or viewer needed for full effect)
print(kable_table_stroke)
```

