###
### name: דוד בורג
### course: אפידמיולוגיה (HIT.ac.il)
### lesson: 3
### date: 19/05/2025
###  
###  סטטיסטיקה תאורית
###

# Reset memory
```{r setup-packages, include false}
rm(list = ls(all.names = TRUE))
lapply(paste('package:',names(sessionInfo()$otherPkgs),sep=""),detach,character.only=TRUE,unload=TRUE,force=TRUE)
gc()
```

# Load required packages
```{r setup-packages, include false}
library(gtsummary)
library(tidyverse)
library(kableExtra)
library(tibble)
library(patchwork)
library(here)
library(readxl)
library(plotly)
library(forcats)  # for fct_reorder
```

# Get data
```{r load data, include false}
library(gapminder)
```

# Check out the variables in Asia
```{r load data, include false}
gap_asia_continent <- gapminder %>% 
  filter(continent == "Asia") %>%
  group_by(country) %>% 
  mutate(mean = mean(lifeExp), sd = sd(lifeExp)) %>%     # add average and standard deviation colums to df
  arrange(desc(mean))                                    # sort from highest to lowest
print(gap_asia_continent)

kable_table_gap_asia_continent <- gap_asia_continent %>%
  ungroup() %>%
  slice_head(n = 200000) %>%  # Show first 20 rows
  kable("html", align = 'c', caption = "Life Expectancy") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    font_size = 14
  ) %>%
  scroll_box(height = "400px")
print(kable_table_gap_asia_continent)
```


# Reanme countries
```{r rename_countries, include false}
# Some countries have difficult or long names. Lets rename a few
gap_asia_continent <- gap_asia_continent %>%
  mutate(country = case_when(
    country == "Korea, Rep." ~ "N.Korea",
    country == "Hong Kong, China" ~ "China (HK)",
    country == "Korea, Dem. Rep." ~ "S.Korea",
    country == "West Bank and Gaza" ~ "P.A.",
    country == "Yemen, Rep." ~ "Yemen",
    TRUE ~ country              # keep other country names unchanged
  ))

kable_table_gap_asia_continent <- gap_asia_continent %>%
  ungroup() %>%
  slice_head(n = 200000) %>%  # Show first 20 rows
  kable("html", align = 'c', caption = "Life Expectancy") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    font_size = 14
  ) %>%
  scroll_box(height = "400px")
print(kable_table_gap_asia_continent)
```

# Show means and standard deviations
```{r mean_sd_graph, include false}
# A graph of means and standard deviations for each country
plotly_mean_sd <- gap_asia_continent %>%   
  ggplot(mapping = aes(x = country, y = mean)) + 
  geom_point(mapping = aes(x = country, y = mean)) +
  geom_errorbar(mapping = aes(ymin = mean - sd, ymax = mean + sd),
                position = position_dodge()) +
  coord_flip() +                      # this flips the x and y axes
  theme_bw()
ggplotly(plotly_mean_sd)
```


# Now sort
```{r select_data, include false}
# the same graph as above - sorted by mean
plotly_sort_mean <- gap_asia_continent %>%   
  ggplot(mapping = aes(x = fct_reorder(country, mean), y = mean)) +  # fct_reorder is a forcats func in tdiverse to reorder - here by mean
  geom_point() +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd),
                position = position_dodge(width = 0.5), width = 0.2) +
  coord_flip() +
  xlab("Country") + ylab("Mean Life Expectancy") +
  ggtitle("Mean Life Expectancy by Country (Asia)") +
  theme_bw()
ggplotly(plotly_sort_mean)
```


# Make Israel first
```{r make_israel_first, include false}
# the same graph as above - Israel is first
plotly_israel_first <- gap_asia_continent %>%
  ungroup() %>%  # important to remove any groupings
  mutate(
    is_israel = ifelse(country == "Israel", 0, 1),
    country = reorder(country, -mean)                 # reorder by descending mean
  ) %>%
  arrange(desc(is_israel), (mean)) %>%
  mutate(country = factor(country, levels = unique(country))) %>%

  ggplot(mapping = aes(x = country, y = mean)) +
  geom_point() +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd),
                position = position_dodge(width = 0.5), width = 0.2) +
  coord_flip() +
  xlab("Country") + ylab("Mean Life Expectancy") +
  ggtitle("Mean Life Expectancy by Country (Asia)") +
  theme(legend.position = "none")
  theme_bw()
ggplotly(plotly_israel_first)
```


# Make Israel first and red
```{r select_data, include false}
# the same graph as above - Israel is red
plotly_Israel_first_red <- gap_asia_continent %>%
  ungroup() %>%
  mutate(
    is_israel = ifelse(country == "Israel", 0, 1),
    country = reorder(country, -mean)
  ) %>%
  arrange(desc(is_israel), (mean)) %>%
  mutate(country = factor(country, levels = unique(country))) %>%

  ggplot(aes(x = country, y = mean)) +
  geom_point(aes(color = country == "Israel")) +
  geom_errorbar(
    aes(ymin = mean - sd, ymax = mean + sd, color = country == "Israel"),
    position = position_dodge(width = 0.5),
    width = 0.2
  ) +
  coord_flip() +
  scale_color_manual(values = c("TRUE" = "red", "FALSE" = "black"), guide = "none") +
  xlab("Country") + ylab("Mean Life Expectancy") +
  ggtitle("Mean Life Expectancy by Country (Asia)") +
  theme(legend.position = "none")
  theme_bw()
ggplotly(plotly_Israel_first_red)
```

# Make Israel first and red and flip x-axis
```{r select_data, include false}
# the same graph as above - the mean is reversed
plotly_Israel_first_red_flip <- gap_asia_continent %>%
  ungroup() %>%
  mutate(
    is_israel = ifelse(country == "Israel", 0, 1),
    country = reorder(country, -mean)
  ) %>%
  arrange(desc(is_israel), (mean)) %>%
  mutate(country = factor(country, levels = unique(country))) %>%

  ggplot(aes(x = country, y = mean)) +
  geom_point(aes(color = country == "Israel")) +
  geom_errorbar(
    aes(ymin = mean - sd, ymax = mean + sd, color = country == "Israel"),
    position = position_dodge(width = 0.5),
    width = 0.2
  ) +
  coord_flip() +
    scale_color_manual(values = c("TRUE" = "red", "FALSE" = "black"), guide = "none") +
  xlab("Country") + ylab("Mean Life Expectancy") +
  ggtitle("Mean Life Expectancy by Country (Asia)") +
  scale_y_reverse() +                 # Reverses the y-axis (mean life expectancy)
  theme(legend.position = "none")
  theme_bw()
ggplotly(plotly_Israel_first_red_flip)
```







# Divide data into middle east and rest of Asia
```{r sort_data, include false}
# Define the countries in the Middle East
middle_east_countries <- c(
  "Israel", "P.A.", "Yemen", "Saudi Arabia", "Jordan", "Lebanon",
  "Syria", "Iraq", "Iran", "Qatar", "Kuwait",
  "Bahrain", "Oman", "Afghanistan"
)
# Add dummy variable to define middle east countries and other Asian countries
gap_asia_continent <- gap_asia_continent %>%
  mutate(region_group = ifelse(country %in% middle_east_countries, 1, 2))

# dataframe of Middle East countries (region_group == 1)
middle_east <- gap_asia_continent %>%
  filter(region_group == 1)

# dataframe Other Asian countries (region_group == 2)
asia <- gap_asia_continent %>%
  filter(region_group == 2)

kable_table_middle_east <- gap_middle_east %>%
  slice_head(n = 20) %>%  # Show first 20 rows
  kable("html", align = 'c', caption = "Tibble for Peptic Ulcer data (First 20 Rows)") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    font_size = 14
  ) %>%
  scroll_box(height = "400px")
print(kable_table_middle_east)

kable_table_asia <- asia %>%
  slice_head(n = 20) %>%  # Show first 20 rows
  kable("html", align = 'c', caption = "Tibble for Peptic Ulcer data (First 20 Rows)") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    font_size = 14
  ) %>%
  scroll_box(height = "400px")
print(kable_table_asia, n=10000000)
```




# Show middle means and standard deviations
```{r select_data, include false}
# means and standard deviations for miidle east countries
plotly_middle_east <- middle_east %>%
  ungroup() %>%
  mutate(
    is_israel = ifelse(country == "Israel", 0, 1),
    country = reorder(country, -mean)
  ) %>%
  arrange(desc(is_israel), (mean)) %>%
  mutate(country = factor(country, levels = unique(country))) %>%

  ggplot(aes(x = country, y = mean)) +
  geom_point(aes(color = country == "Israel")) +
  geom_errorbar(
    aes(ymin = mean - sd, ymax = mean + sd, color = country == "Israel"),
    position = position_dodge(width = 0.5),
    width = 0.2
  ) +
  coord_flip() +
    scale_color_manual(values = c("TRUE" = "red", "FALSE" = "black"), guide = "none") +
  xlab("Country") + ylab("Mean Life Expectancy") +
  ggtitle("Life Expectancy (Middle East)") +
  scale_y_reverse() +                 # Reverses the y-axis (mean life expectancy)
  theme(legend.position = "none")
  theme_bw()
ggplotly(plotly_middle_east)
```


# Show asia means and standard deviations
```{r select_data, include false}
# means and standard deviations for asian countries
plotly_asia <- asia%>%   
  ggplot(mapping = aes(x = fct_reorder(country, mean), y = mean)) +  # reorder by mean
  geom_point() +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd),
                position = position_dodge(width = 0.5), width = 0.2) +
  coord_flip() +
  xlab("Country") + ylab("Mean Life Expectancy") +
  ggtitle("Mean Life Expectancy by Country (Asia)") +
  scale_y_reverse() +                 # Reverses the y-axis (mean life expectancy)
  theme(legend.position = "none")
  theme_bw()
ggplotly(plotly_asia)
```


# combine the two plots
```{r select_data, include false}
# combine with patchwork
plotly_middle_east |  plotly_asia
# combine with plotly
subplot(plotly_middle_east, plotly_asia, nrows = 1, margin = 0.05, shareY = FALSE, titleX = TRUE, titleY = TRUE)
```


