###
### name  : __________________
### ID    : [מספר תעודת זהות]
### course: אפידמיולוגיה (HIT.ac.il)
### lesson: בדיקת השערות 04   
### date  : 22/02/2026
###  
###  נלמד לבצע סטטיטסיקה הסקתית ומבחני השערות


## Reset memory
```{r setup-packages, include = FALSE}
rm(list = ls(all.names = TRUE))          # 👈 Clear all data objects        👇 Clear packages
lapply(paste("package:", names(sessionInfo()$otherPkgs), sep = ""), detach, character.only = TRUE, unload = TRUE, force = TRUE)
gc()                                     # Clear unused data from memory ("garbage cllection")
dev.off()                                # Reset the display output of R
system("powershell -Command Clear-Host") # This will clear the terminal in VScode
```

## Load required packages
```{r load-packages, include = FALSE}  
options("install.lock" = FALSE)
if (!require("pacman")) install.packages("pacman") ; library(pacman)
pacman::p_load(
  Rcpp, devtools, gh, remotes,   # For installing packages/files online (like from GitHub)
  rio, here,              # File import and path support
  labelled,               # deal with labels
  tidyverse,              # data management + ggplot2 graphics,
  broom,                  # For tidying model outputs
  dplyr,                  # For data manipulation
  gt,                     # Nice beuatiful tables
  gtsummary,              # summary statistics and tests
  nortest,                # Anderson-Darling test for normality
  ggplot2,                # For creating plots
  ggpubr,                 # For ggplot2 extensions
  DescTools,              # For statistical functions like skewness and kurtosis
  summarytools,           # to get a table of all summary statistics
  car,                    # For Levene's test and ANOVA
  dunn.test,              # For a Krutal-Wallis post-hoc test
  finalfit                # to get 2x2 crosstables and calculations
              )
# Install/load flexplot
if (!require("flexplot", quietly = TRUE)) devtools::install_github("dustinfife/flexplot", ref = "logistic_plots")  ;  library(flexplot)
```

## Access the data
```{r read-data, echo = TRUE}
# file <- "./EpiData/02 breastcancer.rds"     # 👈 use this if the file is on your computer
file <- "https://raw.githubusercontent.com/Model-Lab-Net/Courses/refs/heads/main/Epi/Data/02%20breastcancer.rds"
# read the data in a file online
df_bc <- import(file, trust = TRUE)
# show variable translations to Hebrew in lables
labels <- data.frame(Variable = names(df_bc), Label = unlist(var_label(df_bc)))   ;   gt(labels)
glimpse(df_bc)               # get an overview of the data
```

## Crosstables, Chi-Square and Fisher Exact tests
```{r chi-square, echo = TRUE}
# Create cross-tabulation with chi-square test
crosstable <- df_bc %>% # nolint
  finalfit("recurrence", "lvi")

crosstable %>%
  gt()
```

## T test for independent samples
```{r t-test, echo = TRUE}
# Plot the data
flexplot(age.dx.yrs ~ recurrence, data = df_bc)

# Test for variance homogeneity (שיוויון שונויות)
levene_test <- leveneTest(age.dx.yrs ~ recurrence, data = df_bc)
levene_test

# Do the independent samples T test
t_test_result <- t.test(age.dx.yrs ~ recurrence, data = df_bc, var.equal = FALSE)
t_test_result
```

## Two-way ANOVA
```{r ANOVA, echo = TRUE}
# Plot the data
flexplot(age.dx.yrs ~ grade, data = df_bc)

# Test for variance homogeneity of variances (שיוויון שונויות)
bartlett_test <- bartlett.test(tumor.size.cm ~ grade, data = df_bc)
bartlett_test

# Do the ANOVA
anova_test <- aov(tumor.size.cm ~ grade, data = df_bc)
summary(anova_test)

# Tukey HSD post-hoc test
tukey_result <- TukeyHSD(anova_test)
tukey_result
```

## Mann-Whitney test (a nonparametric "T test")
```{r M_U-test, echo = TRUE}
# Perform Mann-Whitney test, comparing numerical variable between two groups
MW_test <- wilcox.test(nodal.ratio ~ optimal.systemic.therapy, data = df_bc, na.action = na.exclude)
MW_test
```

## Kruskal-Wallis test (a nonparametric "ANOVA")
```{r M_U-test, echo = TRUE}
# Kruskal-Wallis test, comparing numerical variable between three or more groups
KW_test <- kruskal.test(nodal.ratio ~ grade, data = df_bc, na.action = na.exclude)
KW_test

# Dunn post-hoc test for Kruskal-Wallis 
dunn.test(df_bc$nodal.ratio, df_bc$grade, method = "bh")
```

# Paired T test or Wilcoxon ranked sum
```{r paired-t-test, echo = TRUE}
## Lets get some data on anorexia
file <- "https://raw.githubusercontent.com/mark-andrews/ntupsychology-data/main/data-sets/anorexia.csv"
# read the data in a file online
df_anorexia <- import(file, trust = TRUE)
glimpse(df_anorexia)

# Do a paired T test
paired_t_test_result <- t.test(df_anorexia$Prewt, df_anorexia$Postwt, paired = TRUE)
paired_t_test_result

# Plot the data
ggpaired(df_anorexia, cond1 = "Prewt", cond2 = "Postwt",
         y = "value", x = "variable", fill = "condition",
         palette = "jco", line.color = "gray", line.size = 0.4
         ) #+
         #facet_wrap(~ Treat)

# Do a paired Wilcoxon ranked test if data are not normally distributed
wilcox.test(df_anorexia$Prewt, df_anorexia$Postwt, paired = TRUE)
```

## Spearman regression
```{r spearman, echo = TRUE}
# Lets look at the data
#flexplot(tumor.size.cm ~ grade, data = df_bc)

# Spearman needs the x axis values to be int
#df_bc$grade <- as.integer(df_bc$grade)

# Spearman regrssion (better not to use it....)
#spearman <- cor.test(df_bc$tumor.size.cm, df_bc$grade, method = "spearman", use = "complete.obs")
#spearman
```




### -------------------------------------------------
## תרגול-עצמי

## שאלה 1
```{r question-1, echo = TRUE}
# 

```

## שאלה 2
```{r question-2, echo = TRUE}
# 

```

## שאלה 3
```{r question-3, echo = TRUE}
# 

```









### source: https://www.emilyzabor.com/mmedr/tools-manipulate-dataframes.html

## Clean data
```{r cleam-data, echo = TRUE}
# Load the data
file <- "https://raw.githubusercontent.com/zabore/mmedr/refs/heads/main/data/breastcancer.csv"
df_bc <- import(file, trust = TRUE)
#write.csv(df_bc, here("EpiData", "02 breastcancer.csv"))

# Clean up variable names
install.packages("janitor")   ;    library(janitor)
df_bc <- clean_names(df_bc)
names(df_bc) <- gsub("_", ".", names(df_bc))

# Recode grade from I,II,III to 1,2,3
df_bc <- df_bc %>%
  mutate(grade = case_when(
    grade == "I" ~ 1L,
    grade == "II" ~ 2L,
    grade == "III" ~ 3L,
    TRUE ~ as.integer(NA)  # Convert unexpected values to NA
  ))
df_bc$grade <- factor(df_bc$grade, levels = c("1", "2", "3"), ordered = TRUE)

# Convert to int variable
df_bc$days.onset.hosp <- as.integer(df_bc$days.onset.hosp)

# Rename variables - new_name = old_name
df_bc <- dplyr::rename(df_bc, recurrence = event)

# Convert recurrence to fct
df <- df %>%
  mutate(recurrence = factor(recurrence, levels = c(0, 1), labels = c("no", "yes")))

# Check the updated dataset structure
glimpse(df_bc)

# Save as R data file
saveRDS(df_bc, here("EpiData", "02 breastcancer.rds"))
saveRDS(df_anorexia, here("EpiData", "02 anorexia.rds"))
```

## Give variable explanations in labels
```{r labels, echo = TRUE}
attr(df_bc$time, "label") <- "מספר שנים לאחר גילוי הסרטן"
attr(df_bc$recurrence, "label") <- "הישנות הגידול"
attr(df_bc$rt, "label") <- "טיפול"
attr(df_bc$age.dx.yrs, "label") <- "גיל האבחון"
attr(df_bc$tumor.size.cm , "label") <- "גודל הגידול ב-ס'מ"
attr(df_bc$grade, "label") <- "דרגת המחלה"
attr(df_bc$n.ln.pos.3.vs.1or2, "label") <- "מספר בלוטות לימפה מעורבות"
attr(df_bc$nodal.ratio, "label") <- "מדד למעורבות בלוטות לימפה"
attr(df_bc$lvi, "label") <- "חדירת תאים סרטניים למערכת הלימפה"
attr(df_bc$er.or.pr.pos, "label") <- "תאי סרטן עם קולטנים לאסטרוגן וגם לפרוגסטרון"
attr(df_bc$her2.pos, "label") <- "תאי סרטן עם רמות גבוהות של HER2+"
attr(df_bc$quadrant.inner.vs.upper, "label") <- "מיקום הסרטן בשד (1=פנימי)"
attr(df_bc$optimal.systemic.therapy, "label") <- "טיפול סיסטמי אופטימלי"


labels <- NULL
labels <- data.frame(Variable = names(df_bc), Label = unlist(var_label(df_bc)))
gt(labels)
```