###
### name  :  专
### ID    : [住驻专 转注转 转]
### course: 驻 (HIT.ac.il)
### lesson: 专专住 专转 05   
### date  : 13/07/2025
###  
###  注砖转 专专住 专转 驻砖


## Reset memory
```{r setup-packages, include = FALSE}
rm(list = ls(all.names = TRUE))       #  Clear all data objects         Clear packages
lapply(paste("package:", names(sessionInfo()$otherPkgs), sep = ""), detach, character.only = TRUE, unload = TRUE, force = TRUE)           # Clear packages from memory
gc()                                  # Clear unused data from memory ("garbage cllection")
dev.off()                             # Reset the display output of R
q("no")                               # This will completely shutdown R terminal
```

## Load required packages
```{r load-packages, include = FALSE}
options("install.lock" = FALSE)
if (!require("pacman")) install.packages("pacman") ; library(pacman)
pacman::p_load(
  Rcpp, devtools, gh, remotes,   # For installing packages/files online (like from GitHub)
  rio, here,              # File import and path support
  labelled,               # deal with labels
  tidyverse,              # data management + ggplot2 graphics,
  broom,                  # For tidying model outputs
  dplyr,                  # For data manipulation
  gt,                     # Nice beuatiful tables
  gtsummary,              # summary statistics and tests
  nortest,                # Anderson-Darling test for normality
  ggplot2,                # For creating plots
  ggpubr,                 # For ggplot2 extensions
  DescTools,              # For statistical functions like skewness and kurtosis
  summarytools,           # to get a table of all summary statistics
  car,                    # For Levene's test and ANOVA
  lmtest,                 # do linear regression tests
  parameters              # get standardized parameters/coefficients
              )
# Install/load flexplot
if (!require("flexplot", quietly = TRUE)) devtools::install_github("dustinfife/flexplot", ref = "logistic_plots")  ;  library(flexplot)
```

## Access the data
```{r read-data, echo = TRUE}
# file <- "./EpiData/03 coronary.rds" #  use this if the file is on your computer
file <- "https://raw.githubusercontent.com/Model-Lab-Net/Courses/refs/heads/main/Epi/Data/03%20coronary.rds"
df2 <- import(file, trust = TRUE)
# show variable translations to Hebrew in lables
labels <- data.frame(Variable = names(df2), Label = unlist(var_label(df2)))   ;   gt(labels)
glimpse(df2)               # get an overview of the data
```

## Summary statistics for numerical variables
```{r summary-stats, echo = TRUE}
# Create summary statistics for all numerical variables grouped by "race"
df2_summary_stats <- df2 %>%
  select(where(is.numeric), gender) %>%                # Select all numeric columns and gender
  tbl_summary(
    by = gender,                                       # Stratify by gender
    statistic = list(
      all_continuous() ~ "{mean} ({sd})"#,              # Mean (SD) for continuous variables
      #all_continuous2() ~ "{median} ({p25}, {p75})"    # Median (Q1, Q3) for non-normal
    ),
    digits = list(all_continuous() ~ 1),               # Round to 2 decimal places
    # missing = "ifany",                               # Show missing values if any
    label = list(                                      # Optional: Customize variable labels
      # Example: age ~ "Age (years)"
    )
  ) %>%
  # add_p() %>%                                        # Add p-values for comparisons
  # add_overall() %>%                                    # Add overall column
  modify_header(label = "**砖转**")                # Customize header

# Display the summary statistics table
df2_summary_stats
```

## Choose variables
```{r choose-vars, echo = TRUE}
x_var <- "chol"   # Change this to the categorial variable
y_var <- "sbp"    # Change this to the numerical variable
f <- as.formula(paste(y_var, "~", x_var))     # This just helps further down....
```

## Simple linear regression model
```{r slr, echo = TRUE}
# Perform a Simple Linear Regression
slr <- lm(df2[[y_var]] ~ df2[[x_var]], data = df2)
slr <- lm(f, data = df2)                            # useful for later down.....

# print the summary of the simple linear regression
summary(slr)
tidy(slr)
```

## Calculate pearson correlation coefficient
```{r pearson, echo = TRUE}
# Calculate the pearson coefficient
pearson_result <- cor.test(df2[[x_var]], df2[[y_var]], method = "pearson", use = "complete.obs")
pearson_result
```

## Linear regression parameter values with nice output
```{r slr-summary, echo = TRUE}
# use gtsummary to create a nice table
slr_summary <- slr %>%
  tbl_regression(intercept = TRUE) %>%
  add_glance_table(include = c(
    nobs,
    r.squared,
    adj.r.squared,
    statistic,
    sigma,
    logLik,
    AIC
  ))

# Show the model results table
slr_summary
```

## Old way to get scatter plot with regression model
```{r polt-regression, echo = TRUE}
# use ggplot to show the scatter plot with regression line with linear model and confidence interval
ggplot(df2, aes_string(x = x_var, y = y_var)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, color = "blue") + # Adding the trend line with linear model
  labs(
    title = paste("Scatter Plot of", y_var, "vs", x_var),
    x = x_var,
    y = y_var
  ) +
  theme_minimal()
```

## Flexplot scatter plot and analysis
```{r polt-regression, echo = TRUE}
# Plot the regression
flexplot(f, data = df2, method = "lm", se = TRUE, ghost.line = "red")

# Analyze the residuals
visualize(slr)

# Get the model parameter values
estimates(slr)
```

## Test for normal distribution of model residuals
```{r residuals, echo = TRUE}
# Get nmodel residuals
res <- residuals(slr)

# Plot the residuals
hist(res,
  breaks = 20, main = "Histogram of Residuals",
  xlab = "Residuals", col = "darkgrey"
)

# Q-Q plot
qqnorm(res)
qqline(res, col = "red")

# Kolmogorov-Smirnov test  (Ho: residuals are normal)
ks_test_result <- ks.test(res, "pnorm", mean = mean(res), sd = sd(res))
ks_test_result
```

## Test for homoscedacity
```{r homoscedacity, echo = TRUE}
# Plot Residuals vs. Fitted Values Plot
# (homoscedacity = scattered randomly around the horizontal line)
# (heteroscedacity = Patterns like a funnel shape (wider spread at higher/lower fitted values), curves, or systematic trends)
plot(slr$fitted.values, residuals(slr),
  main = "Residuals vs. Fitted Values",
  xlab = "Fitted Values",
  ylab = "Residuals"
)
abline(h = 0, col = "red", lty = 2) # Add horizontal line at y = 0

# Run Breusch-Pagan test  (Ho: no heteroscedacity)
bp_test <- bptest(slr)
bp_test
```

## Test for autocorrelation 
```{r autocorrelation , echo = TRUE}
# Plot ACF residuals
acf(residuals(slr), main = "ACF of Residuals")

# Run Durbin-Watson test  (Ho: no autocorelation)
dw_result <- dwtest(slr)
dw_result
```

## Calculate the standaradized parameters
```{r std-parms , echo = TRUE}
model_parameters(slr, standardize = "refit")
```





### -------------------------------------------------
## 转专-注爪

## 砖 1
```{r question-1, echo = TRUE}
# 

```

## 砖 2
```{r question-2, echo = TRUE}
# 

```

## 砖 3
```{r question-3, echo = TRUE}
# 

```









### source: https://bookdown.org/drki_musa/dataanalysis/linear-regression.html
###         https://github.com/drkamarul/multivar_data_analysis/tree/main/data

## Clean data
```{r clean-data, echo = TRUE}
# Load the data
file <- "https://raw.githubusercontent.com/drkamarul/multivar_data_analysis/refs/heads/main/data/coronary.dta"
df2 <- import(file, trust = TRUE)
#write.csv(df2, here("EpiData", "03 coronary.csv"))

# Change id to factor
df2 <- df2 %>%
  mutate(id = as.factor(id))

# Change cad to factor
df2 <- df2 %>%
  mutate(cad = as.factor(cad))

# Change gender column from 1, 2 to "F", "M"
df2 <- df2 %>%
  mutate(gender = case_when(
    gender == 1 ~ "F",
    gender == 2 ~ "M",
    TRUE ~ as.factor(gender) # Preserve other values (e.g., NA)
  ))

# Change race column from 1, 2, 3  to "W", "B", "J"
df2 <- df2 %>%
  mutate(race = case_when(
    race == 1 ~ "W",
    race == 2 ~ "B",
    race == 3 ~ "J",
    TRUE ~ as.factor(race) # Preserve other values (e.g., NA)
  ))

# Convert all chr variables to fct
df2 <- mutate_if(df2, is.character, as.factor)

# Check the updated dataset structure
glimpse(df2)

# Save as R data file
saveRDS(df2, here("EpiData", "03 coronary.rds"))
```

## Give variable explanations in labels
```{r labels, echo = TRUE}
attr(df2$id, "label") <- "拽 驻"
attr(df2$cad, "label") <- "转 注专拽 "
attr(df2$sbp, "label") <- "抓  住住"
attr(df2$dbp, "label") <- "抓  住"
attr(df2$chol , "label") <- "住专 "
attr(df2$age, "label") <- ""
attr(df2$bmi, "label") <- "BMI"
attr(df2$race, "label") <- " / 砖专 / "
attr(df2$gender, "label") <- "专"

labels <- NULL
labels <- data.frame(Variable = names(df2), Label = unlist(var_label(df2)))
gt(labels)
```