###
### name  : __________________
### ID    : [מספר תעודת זהות]
### course: אפידמיולוגיה (HIT.ac.il)
### lesson: רגרסיה לינארית 05   
### date  : 12/04/2026
###  
### נלמד לעשות רגרסיה לינארית פשוטה
###
### source: https://bookdown.org/drki_musa/dataanalysis/linear-regression.html#simple-linear-regression


## Reset memory
```{r setup-packages, include = FALSE}
rm(list = ls(all.names = TRUE))          # 👈 Clear all data objects        👇 Clear packages
lapply(paste("package:", names(sessionInfo()$otherPkgs), sep = ""), detach, character.only = TRUE, unload = TRUE, force = TRUE)
gc()                                     # Clear unused data from memory ("garbage cllection")
dev.off()                                # Reset the display output of R
system("powershell -Command Clear-Host") # This will clear the terminal in VScode
```

## Load required packages
```{r load-packages, include = FALSE}
options("install.lock" = FALSE)
if (!require("pacman")) install.packages("pacman") ; library(pacman)
pacman::p_load(
  Rcpp, devtools, gh, remotes,   # For installing packages/files online (like from GitHub)
  rio, here,              # File import and path support
  labelled,               # deal with labels
  tidyverse,              # data management + ggplot2 graphics,
  broom,                  # For tidying model outputs
  dplyr,                  # For data manipulation
  gt,                     # Nice beuatiful tables
  gtsummary,              # summary statistics and tests
  nortest,                # Anderson-Darling test for normality
  ggplot2,                # For creating plots
  ggpubr,                 # For ggplot2 extensions
  DescTools,              # For statistical functions like skewness and kurtosis
  summarytools,           # to get a table of all summary statistics
  car,                    # For Levene's test and ANOVA
  lmtest,                 # do linear regression tests
  parameters,             # get standardized parameters/coefficients
  performance
              )
# Install/load flexplot
if (!require("flexplot", quietly = TRUE)) devtools::install_github("dustinfife/flexplot", ref = "logistic_plots")  ;  library(flexplot)
```

## Access the data
```{r read-data, echo = TRUE}
# file <- "./EpiData/03 coronary.rds" # 👈 use this if the file is on your computer
file <- "https://raw.githubusercontent.com/Model-Lab-Net/Courses/refs/heads/main/Epi/Data/03%20coronary.rds"
df2 <- import(file, trust = TRUE)
# show variable translations to Hebrew in lables
labels <- data.frame(Variable = names(df2), Label = unlist(var_label(df2)))   ;   gt(labels)
glimpse(df2)               # get an overview of the data
```

## Summary statistics for numerical variables
```{r summary-stats, echo = TRUE}
# Create summary statistics for all numerical variables grouped by "race"
summary_stats <- df2 %>%
  select(where(is.numeric), gender) %>%                # Select all numeric columns and gender
  tbl_summary(
    by = gender,                                       # Stratify by gender
    statistic = list(
      all_continuous() ~ "{mean} ({sd})"#,              # Mean (SD) for continuous variables
      #all_continuous2() ~ "{median} ({p25}, {p75})"    # Median (Q1, Q3) for non-normal
    ),
    digits = list(all_continuous() ~ 1),               # Round to 2 decimal places
    # missing = "ifany",                               # Show missing values if any
    label = list(                                      # Optional: Customize variable labels
      # Example: age ~ "Age (years)"
    )
  ) %>%
  # add_p() %>%                                        # Add p-values for comparisons
  # add_overall() %>%                                    # Add overall column
  modify_header(label = "**משתנה**")                # Customize header

# Display the summary statistics table
summary_stats
```

## Simple linear regression model
```{r slr, echo = TRUE}
# Perform a Simple Linear Regression
slr <- lm(chol ~ sbp, data = df2)

# print the summary of the simple linear regression
summary(slr)
```

## Calculate pearson correlation coefficient
```{r pearson, echo = TRUE}
# Calculate the pearson coefficient
pearson_result <- cor.test(df2$chol, df2$sbp, method = "pearson", use = "complete.obs")
pearson_result
```

## Linear regression parameter values with nice output
```{r slr-model, echo = TRUE}
slr_model <- slr %>%
  tbl_regression(intercept = TRUE) %>%
  add_glance_table(include = c(
    nobs,
    r.squared,
    adj.r.squared,
    statistic,
    sigma,
    logLik,
    AIC,
    p.value
  ))

slr_model
```

## Flexplot scatter plot and analysis
```{r polt-regression, echo = TRUE}
# Plot the regression
flexplot(chol ~ sbp, data = df2, method = "lm", se = TRUE)

# Analyze the residuals
visualize(slr)

# Get the model parameter values
estimates(slr)
```

# Diagnostics  (Ho: residuals are normal)
```{r residuals, echo = TRUE}
# Get nmodel residuals
res <- residuals(slr)
# Do the tests:
sw_test <- shapiro.test(residuals(slr))$p.value       # Shapiro-Wilkn for normality
ks_test <- nortest::lillie.test(res)$p.value                     # do _NOT_ use Kolmogorov–Smirnov test!!!
bp_test <- bptest(slr)$p.value                        # Breusch-Pagan for heteroscedacity
dw_test <- dwtest(slr)$p.value                      # Durbin-Watson test for autocorelation


res <- check_normality(slr)
res["p"]

# Add these to the model summary table
slr_model %>%
   modify_table_styling(
    column = label,
    rows = row_type == "glance_statistic",
    footnote = sprintf("Diagnostics:<br> Shapiro-Wilk (normality) = %.3f <br>Breusch-Pagan (heteroscedcity) = %.3f <br>Durbin-Watson (autocorrelation) = %.3f", 
                       sw_test, bp_test, dw_test)
  )
```

## Calculate the standaradized parameters
```{r std-parms , echo = TRUE}
model_parameters(slr, standardize = "refit")
```





### -------------------------------------------------
## תרגול-עצמי

## שאלה 1
```{r question-1, echo = TRUE}
# 

```

## שאלה 2
```{r question-2, echo = TRUE}
# 

```

## שאלה 3
```{r question-3, echo = TRUE}
# 

```









### source: https://bookdown.org/drki_musa/dataanalysis/linear-regression.html
###         https://github.com/drkamarul/multivar_data_analysis/tree/main/data

## Clean data
```{r clean-data, echo = TRUE}
# Load the data
file <- "https://raw.githubusercontent.com/drkamarul/multivar_data_analysis/refs/heads/main/data/coronary.dta"
df2 <- import(file, trust = TRUE)
#write.csv(df2, here("EpiData", "03 coronary.csv"))

# Change id to factor
df2 <- df2 %>%
  mutate(id = as.factor(id))

# Change cad to factor
df2 <- df2 %>%
  mutate(cad = as.factor(cad))

# Change gender column from 1, 2 to "F", "M"
df2 <- df2 %>%
  mutate(gender = case_when(
    gender == 1 ~ "F",
    gender == 2 ~ "M",
    TRUE ~ as.factor(gender) # Preserve other values (e.g., NA)
  ))

# Change race column from 1, 2, 3  to "W", "B", "J"
df2 <- df2 %>%
  mutate(race = case_when(
    race == 1 ~ "W",
    race == 2 ~ "B",
    race == 3 ~ "J",
    TRUE ~ as.factor(race) # Preserve other values (e.g., NA)
  ))

# Convert all chr variables to fct
df2 <- mutate_if(df2, is.character, as.factor)

# Check the updated dataset structure
glimpse(df2)

# Save as R data file
saveRDS(df2, here("EpiData", "03 coronary.rds"))
```

## Give variable explanations in labels
```{r labels, echo = TRUE}
attr(df2$id, "label") <- "קוד המטופל"
attr(df2$cad, "label") <- "מחלת עורקים כליליים"
attr(df2$sbp, "label") <- "לחץ דם סיסטולי"
attr(df2$dbp, "label") <- "לחץ דם דיאסטולי"
attr(df2$chol , "label") <- "כולסטרול בדם"
attr(df2$age, "label") <- "גיל"
attr(df2$bmi, "label") <- "BMI"
attr(df2$race, "label") <- "לבן / שלור / יהודי"
attr(df2$gender, "label") <- "מגדר"

labels <- NULL
labels <- data.frame(Variable = names(df2), Label = unlist(var_label(df2)))
gt(labels)
```