###
### name: דוד בורג
### ID  : [מספר תעודת זהות]
### course: אפידמיולוגיה (HIT.ac.il)
### lesson: רגרסיה לינארית 05   
### date  : 15/06/2025
###  
### נלמד לעשות רגרסיה לינארית פשוטה
###
### source:  https://bookdown.org/drki_musa/dataanalysis/
###          https://github.com/drkamarul/multivar_data_analysis/tree/main/data


## Reset memory
```{r setup-packages, include false}
rm(list = ls(all.names = TRUE))
lapply(paste('package:',names(sessionInfo()$otherPkgs),sep=""),detach,character.only=TRUE,unload=TRUE,force=TRUE)
gc()
dev.off()
```

## Load required packages
```{r load-packages, include false}
options("install.lock"=FALSE)
install.packages("pacman")
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  rio,          # File import
  tidyverse,    # data management + ggplot2 graphics, 
  gt,           # Nice beuatiful tables
  nortest,      # Anderson-Darling test for normality
  ggplot2,      # For creating plots
  DescTools,    # For statistical functions like skewness and kurtosis
  janitor,      # adding totals and percents to tables
  car,           # For Levene's test and ANOVA
  gtsummary,    # summary statistics and tests
  corrr        # correlation analayis for numeric variables
            )
```

## Access the data
```{r read-data, echo=FALSE}
## read the data in a file online
# file <- "./EpiData/coronary.dta"     # use this if the file is on your computer
file <- "https://raw.githubusercontent.com/drkamarul/multivar_data_analysis/refs/heads/main/data/coronary.dta"
df2 <- import(file, trust  = TRUE)
summary(df2)
glimpse(df2)
```

## Change categorial variables to factors
```{r read-data, echo=FALSE}
# Mutate id to factor
df2 <- df2 %>%
  mutate(id = as.factor(id))

# Mutate cad to factor
df2 <- df2 %>%
  mutate(cad = as.factor(cad))

# Mutate gender column from 1, 2 to "F", "M"
df2 <- df2 %>%
  mutate(gender = case_when(
    gender == 1 ~ "F",
    gender == 2 ~ "M",
    TRUE ~ as.character(gender)  # Preserve other values (e.g., NA)
  ))

# Mutate race column from 1, 2, 3  to "W", "B", "J"
df2 <- df2 %>%
  mutate(race = case_when(
    race == 1 ~ "W",
    race == 2 ~ "B",
    race == 3 ~ "J",
    TRUE ~ as.character(gender)  # Preserve other values (e.g., NA)
  ))

glimpse(df2)
```

# Summary statistics for numerical variables
```{r sort_data, include false}
# Show median and IQR for all numerical variables (5)
df2_summary_stats <- df2 %>%
select(where(is.numeric), race) %>%  # Select all numeric columns and gender
  tbl_summary(
    by = race,                        # Stratify by gender
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",  # Mean (SD) for continuous variables
      all_continuous2() ~ "{median} ({p25}, {p75})"  # Median (Q1, Q3) for non-normal
    ),
    digits = list(all_continuous() ~ 2),  # Round to 2 decimal places
    #missing = "ifany",                    # Show missing values if any
    label = list(                         # Optional: Customize variable labels
      # Example: age ~ "Age (years)"
    )
  ) %>%
  #add_p() %>%                           # Add p-values for comparisons
  add_overall() %>%                     # Add overall column
  modify_header(label = "**Variable**") # Customize header

df2_summary_stats
```

## Choose variables
```{r pearson, include=FALSE}
###pearson
x_var <- "chol"  # Change this to the categorial variable
y_var <- "sbp"   # Change this to the numerical variable
```

## Calculate pearson correlation coefficient
```{r pearson, include=FALSE}
###pearson
pearson_result <- cor.test(df2[[x_var]], df2[[y_var]], method = "pearson", use = "complete.obs")
pearson_result
```

## Linear regression
```{r slr, include=FALSE}
# Perform a Simple Linear Regression
slr <- lm(df2[[y_var]] ~ df2[[x_var]], data = df2)

# print the summary of the simple linear regression
summary(slr)
```

## Linear regression with nice output
```{r slr-gtsummary, include=FALSE}
# use gtsummary to create a nice table
slr_summary <-
  lm(df2[[y_var]] ~ df2[[x_var]]) %>% 
  tbl_regression() %>%
  add_glance_table(include = c(nobs,
                               r.squared,
                               statistic,
                               logLik,
                               AIC,
                               sigma,
                               p.value))

# Show the table
slr_summary
```

## Show the scatter plot with regression line
```{r slr-gtsummary, include=FALSE}
# use ggplot to show the scatter plot with regression line with linear model and confidence interval
ggplot(df2, aes_string(x = x_var, y = y_var)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, color = "blue") +  # Adding the trend line with linear model
  labs(title = paste("Scatter Plot of", y_var, "vs", x_var),
       x = x_var,
       y = y_var) +
  theme_minimal()
```


### ------------------------------------


## Question 2
```{r Q2, include=FALSE}
model2 <- lm(weight1 ~ newborn_gender, data = data)

summary_model2 <- summary(model2)
summary_model2
```


## Question 3
```{r Q3, include=FALSE}
install.packages("lm.beta")
library(lm.beta)

###colinearity test
test2 <- cor.test(data$height4,  data$newborn_gender , method = "spearman", use = "complete.obs")
test3 <- cor.test(data$height4,  data$mother_weight , method = "spearman", use = "complete.obs")
test4 <- cor.test(data$mother_weight,  data$newborn_gender , method = "spearman", use = "complete.obs")
test2
test3
test4
```