###
### name:  专
### ID  : [住驻专 转注转 转]
### course: 驻 (HIT.ac.il)
### lesson: 专专住 专转 05   
### date  : 03/07/2025
###  
###  注砖转 专专住 专转 驻砖
###
### source: https://bookdown.org/drki_musa/dataanalysis/linear-regression.html
###         https://github.com/drkamarul/multivar_data_analysis/tree/main/data


## Reset memory
```{r setup-packages, include = FALSE}
rm(list = ls(all.names = TRUE))
lapply(paste("package:", names(sessionInfo()$otherPkgs), sep = ""), detach, character.only = TRUE, unload = TRUE, force = TRUE)
gc()
dev.off()
```

## Load required packages
```{r load-packages, include = FALSE}
options("install.lock" = FALSE)
install.packages("pacman")
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  rio,                    # File import
  tidyverse,              # data management + ggplot2 graphics,
  broom,                  # For tidying model outputs
  dplyr,                  # For data manipulation
  gt,                     # Nice beuatiful tables
  nortest,                # Anderson-Darling test for normality
  ggplot2,                # For creating plots
  DescTools,              # For statistical functions like skewness and kurtosis
  psych,                  # to get a table of all summary statistics
  janitor,                # adding totals and percents to tables
  car,                    # For Levene's test and ANOVA
  gtsummary,              # summary statistics and tests
  corrr,                  # correlation matrix
  lmtest,                 # do linear regression tests
  parameters              # get standardized parameters/coefficients
              )
# Install/load flexplot
if (!require("devtools")) devtools::install_github("dustinfife/flexplot", ref = "development")  ;  if (!require("pacman")) install.packages("pacman")  ;  library(flexplot)
```

## Access the data
```{r read-data, echo = TRUE}
# file <- "./EpiData/coronary.dta" #  use this if the file is on your computer
file <- "https://raw.githubusercontent.com/drkamarul/multivar_data_analysis/refs/heads/main/data/coronary.dta"
df2 <- import(file, trust = TRUE)
summary(df2)
glimpse(df2)
```

## Clean data
```{r clean-data, echo = TRUE}
# Change id to factor
df2 <- df2 %>%
  mutate(id = as.factor(id))

# Change cad to factor
df2 <- df2 %>%
  mutate(cad = as.factor(cad))

# Change gender column from 1, 2 to "F", "M"
df2 <- df2 %>%
  mutate(gender = case_when(
    gender == 1 ~ "F",
    gender == 2 ~ "M",
    TRUE ~ as.factor(gender) # Preserve other values (e.g., NA)
  ))

# Change race column from 1, 2, 3  to "W", "B", "J"
df2 <- df2 %>%
  mutate(race = case_when(
    race == 1 ~ "W",
    race == 2 ~ "B",
    race == 3 ~ "J",
    TRUE ~ as.factor(race) # Preserve other values (e.g., NA)
  ))

df <- mutate_if(df2, is.character, as.factor)

# Look at the variables list
glimpse(df2)
```

# Summary statistics for numerical variables
```{r summary-stats, echo = TRUE}
# Create summary statistics for all numerical variables grouped by "race"
df2_summary_stats <- df2 %>%
  select(where(is.numeric), gender) %>%                # Select all numeric columns and gender
  tbl_summary(
    by = gender,                                       # Stratify by gender
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",              # Mean (SD) for continuous variables
      all_continuous2() ~ "{median} ({p25}, {p75})"    # Median (Q1, Q3) for non-normal
    ),
    digits = list(all_continuous() ~ 2),               # Round to 2 decimal places
    # missing = "ifany",                               # Show missing values if any
    label = list(                                      # Optional: Customize variable labels
      # Example: age ~ "Age (years)"
    )
  ) %>%
  # add_p() %>%                                        # Add p-values for comparisons
  add_overall() %>%                                    # Add overall column
  modify_header(label = "**Variable**")                # Customize header

# Display the summary statistics table
df2_summary_stats
```

## Choose variables
```{r choose-vars, echo = TRUE}
x_var <- "chol"   # Change this to the categorial variable
y_var <- "sbp"    # Change this to the numerical variable
f <- as.formula(paste(y_var, "~", x_var))     # This just helps further down....
```

## Linear regression
```{r slr, echo = TRUE}
# Perform a Simple Linear Regression
slr <- lm(df2[[y_var]] ~ df2[[x_var]], data = df2)
slr <- lm(f, data = df2)                            # useful for later down.....

# print the summary of the simple linear regression
summary(slr)
tidy(slr)
```

## Calculate pearson correlation coefficient
```{r pearson, echo = TRUE}
# Calculate the pearson coefficient
pearson_result <- cor.test(df2[[x_var]], df2[[y_var]], method = "pearson", use = "complete.obs")
pearson_result
```

## Linear regression with nice output
```{r slr-summary, echo = TRUE}
# use gtsummary to create a nice table
slr_summary <- slr %>%
  tbl_regression(intercept = TRUE) %>%
  add_glance_table(include = c(
    nobs,
    r.squared,
    adj.r.squared,
    statistic,
    sigma,
    logLik,
    AIC
  ))

# Show the model results table
slr_summary
```

## Show the scatter plot with regression line
```{r polt-regression, echo = TRUE}
# use ggplot to show the scatter plot with regression line with linear model and confidence interval
ggplot(df2, aes_string(x = x_var, y = y_var)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, color = "blue") + # Adding the trend line with linear model
  labs(
    title = paste("Scatter Plot of", y_var, "vs", x_var),
    x = x_var,
    y = y_var
  ) +
  theme_minimal()
```

## Flexplot scatter plot
```{r polt-regression, echo = TRUE}
# Plot the regression
flexplot(f, data = df2, method = "lm", se = TRUE, ghost.line = "red")

# Analyze the residuals
visualize(slr)

# Get the model parameter values
estimates(slr)
```

## Test for normal distribution of residuals
```{r residuals, echo = TRUE}
# Get nmodel residuals
residuals_slr <- residuals(slr)

# Plot the residuals
hist(residuals_slr,
  breaks = 20, main = "Histogram of Residuals",
  xlab = "Residuals", col = "darkgrey"
)

# Q-Q plot
qqnorm(residuals_slr)
qqline(residuals_slr, col = "red")

# Kolmogorov-Smirnov test  (Ho: residuals are normal)
ks_test_result <- ks.test(residuals_slr, "pnorm", mean = mean(residuals_slr), sd = sd(residuals_slr))
ks_test_result
```

## Test for homoscedacity
```{r homoscedacity, echo = TRUE}
# Run Breusch-Pagan test  (Ho: no heteroscedacity)
bp_test <- bptest(slr)
bp_test

# Plot Residuals vs. Fitted Values Plot
# (homoscedacity = scattered randomly around the horizontal line)
# (heteroscedacity = Patterns like a funnel shape (wider spread at higher/lower fitted values), curves, or systematic trends)
plot(slr$fitted.values, residuals(slr),
  main = "Residuals vs. Fitted Values",
  xlab = "Fitted Values",
  ylab = "Residuals"
)
abline(h = 0, col = "red", lty = 2) # Add horizontal line at y = 0
```

## Test for autocorrelation 
```{r autocorrelation , echo = TRUE}
# Run Durbin-Watson test  (Ho: no autocorelation)
dw_result <- dwtest(slr)
dw_result

# Plot ACF residuals
acf(residuals(slr), main = "ACF of Residuals")
```

## Calculate the standaradized parameters
```{r autocorrelation , echo = TRUE}
model_parameters(slr, standardize = "refit")
```




### ------------------------------------




## Question 2
```{r Q2, include=FALSE}
model2 <- lm(weight1 ~ newborn_gender, data = data)

summary_model2 <- summary(model2)
summary_model2
```


## Question 3
```{r Q3, include=FALSE}
install.packages("lm.beta")
library(lm.beta)

### Pearson coefficient
test2 <- cor.test(data$height4, data$newborn_gender, method = "spearman", use = "complete.obs")
test3 <- cor.test(data$height4, data$mother_weight, method = "spearman", use = "complete.obs")
test4 <- cor.test(data$mother_weight, data$newborn_gender, method = "spearman", use = "complete.obs")
test2
test3
test4
```