###
### name: דוד בורג
### ID  : [מספר תעודת זהות]
### course: אפידמיולוגיה (HIT.ac.il)
### lesson: סרגרסיה לינארית 05   
### date  : 15/06/2025
###  
###  מלמד לעשות רגסיה מרובה
###


## Reset memory
```{r setup-packages, include false}
rm(list = ls(all.names = TRUE))
lapply(paste('package:',names(sessionInfo()$otherPkgs),sep=""),detach,character.only=TRUE,unload=TRUE,force=TRUE)
gc()
```

## Load required packages
```{r load-packages, include false}  
options("install.lock"=FALSE)
install.packages("pacman")
library(pacman)
pacman::p_load(
  rio,          # File import
  here,         # File locator
  skimr,        # get overview of data
  tidyverse,    # data management + ggplot2 graphics, 
  gtsummary,    # summary statistics and tests
  rstatix,      # statistics
  nortest,      # Anderson-Darling test for normality
  moments,      # skewness and kurtosis
  corrr,        # correlation analayis for numeric variables
  janitor,      # adding totals and percents to tables
  flextable,    # converting tables to HTML
  dplyr,        # For data manipulation
  ggplot2,      # For creating plots
  DescTools,    # For statistical functions like skewness and kurtosis
  car,          # For Levene's test and ANOVA
  vctrs,        # For data type handling
  reshape2     # For reshaping data
            )
```

## Access the data
```{r read-data, echo=FALSE}
## read the data in a file online
# file <- "coronary.dta"     # use this if the file is on your computer
file <- "https://raw.githubusercontent.com/drkamarul/multivar_data_analysis/refs/heads/main/data/coronary.dta"
df2 <- import(file, trust  = TRUE)
summary(df2)
glimpse(df2)
```

## Change categorial variables to factors
```{r read-data, echo=FALSE}
# Mutate id to factor
df2 <- df2 %>%
  mutate(id = as.factor(id))

# Mutate cad to factor
df2 <- df2 %>%
  mutate(cad = as.factor(cad))

# Mutate gender column from 1, 2 to "F", "M"
df2 <- df2 %>%
  mutate(gender = case_when(
    gender == 1 ~ "F",
    gender == 2 ~ "M",
    TRUE ~ as.character(gender)  # Preserve other values (e.g., NA)
  ))

# Mutate race column from 1, 2,3  to "W", "B", "J"
df2 <- df2 %>%
  mutate(race = case_when(
    race == 1 ~ "W",
    race == 2 ~ "B",
    race == 3 ~ "J",
    TRUE ~ as.character(gender)  # Preserve other values (e.g., NA)
  ))

glimpse(df2)  
```

# Multi-variable regression with lm
```{r num-to-cat, include false}
# Do a simple linear regression (SLR) with chol as the dependent variable 
mlr <- lm(chol ~ sbp + dbp + bmi + age + race, data = df2)
summary(mlr)
```

## Correlation matrix
```{r pearson, include=FALSE}
# Select numeric variables
numeric_vars <- df2 %>%
  select(where(is.numeric))

# Compute correlation matrix
corr_matrix <- cor(numeric_vars, use = "complete.obs")

# Melt the matrix
corr_melted <- melt(corr_matrix)

# Filter to keep only upper triangle
corr_melted <- corr_melted %>%
  filter(as.numeric(Var1) < as.numeric(Var2))

# Plot
ggplot(corr_melted, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "#ff0000", high = "#0000ff", mid = "#ffffff", 
                       midpoint = 0, limit = c(-1, 1), name = "Correlation") +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        axis.title = element_blank()) +
  geom_text(aes(label = round(value, 2)), color = "black", size = 3) +
  coord_fixed()
```

## Perform a multiple linear regression
```{r pearson, include=FALSE}
### Colinearity test
mlr <- lm(chol ~ dbp + bmi + age, data = df2)

# Show results
summary(mlr)

mlr_summary <- tbl_regression(
  mlr,
  intercept = TRUE,
  estimate_fun = ~round(.x, 2)
) %>%
  add_glance_table(
    include = c(r.squared, adj.r.squared, AIC, p.value)
  )
mlr_summary
```

## Test for multicolinearity
```{r pearson, include=FALSE}
### Run VIF colinearity test
mlr_vif <- vif(mlr)
mlr_vif
```

## Show multiple linear regression summary table
```{r pearson, include=FALSE}
mlr_summary <- tbl_regression(mlr,
                             #label = list(dbp = "Diastolic Blood Pressure"),
                             conf.int = TRUE,
                             show_single_row = "dbp",
                             intercept = TRUE) %>%
  modify_header(label = "**coeff**",
                estimate = "**beta**",
                std.error = "**SE**",
                statistic = "**F**",
                p.value = "**p**",
                conf.low = "**CI<sub>95%**") %>%
  #modify_table_body(~ .x %>% filter(label != "Abbreviations")) %>%
  add_glance_source_note(
    include = c(r.squared,
                adj.r.squared,
                statistic,
                p.value,
                #logLik,
                AIC
                #sigma
                ),
    label = list(
      r.squared = "R²",
      adj.r.squared = "Adj.R²",
      statistic = "F",
      p.value = "p",
      logLik = "log-L",
      sigma = "\U03C3",
      AIC = "AIC"
    )
  ) %>%
  modify_caption("**Table 2: Model Coefficients<br>(fit statistics)**")

mlr_summary
```