###
### name  : __________________
### ID    : [מספר תעודת זהות]
### course: אפידמיולוגיה (HIT.ac.il)
### lesson: רגרסיה לינארית 05   
### date  : 12/04/2026
###  
### נלמד לעשות רגרסיה לינארית פשוטה
###
### source: https://bookdown.org/drki_musa/dataanalysis/linear-regression.html#simple-linear-regression


## Reset memory
```{r setup-packages, include = FALSE}
rm(list = ls(all.names = TRUE))          # 👈 Clear all data objects        👇 Clear packages
lapply(paste("package:", names(sessionInfo()$otherPkgs), sep = ""), detach, character.only = TRUE, unload = TRUE, force = TRUE)
gc()                                     # Clear unused data from memory ("garbage cllection")
dev.off()                                # Reset the display output of R
system("powershell -Command Clear-Host") # This will clear the terminal in VScode
```

## Load required packages
```{r load-packages, include = FALSE}
options("install.lock" = FALSE)
if (!require("pacman")) install.packages("pacman") ; library(pacman)
pacman::p_load(
  Rcpp, devtools, gh, remotes,   # For installing packages/files online (like from GitHub)
  rio, here,              # File import and path support
  labelled,               # deal with labels
  tidyverse,              # data management + ggplot2 graphics,
  broom,                  # For tidying model outputs
  dplyr,                  # For data manipulation
  gt,                     # Nice beuatiful tables
  gtsummary,              # summary statistics and tests
  nortest,                # Anderson-Darling test for normality
  ggplot2,                # For creating plots
  ggpubr,                 # For ggplot2 extensions
  DescTools,              # For statistical functions like skewness and kurtosis
  summarytools,           # to get a table of all summary statistics
  car,                    # For Levene's test and ANOVA
  lmtest,                 # do linear regression tests
  parameters,             # get standardized parameters/coefficients
  performance,
  GGally
              )
# Install/load flexplot
if (!require("flexplot", quietly = TRUE)) devtools::install_github("dustinfife/flexplot", ref = "logistic_plots")  ;  library(flexplot)
```

## Access the data
```{r read-data, echo = TRUE}
# file <- "./EpiData/03 coronary.rds" # 👈 use this if the file is on your computer
file <- "https://raw.githubusercontent.com/Model-Lab-Net/Courses/refs/heads/main/Epi/Data/03%20coronary.rds"
df2 <- import(file, trust = TRUE)
# show variable translations to Hebrew in lables
labels <- data.frame(Variable = names(df2), Label = unlist(var_label(df2)))   ;   gt(labels)
glimpse(df2)               # get an overview of the data
```

## Simple linear regression model
```{r slr, echo = TRUE}
# Perform a Simple Linear Regression
slr <- lm(chol ~ sbp, data = df2)
# print the summary of the simple linear regression
summary(slr)
```

## Multivariate linear regression model
```{r mlr, echo = TRUE}
# Perform a Simple Linear Regression
mlr <- lm(chol ~ sbp + race + bmi + age + dbp, data = df2)
# print the summary of the simple linear regression
summary(mlr)
```

## Model parameter values with nice output
```{r slr-model, echo = TRUE}
mlr_model <- mlr %>%
  tbl_regression(intercept = TRUE) %>%
  add_glance_source_note(
    include = c(r.squared, adj.r.squared, AIC, p.value)
  )

mlr_model
```

## Flexplot scatter plot and analysis
```{r polt-regression, echo = TRUE}
# Plot the regression
flexplot(chol ~ dbp + bmi | age + race, data = df2, method = "lm", se = TRUE)

# Analyze the residuals
visualize(mlr)

# Get the model parameter values
estimates(mlr)
```

## Diagnostics  (Ho: residuals are normal)
```{r residuals, echo = TRUE}
# Get nmodel residuals
res <- residuals(mlr)
# Do the tests:
sw_test <- shapiro.test(residuals(mlr))$p.value       # Shapiro-Wilkn for normality
ks_test <- nortest::lillie.test(res)$p.value                     # do _NOT_ use Kolmogorov–Smirnov test!!!
bp_test <- bptest(mlr)$p.value                        # Breusch-Pagan for heteroscedacity
dw_test <- dwtest(mlr)$p.value                      # Durbin-Watson test for autocorelation

# Add these to the model summary table
mlr_model %>%
  as_gt() %>% # Convert to gt engine first
  tab_source_note(
    source_note = html(sprintf(
      "Shapiro-Wilk (Normality): p = %.3f <br>
       Breusch-Pagan (Heteroscedasticity): p = %.3f <br>
       Durbin-Watson (Autocorrelation): p = %.3f", 
      sw_test, bp_test, dw_test
    ))
  )
```




### -------------------------------------------------
## תרגול-עצמי

## שאלה 1
```{r question-1, echo = TRUE}
# 

```

## שאלה 2
```{r question-2, echo = TRUE}
# 

```

## שאלה 3
```{r question-3, echo = TRUE}
# 

```









### source: https://bookdown.org/drki_musa/dataanalysis/linear-regression.html#multiple-linear-regression
###         https://github.com/drkamarul/multivar_data_analysis/tree/main/data

## Clean data
```{r clean-data, echo = TRUE}
# Load the data
file <- "https://raw.githubusercontent.com/drkamarul/multivar_data_analysis/refs/heads/main/data/coronary.dta"
df2 <- import(file, trust = TRUE)
#write.csv(df2, here("EpiData", "03 coronary.csv"))

# Change id to factor
df2 <- df2 %>%
  mutate(id = as.factor(id))

# Change cad to factor
df2 <- df2 %>%
  mutate(cad = as.factor(cad))

# Change gender column from 1, 2 to "F", "M"
df2 <- df2 %>%
  mutate(gender = case_when(
    gender == 1 ~ "F",
    gender == 2 ~ "M",
    TRUE ~ as.factor(gender) # Preserve other values (e.g., NA)
  ))

# Change race column from 1, 2, 3  to "W", "B", "J"
df2 <- df2 %>%
  mutate(race = case_when(
    race == 1 ~ "W",
    race == 2 ~ "B",
    race == 3 ~ "J",
    TRUE ~ as.factor(race) # Preserve other values (e.g., NA)
  ))

# Convert all chr variables to fct
df2 <- mutate_if(df2, is.character, as.factor)

# Check the updated dataset structure
glimpse(df2)

# Save as R data file
saveRDS(df2, here("EpiData", "03 coronary.rds"))
```

