###
### name  : __________________
### ID    : [מספר תעודת זהות]
### course: אפידמיולוגיה (HIT.ac.il)
### lesson: רגרסיה לינארית 05   
### date  : 19/08/2025
###  
###  נלמד לעשות רגסיה עם אינטראקציה
###
### source: https://bookdown.org/drki_musa/dataanalysis/linear-regression.html#interaction


## Reset memory
```{r setup-packages, include = FALSE}
rm(list = ls(all.names = TRUE))          # 👈 Clear all data objects        👇 Clear packages
lapply(paste("package:", names(sessionInfo()$otherPkgs), sep = ""), detach, character.only = TRUE, unload = TRUE, force = TRUE)
gc()                                     # Clear unused data from memory ("garbage cllection")
dev.off()                                # Reset the display output of R
system("powershell -Command Clear-Host") # This will clear the terminal in VScode
```

## Load required packages
```{r setup-packages, include = FALSE}
options("install.lock" = FALSE)
if (!require("pacman")) install.packages("pacman") ; library(pacman)
pacman::p_load(
  Rcpp, devtools, gh, remotes,   # For installing packages/files online (like from GitHub)
  rio, here,              # File import and path support
  labelled,               # deal with labels
  tidyverse,              # data management + ggplot2 graphics,
  broom,                  # For tidying model outputs
  dplyr,                  # For data manipulation
  gt,                     # Nice beuatiful tables
  gtsummary,              # summary statistics and tests
  nortest,                # Anderson-Darling test for normality
  ggplot2,                # For creating plots
  ggpubr,                 # For ggplot2 extensions
  DescTools,              # For statistical functions like skewness and kurtosis
  summarytools,           # to get a table of all summary statistics
  car,                    # For Levene's test and ANOVA
  lmtest,                 # do linear regression tests
  parameters,             # get standardized parameters/coefficients
  performance,
  GGally                  # to make a correlation matrix
              )
# Install/load flexplot
if (!require("flexplot", quietly = TRUE)) devtools::install_github("dustinfife/flexplot", ref = "logistic_plots")  ;  library(flexplot)
```

## Access the data
```{r read-data, echo = TRUE}
#file <- "./EpiData/03 covid.rds"     # 👈 use this if the file is on your computer
file <- "https://raw.githubusercontent.com/Model-Lab-Net/Courses/refs/heads/main/Epi/Data/03%20covid.rds"
df_covid <- import(file, trust = TRUE)
# show variable translations to Hebrew in lables
labels <- data.frame(Variable = names(df_covid), Label = unlist(var_label(df_covid)))   ;   gt(labels)
glimpse(df_covid)               # get an overview of the data
```

## Histogram of sympotms
```{r read-data, echo = TRUE}
flexplot(symptoms ~ 1, data = df_covid)
```

## Correlation matrix
```{r correlation-matrix, echo = TRUE}
ggcorr(df_covid, label = TRUE)
```

# Simple regression model
```{r multi-reg, echo = TRUE}
# Do a reduced linear regression (SLR) with chol as the dependent variable
slr <- lm(symptoms ~ bmi, data = df_covid)

# Show results in a nice table
slr_model <- slr %>%
  tbl_regression(intercept = TRUE) %>%
  add_glance_source_note(
    include = c(r.squared, adj.r.squared, AIC, p.value)
  )

slr_model

# Lets check out the plot
flexplot(symptoms ~ bmi, data = df_covid, method = "lm")
```

# Multi-variate regression model (no interaction)
```{r multi-reg, echo = TRUE}
# Do a MLR 
mlr <- lm(symptoms ~ bmi + age + copd, data = df_covid)

# Show results in a nice table
mlr_model <- mlr %>%
  tbl_regression(intercept = TRUE) %>%
  add_glance_source_note(
    include = c(r.squared, adj.r.squared, AIC, p.value)
  )

mlr_model

# Lets check out the plot
flexplot(symptoms ~ bmi + copd + age , data = df_covid, method = "lm")
```

# Multi-variate regression model with interaction /2/
```{r multi-reg, echo = TRUE}
# Do a MLR with interaction
mlrIA <- lm(symptoms ~ bmi + copd + bmi*copd, data = df_covid)

# Show results in a nice table
mlrIA_model <- mlrIA %>%
  tbl_regression(intercept = TRUE) %>%
  add_glance_source_note(
    include = c(r.squared, adj.r.squared, AIC, p.value)
  )

mlrIA_model

# Lets check out the plot
flexplot(symptoms ~ bmi + copd | age , data = df_covid, method = "lm")
```


## Model comparison
```{r mdodel-comparison, echo = TRUE}
# Check all combinations to select "best" model
model.comparison(slr, mlr)
model.comparison(slr, mlrIA)
model.comparison(mlr, mlrIA)
```


## Diagnostics  (Ho: residuals are normal)
```{r residuals, echo = TRUE}
# Get nmodel residuals
res <- residuals(mlrIA)
# Do the tests:
sw_test <- shapiro.test(residuals(mlr))$p.value       # Shapiro-Wilkn for normality
ks_test <- nortest::lillie.test(res)$p.value                     # do _NOT_ use Kolmogorov–Smirnov test!!!
bp_test <- bptest(mlr)$p.value                        # Breusch-Pagan for heteroscedacity
dw_test <- dwtest(mlr)$p.value                      # Durbin-Watson test for autocorelation

# Add these to the model summary table
mlrIA_model %>%
  as_gt() %>% # Convert to gt engine first
  tab_source_note(
    source_note = html(sprintf(
      "Shapiro-Wilk (Normality): p = %.3f <br>
       Breusch-Pagan (Heteroscedasticity): p = %.3f <br>
       Durbin-Watson (Autocorrelation): p = %.3f", 
      sw_test, bp_test, dw_test
    ))
  )
```





### -------------------------------------------------
## תרגול-עצמי

## שאלה 1
```{r question-1, echo = TRUE}
# 

```

## שאלה 2
```{r question-2, echo = TRUE}
# 

```

## שאלה 3
```{r question-3, echo = TRUE}
# 

```









### source: 
###         

## Make party enjoyment data
```{r make-data, echo = TRUE}
set.seed(123)
n <- 200  

# personality
personality <- sample(c("introvert", "extrovert"), n, replace = TRUE, prob = c(0.4, 0.6))

# party size
party.size <- ifelse(personality == "introvert",
                     rpois(n, lambda = 5) + 2,
                     rpois(n, lambda = 15) + 5)

# music
# music <- ifelse(runif(n) < plogis((party.size - 8)/3 + ifelse(party.size=="extrovert", 0.5, -0.5)), "yes", "no")
music <- sample(c("yes", "no"), n, replace = TRUE)
music <- factor(music, levels = c("no", "yes"))

# enjoyment (dependent variable)
enjoyment <- ifelse(personality == "introvert",
                    8 - 0.3 * (party.size - 5),   # introverts: lower enjoyment as size ↑
                    5 + 0.3 * (party.size - 10))  # extroverts: higher enjoyment as size ↑

# music effect: +1 if yes
enjoyment <- enjoyment + ifelse(music == "yes", 1, -0.5)

# add random noise
enjoyment <- enjoyment + rnorm(n, mean = 0, sd = 1)

# keep within 1–10 range
enjoyment <- pmin(pmax(enjoyment, 1), 10)

# dataframe
df_party <- data.frame(
  party.size = party.size,
  personality = factor(personality),
  music = factor(music),
  enjoyment = enjoyment
)

df_party <- df_party %>% filter(enjoyment < 10)

glimpse(df_party)
```

flexplot(enjoyment ~ 1, data = df_party, method = "lm")
flexplot(enjoyment ~ party.size, data = df_party, method = "lm")
flexplot(enjoyment ~ party.size | music , data = df_party, method = "lm")
flexplot(enjoyment ~ party.size | personality + music , data = df_party, method = "lm")

slr <- lm(enjoyment ~ party.size, data = df_party)
summary(slr)

mlr1 <- lm(enjoyment ~ party.size + music, data = df_party)
summary(mlr1)

mlr2 <- lm(enjoyment ~ party.size + music + personality, data = df_party)
summary(mlr2)

mlr3 <- lm(enjoyment ~ music + party.size:personality, data = df_party)
summary(mlr3)



TABLE <- mlr3 %>%
  tbl_regression(intercept = T,
                 estimate_fun = function(x) style_sigfig(x, digits = 3),
                 pvalue_fun   = function(x) style_pvalue(x, digits = 3),
                               ) %>%
  #add_global_p(keep = T) %>% 
  modify_caption("Interaction effect on enjoyment")
TABLE


#write.csv(df_party, here("EpiData", "03 party.csv"))

## Cleanup data
```{r clean-data, echo = TRUE}
# Clean up variable names
install.packages("janitor")   ;    library(janitor)
df_party <- clean_names(df_party)
names(df_party) <- gsub("_", ".", names(df_party))

glimpse(df_party)

saveRDS(df_party, here("EpiData", "03 party.rds"))
```

## Give variable explanations in labels
```{r labels, echo = TRUE}
attr(df_party$party.size, "label") <- "מספר חוגגים במסיבה"
attr(df_party$personality, "label") <- "מאפיין אישיות"
attr(df_party$music, "label") <- "האם הייתה מוסיקה במסיבה"
attr(df_party$enjoyment, "label") <- "מידת ההנאה במסביה"

labels <- NULL
labels <- data.frame(Variable = names(df_party), Label = unlist(var_label(df_party)))
gt(labels)
```


## Make covid data
```{r analysis, echo = TRUE}

flexplot(symptoms ~ 1, data = df_covid)
flexplot(symptoms ~ bmi, data = df_covid, method = "lm")
flexplot(symptoms ~ bmi | age , data = df_covid, method = "lm")
flexplot(symptoms ~ bmi | copd , data = df_covid, method = "lm")
flexplot(symptoms ~ bmi | copd + age , data = df_covid, method = "lm")


slr <- lm(symptoms ~ bmi, data = df_covid)
summary(slr)

mlr1 <- lm(symptoms ~ bmi + music, data = df_covid)
summary(mlr1)

mlr2 <- lm(symptoms ~ bmi + music + personality, data = df_covid)
summary(mlr2)

mlr3 <- lm(symptoms ~ age + bmi:copd, data = df_covid)
summary(mlr3)



TABLE <- mlr3 %>%
  tbl_regression(intercept = T,
                 estimate_fun = function(x) style_sigfig(x, digits = 3),
                 pvalue_fun   = function(x) style_pvalue(x, digits = 3),
                               ) %>%
  #add_global_p(keep = T) %>% 
  modify_caption("Interaction effect on enjoyment")
TABLE


#write.csv(df_party, here("EpiData", "03 party.csv"))
```

## Make covid data
```{r covid-data, echo = TRUE}
# Copy
df_covid <- df_party

# Remane variables
df_covid <- df_covid %>%
  rename(
    bmi = party.size,
    copd = personality,
    age = music,
    symptoms = enjoyment
  )

df_covid <- df_covid %>%
  mutate(copd = ifelse(copd == "extrovert", "yes", "no")) %>%
  mutate(copd = factor(copd, levels = c("yes", "no")))

df_covid <- df_covid %>%
  mutate(age = ifelse(age == "no", "<60", "60+")) %>%
  mutate(age = factor(age, levels = c("<60", "60+")))

glimpse(df_covid)

saveRDS(df_covid, here("EpiData", "03 covid_1.rds"))
```


## Give variable explanations in labels
```{r labels, echo = TRUE}
attr(df_covid$bmi, "label") <- "עודף משקל"
attr(df_covid$copd, "label") <- "מחלת ראיה חסימתית"
attr(df_covid$age, "label") <- "גיל"
attr(df_covid$symptoms, "label") <- "עוצמת סימפטומים"


# Convert 
df_covid <- df_covid %>%
  mutate(
    # Convert age factor to double
    age = case_when(
      age == "<60" ~ 40,
      age == "60+" ~ 65,
      TRUE ~ NA_real_ 
    ),
    # Convert copd factor to double (0/1)
    copd = case_when(
      copd == "no"  ~ 0L,
      copd == "yes" ~ 1L,
      TRUE ~ NA_integer_
    )
  )

df_covid_1 <- df_covid
glimpse(df_covid_1)
saveRDS(df_covid_1, here("EpiData", "03 covid_1.rds"))


labels <- NULL
labels <- data.frame(Variable = names(df_covid), Label = unlist(var_label(df_covid)))
gt(labels)
```