###
### name:  专
### ID  : [住驻专 转注转 转]
### course: 驻 (HIT.ac.il)
### lesson: 专专住 住转 06   
### date  : 15/06/2025
###  
###   注砖转 专住 住转
###
### source: https://github.com/drkamarul/multivar_data_analysis/tree/main/data


## Reset memory
```{r setup-packages, include = FALSE}
rm(list = ls(all.names = TRUE))
lapply(paste('package:',names(sessionInfo()$otherPkgs),sep=""),detach,character.only=TRUE,unload=TRUE,force=TRUE)
gc()
dev.off()
```

## Load required packages
```{r load-packages, include = FALSE}
options("install.lock"=FALSE)
install.packages("pacman")
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  rio,          # File import
  tidyverse,    # data management + ggplot2 graphics, 
  gt,           # Nice beuatiful tables
  nortest,      # Anderson-Darling test for normality
  ggplot2,      # For creating plots
  DescTools,    # For statistical functions like skewness and kurtosis
  janitor,      # adding totals and percents to tables
  car,           # For Levene's test and ANOVA
  gtsummary,    # summary statistics and tests
  corrr,        # correlation analayis for numeric variables
  lmtest,
  dplyr,        # For data manipulation
  reshape2,     # For reshaping data
  skimr,        # get overview of data
  performance  # For model performance metrics
            )
#library(LogisticDx)    # requires complex install             
```

## Access the data
```{r read-data, echo = TRUE}
# file <- "./EpiData/stroke.dta"     #  use this if the file is on your computer
file <- "https://raw.githubusercontent.com/drkamarul/multivar_data_analysis/refs/heads/main/data/stroke.dta"
df3 <- import(file, trust  = TRUE)
summary(df3)
glimpse(df3)  # get an overview of the data
```

## Fix the data
```{r fix-data, echo = TRUE}
# Need to change "status" to be 1,0 
df3 <- df3 %>% 
  mutate(status = ifelse(status == 2, 0, 1))

df3 %>%
  tbl_summary(by = status) %>%
  as_gt()
```

## Logistic regression
```{r logistic-regression, echo = TRUE}
# Perform logistic regression
log_reg <- glm(status ~ gcs, data = df3, family = binomial(link = 'logit'))
tidy(mlog_reg)
```

## Model performance measures
```{r model-performance, echo = TRUE}
# Get -2LL and AIC
minus_2LL_log_reg <- -2 * as.numeric(logLik(log_reg))
AIC(log_reg)

# Calculate Nagelkerke's R虏
r2_val <- as.numeric(r2_nagelkerke(log_reg))

# Disiplay the summary of the logistic regression model
summary(log_reg)
r2_val
```

# Calculate log OR
```{r calc-log_OR, echo = TRUE}
log_reg_tbl <- tbl_regression(log_reg, exponentiate = FALSE)
log_reg_tbl
```

# Calculate OR
```{r calc-OR, include=FALSE}
log_reg_OR <- tbl_regression(log_reg, exponentiate = TRUE)
log_reg_OR
```

# Combine logOR and OR results with model performance 
```{r log_reg_tbl, echo = TRUE}
combined_tbl <- tbl_merge(
  tbls = list(log_reg_tbl, log_reg_OR),
  tab_spanner = c("**log Odds**", "**Odds Ratio**")  # Optional: Add spanning headers
) %>%
  as_gt() %>%
  tab_source_note(

    source_note = paste0("R虏 = ", round(r2_val,3),
                          ", -2LL = ", round(minus_2LL_log_reg, 0), 
                          ", AIC = ", round(AIC(log_reg), 0)
                        )
  )

# Display the combined table
combined_tbl
```

## Logistic regression plot
```{r log_reg-plt, echo = TRUE}
# use ggplot to show the scatter plot with regression line with linear model and confidence interval
y_var <- "status"
x_var <- "gcs"
log_reg_plot <- ggplot(df3, aes_string(x = x_var, y = y_var)) +
  geom_point() +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = TRUE, color = "blue") +  # Adding the trend line with linear model
  labs(title = paste("Scatter Plot of", y_var, "vs", x_var),
       x = x_var,
       y = y_var) +
  theme_minimal()
log_reg_plot
```





## 2nd Logistic regression
```{r logistic_regression,2 echo = TRUE}
# Perform logistic regression
log_reg2 <- glm(status ~ age, data = df3, family = binomial(link = 'logit'))
tidy(mlog_reg2)
```

## 2nd Logistic regression performance measures
```{r model-performance2, echo = TRUE}
# Calculate Nagelkerke's R虏
r2_val2 <- as.numeric(r2_nagelkerke(log_reg2))

# Get -2LL and AIC
minus_2LL_log_reg2 <- -2 * as.numeric(logLik(log_reg2))
AIC(log_reg2)

# Disiplay the summary of the logistic regression model
summary(log_reg2)
r2_val2
```

## Calculate log OR for 2nd Logistic regression
```{r calc-log_OR2, echo = TRUE}
log_reg_tbl2 <- tbl_regression(log_reg2, exponentiate = FALSE)
log_reg_tbl2
```

## Calculate OR for 2nd Logistic regression
```{r calc-OR2, echo = TRUE}
log_reg_OR2 <- tbl_regression(log_reg2, exponentiate = TRUE)
log_reg_OR2
```

## Combine 2nd Logistic regression logOR and OR results with model performance 
```{r log_reg_tbl2, echo = TRUE}
combined_tbl <- tbl_merge(
  tbls = list(log_reg_tbl2, log_reg_OR2),
  tab_spanner = c("**log Odds**", "**Odds Ratio**")  # Optional: Add spanning headers
) %>%
  as_gt() %>%
  tab_source_note(

    source_note = paste0("R虏 = ", round(r2_val2,3),
                          ", -2LL = ", round(minus_2LL_log_reg2, 0), 
                          ", AIC = ", round(AIC(log_reg2), 0)
                        )
  )

# Display the combined table
combined_tbl2
```

## 2st Logistic regression plot
```{r log_reg-plot2, echo = TRUE}
# use ggplot to show the scatter plot with regression line with linear model and confidence interval
y_var2 <- "status"
x_var2 <- "age"
log_reg2_plot <- ggplot(df3, aes_string(x = x_var2, y = y_var2)) +
  geom_point() +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = TRUE, color = "blue") +  # Adding the moddel line
  labs(title = paste("Scatter Plot of", y_var2, "vs", x_var2),
       x = x_var2,
       y = y_var2) +
  theme_minimal()
log_reg2_plot
```


### --------------------------------------


## Compare Odds Ratio between models
```{r compare, echo = TRUE}
stacked_tbl <- tbl_stack(
  tbls = list(log_reg_OR, log_reg_OR2),
  quiet = TRUE  # Suppress header difference warning
) %>%
  as_gt() %>%
  tab_source_note(
    source_note = html("<b>1st model: </b>",
                       "R虏 = ", round(r2_val,2),
                       ", -2LL = ", round(minus_2LL_log_reg, 0), 
                       ", AIC = ", round(AIC(log_reg), 0),
                       "<br>",
                       "<b>2nd model: </b>",
                       "R虏 = ", round(r2_val2,3),
                       ", -2LL = ", round(minus_2LL_log_reg2, 0), 
                       ", AIC = ", round(AIC(log_reg2), 0) 
                      )
  )

stacked_tbl
```