###
### name:  专
### ID  : [住驻专 转注转 转]
### course: 驻 (HIT.ac.il)
### lesson: 专专住 专转 06   
### date  : 13/07/2025
###  
###   注砖转 专专住 住 专 注 砖转 专


## Reset memory
```{r setup-packages, include = FALSE}
rm(list = ls(all.names = TRUE))       #  Remove all data objects         Remove packages
lapply(paste("package:", names(sessionInfo()$otherPkgs), sep = ""), detach, character.only = TRUE, unload = TRUE, force = TRUE)           # Clear packages from memory
gc()                                  # Remove unused data from memory ("garbage cllection")
dev.off()                             # Reset the display output of R
q("no")                               # This will completely shutdown R terminal
```

## Load required packages
```{r load-packages, include = FALSE}  
options("install.lock" = FALSE)
if (!require("pacman")) install.packages("pacman") ; library(pacman)
pacman::p_load(
  Rcpp, devtools, gh, remotes,   # For installing packages/files online (like from GitHub)
  rio, here,              # File import and path support
  labelled, DT,           # deal with labels
  tidyverse,              # data management + ggplot2 graphics,
  broom,                  # For tidying model outputs
  dplyr,                  # For data manipulation
  gt,                     # Nice beuatiful tables
  gtsummary,              # summary statistics and tests
  nortest,                # Anderson-Darling test for normality
  ggplot2,                # For creating plots
  ggpubr,                 # For ggplot2 extensions
  DescTools,              # For statistical functions like skewness and kurtosis
  summarytools,           # to get a table of all summary statistics
  car,                    # For Levene's test and ANOVA
  lmtest,                 # do linear regression tests
  parameters,             # get standardized parameters/coefficients
  GGally,                 # to make a correlation matrix
  skimr,                  # get overview of data
  pROC,                   # For ROC analysis
  cvms,                   # For confusion matrix and classification metrics
  ggeffects,              # To display the confusion matrix
  performance,            # For model performance metrics
  generalhoslem,          # For Hosmer-Lemeshow test for goodness of fit 
  # MASS,                   # For statistical functions (polr for ordinal regression)ordinal
  ordinal                 # For ordinal regression statistical functions 
              )
# Install/load flexplot
if (!require("devtools")) devtools::install_github("dustinfife/flexplot", ref = "development")  ;  if (!require("pacman")) install.packages("pacman")  ;  library(flexplot)
```

## Access the data
```{r read-data, echo = TRUE}
## read the data in a file online
# file <- "./EpiData/05 peptic_ulcer.rds"     # use this if the file is on your computer
file <- "https://raw.githubusercontent.com/Model-Lab-Net/Courses/refs/heads/main/Epi/Data/05%20peptic_ulcer.rds"
df5 <- import(file, trust = TRUE)
# show variable translations to Hebrew in lables
labels <- data.frame(Variable = names(df5), Label = unlist(var_label(df5)))   ;   datatable(labels, options = list(pageLength = 100, autoWidth = TRUE, dom = 't'), caption = "Variable Labels for df5", rownames = FALSE)
glimpse(df5)               # get an overview of the data
```

## Perfrom an ordinal linear regression
```{r ord-logisitc-reg, echo = TRUE}
# Ordinal regression model
ord_lr <- clm(perforation.ord ~ pulp + hemoglobin + systolic, data = df5, link = "logit")
summary(ord_lr)

# Variance-covariance matraix
vcov_matrix <- vcov(ord_lr)
vcov_matrix
```

## Model performance
```{r ord-logisitc-reg, echo = TRUE}
# Get -2LL and AIC
minus_2LL_val <- -2 * as.numeric(logLik(ord_lr))
minus_2LL_val <- round(minus_2LL_val,0)

AIC_val <- AIC(ord_lr)
AIC_val <- round(AIC_val, 0)

# Get Efron's R虏
Nagelkerke <- r2_nagelkerke(ord_lr)
r2_val <- as.numeric(Nagelkerke)
r2_val <- round(r2_val, 3)

# Disiplay the summary of the logistic regression model
paste("R2= ", r2_val)
```

# Display regression results
```{r ord-logisitc-reg, echo = TRUE}
ord_lr_summary <- tbl_regression(
  ord_lr,
  exponentiate = TRUE # Gives odds ratios instead of log-odds
) %>%
  as_gt() %>%
  tab_source_note(
    source_note = paste0(
      "R虏 = ", r2_val,
      ", -2LL = ", minus_2LL_val,
      ", AIC = ", AIC_val
    )
  )
ord_lr_summary
```

## Classification table
```{r predict, echo = TRUE}
# Calculate predicted probabilities and append to df
df5$predict.prob <- unlist(predict(ord_lr, type = "class"))

# Calculate the "confusion_matrix"
```{r matrix, echo=FALSE}
cm <- confusion_matrix(targets = df5$perforation.ord, predictions = df5$predict.prob)

# Plot the "confusion_matrix"
plot_confusion_matrix(cm)

# Display the "classification metrics"
cm_metrics <- select_metrics(cm)
cm_metrics <- cm_metrics %>%
  pivot_longer(cols = everything(),
               names_to = "Metric",
               values_to = "Value"
              )
cm_metrics <- as.data.frame(cm_metrics)
cm_metrics <- cm_metrics %>%
  mutate(across(where(is.numeric), ~ round(., 3)))
cm_metrics

cm_metrics %>%
  gt() 
```

## Nominal effects test
```{r , echo=FALSE}
# Test to make sure model assumption are not violated
nominal_test(ord_lr)
          # Note: If p<0.05 for all variables then fit is good.
          ### ------------------------------------------------
          ### Important: HosmerLemeshow only works for binary 
          ### this is more genralizable logistic regression
```

## Create distribution plots
```{r predict, echo = TRUE}
# Distribution of degree_perforation
flexplot(perforation.ord ~  pulp | hemoglobin + systolic, data = df5, alpha = 0.6, method = "lm")

flexplot(perforation.ord ~ 1, data = df5, alpha = 0.6)

# Distribution of hemoglobin vs degree_perforation
flexplot(hemoglobin ~ perforation.ord, data = df5, alpha = 0.6)

# Distribution of systolic vs degree_perforation
flexplot(systolic ~ perforation.ord, data = df5, alpha = 0.6)

# # Distribution of PULP vs degree_perforation (assuming numeric)
p <- flexplot(pulp ~ perforation.ord, data = df5, alpha = 0.6)
p <- p + ggtitle("PULP Score by Degree of Perforation")
p
```

### Create prediction plots
```{r predict, echo = TRUE}
# Predicted probabilities for pulp
preds_pulp <- ggpredict(ord_lr, terms = "pulp")
# Build the plot with ggplot2 directly
ggplot(preds_pulp, aes(x = x, y = predicted, color = response.level, group = response.level)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = response.level, group = response.level),
              alpha = 0.2,
              color = NA) +
  labs(
    title = "Predicted Probability of Degree of Perforation by pulp",
    x = "pulp",
    y = "Predicted Probability",
    color = "Outcome",
    fill = "Outcome"
  ) +
  theme_minimal()


# Predicted probabilities for hemoglobin
preds_hemoglobin <- ggpredict(ord_lr, terms = "hemoglobin")
# Build the plot with ggplot2 directly
ggplot(preds_hemoglobin, aes(x = x, y = predicted, color = response.level, group = response.level)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = response.level, group = response.level),
              alpha = 0.2,
              color = NA) +
  labs(
    title = "Predicted Probability of Degree of Perforation by pulp",
    x = "pulp",
    y = "Predicted Probability",
    color = "Outcome",
    fill = "Outcome"
  ) +
  theme_minimal()


# Predicted probabilities for systolic
preds_systolic <- ggpredict(ord_lr, terms = "systolic")
# Build the plot with ggplot2 directly
ggplot(preds_systolic, aes(x = x, y = predicted, color = response.level, group = response.level)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = response.level, group = response.level),
              alpha = 0.2,
              color = NA) +
  labs(
    title = "Predicted Probability of Degree of Perforation by pulp",
    x = "pulp",
    y = "Predicted Probability",
    color = "Outcome",
    fill = "Outcome"
  ) +
  theme_minimal()
```





### -------------------------------------------------
## 转专-注爪

## 砖 1
```{r question-1, echo = TRUE}
# 

```

## 砖 2
```{r question-2, echo = TRUE}
# 

```

## 砖 3
```{r question-3, echo = TRUE}
# 

```









### source: https://www.scielosp.org/pdf/rsp/2009.v43n1/183-194/en

## Clean the data
```{r clean-data, echo = TRUE}
# Load the data
file <- "https://raw.githubusercontent.com/drkamarul/multivar_data_analysis/refs/heads/main/data/peptic_ulcer.xlsx"
df5 <- import(file, trust = TRUE)
#write.csv(df5, here("EpiData", "05 peptic_ulcer.csv"))

# Clean up variable names
df5 <- clean_names(df5)
names(df5) <- gsub("_", ".", names(df5))

# Swap all char variables to fct
df5 <- mutate_if(df5, is.character, as.factor)

# Create an ordinal variable from perforation
df5$perforation.ord <- factor(
  ifelse(df5$perforation <= 1.0, 0L, 1L),  # use integer values
  levels = c(0L, 1L),                     # explicitly set levels
  ordered = TRUE                          # make it ordinal
)

# Check the updated dataset structure
glimpse(df5)

# Save as R data file
saveRDS(df5, here("EpiData", "05 peptic_ulcer.rds"))
```

## Give variable explanations in labels
```{r labels, echo = TRUE}
attr(df5$age, "label") <- ""
attr(df5$gender, "label") <- "专"
attr(df5$epigastric.pain, "label") <- " 注"
attr(df5$vomiting, "label") <- "拽转"
attr(df5$nausea , "label") <- "转"
attr(df5$fever, "label") <- ""
attr(df5$diarrhea, "label") <- "砖砖"
attr(df5$malena, "label") <- " 注"
attr(df5$onset.more.24.hrs, "label") <- " 砖 转专 -24 砖注转"
attr(df5$nsaids, "label") <- "转专驻转 -拽转转"
attr(df5$septic.shock, "label") <- " 住驻"
attr(df5$previous.ogds, "label") <- "拽转 注"
attr(df5$asa, "label") <- "转 住驻专"
attr(df5$systolic, "label") <- "抓  住住"
attr(df5$diastolic, "label") <- "抓  住"
attr(df5$inotropes, "label") <- "转专驻转 转 "
attr(df5$pulse, "label") <- "驻拽"
attr(df5$tenderness, "label") <- "专砖转 专 注"
attr(df5$guarding, "label") <- "抓 砖专专 "
attr(df5$hemoglobin, "label") <- " "
attr(df5$twc, "label") <- " 专转  砖 Truelove-Witts"
attr(df5$platelet, "label") <- "转 住转 "
attr(df5$creatinine, "label") <- " 驻注转 转"
attr(df5$albumin, "label") <- " 驻注转 "
attr(df5$pulp, "label") <- "拽  驻驻"
attr(df5$admission.to.op.hrs , "label") <- "住驻专 砖注转  砖驻  转"
attr(df5$perforation, "label") <- " 拽 注"
attr(df5$degree.perforation, "label") <- "专转 拽 注"
attr(df5$side.perforation, "label") <- "拽 爪"
attr(df5$icu, "label") <- "砖驻 驻 专抓"
attr(df5$sssi, "label") <- " 拽"
attr(df5$anast.leak, "label") <- " 爪专  拽 专"
attr(df5$sepsis, "label") <- " "
attr(df5$outcome, "label") <- " / 驻专"
attr(df5$perforation.ord, "label") <- "专转 拽 注 (砖转 专)"


labels <- NULL
labels <- data.frame(Variable = names(df5), Label = unlist(var_label(df5)))
datatable(labels, options = list(pageLength = 100, autoWidth = TRUE, dom = 't'), caption = "Variable Labels for df5", rownames = FALSE)
```