###
### name: דוד בורג
### ID  : [מספר תעודת זהות]
### course: אפידמיולוגיה (HIT.ac.il)
### lesson: רגרסיה אורדינלית 06   
### date  : 03/07/2025
###  
###  נלמד לעשות רגרסיה לוגיסטי מרובה על משתנים אורדינליים
### 
### source: https://www.scielosp.org/pdf/rsp/2009.v43n1/183-194/en


## Reset memory
```{r setup-packages, include = FALSE}
rm(list = ls(all.names = TRUE))
lapply(paste("package:", names(sessionInfo()$otherPkgs), sep = ""), detach, character.only = TRUE, unload = TRUE, force = TRUE)
gc()
dev.off()
```

## Load required packages
```{r load-packages, include = FALSE}  
options("install.lock" = FALSE)
if (!require("pacman")) install.packages("pacman")
library(pacman)
pacman::p_load(
  rio,                    # File import
  tidyverse,              # data management + ggplot2 graphics,
  broom,                  # For tidying model outputs
  dplyr,                  # For data manipulation
  gt,                     # Nice beuatiful tables
  nortest,                # Anderson-Darling test for normality
  ggplot2,                # For creating plots
  DescTools,              # For statistical functions like skewness and kurtosis
  psych,                  # to get a table of all summary statistics
  janitor,                # adding totals and percents to tables
  car,                    # For Levene's test and ANOVA
  gtsummary,              # summary statistics and tests
  corrr,                  # correlation
  lmtest,                 # linear regression tets
  parameters,             # calculate standardized parameters/coefficients
  GGally,                 # to make a correlation matrix
  skimr,                  # get overview of data
  pROC,                   # For ROC analysis
  performance,            # For model performance metrics
  caret,                  # For confusion matrix and classification metrics
  ggeffects,              # To display the confusion matrix
  ResourceSelection,       # For Hosmer-Lemeshow test for goodness of fit 
  MASS,                   # For statistical functions (polr for ordinal regression)ordinal
  ordinal                 # For ordinal regression statistical functions 
              )
# Install/load flexplot
if (!require("devtools")) devtools::install_github("dustinfife/flexplot", ref = "development")  ;  if (!require("pacman")) install.packages("pacman")  ;  library(flexplot)
```

## Access the data
```{r read-data, echo = TRUE}
## read the data in a file online
# file <- "./EpiData/peptic_ulcer.xlsx"     # use this if the file is on your computer
file <- "https://raw.githubusercontent.com/drkamarul/multivar_data_analysis/refs/heads/main/data/peptic_ulcer.xlsx"
df5 <- import(file, trust = TRUE)
summary(df5)
glimpse(df5)
```

## Fix the data
```{r fix-data, echo = TRUE}
# clean up variable names
df5 <- clean_names(df5)
names(df5) <- gsub("_", ".", names(df5))

# Swap all char variables to fct
df5 <- mutate_if(df5, is.character, as.factor)

# Make sure degree_perforation is an ordered factor
df5 <- df5 %>%
  mutate(degree.perforation = factor(degree.perforation,
    levels = c("small", "moderate", "large"),
    ordered = TRUE
  ))
```

## Perfrom an ordinal linear regression
```{r ord-logisitc-reg, echo = TRUE}
# Ordinal regression model
# ord_lr <- polr(degree.perforation ~ puplp + hemoglobin + systolic , data = df5, Hess = TRUE)
ord_lr <- clm(degree.perforation ~ pulp + hemoglobin + systolic, data = df5, link = "logit")
summary(ord_lr)

# Variance-covariance matraix
vcov_matrix <- vcov(ord_lr)
```

## Model performance
```{r ord-logisitc-reg, echo = TRUE}
ord_lr_perf <- polr(degree.perforation ~ pulp + hemoglobin + systolic, data = df5, Hess = TRUE)
# Get -2LL and AIC
minus_2LL_val <- -2 * as.numeric(logLik(ord_lr_perf))
AIC_val <- AIC(ord_lr_perf)

# Get Efron's R²
Nagelkerke <- PseudoR2(ord_lr_perf, which = "Nagelkerke")
r2_val <- as.numeric(Nagelkerke)

# Disiplay the summary of the logistic regression model
paste("R2= ", r2_val)
```

# Display regression results
```{r ord-logisitc-reg, echo = TRUE}
ord_lr_summary <- tbl_regression(
  ord_lr,
  exponentiate = TRUE # Gives odds ratios instead of log-odds
) %>%
  as_gt() %>%
  tab_source_note(
    source_note = paste0(
      "R² = ", round(r2_val, 3),
      ", -2LL = ", round(minus_2LL_val, 0),
      ", AIC = ", round(AIC_val, 0)
    )
  )
ord_lr_summary
```

## Classification table
```{r predict, echo = TRUE}
# Create predicted values on df
df5$predict_prob <- predict(ord_lr_perf, type = "class")

## Classification table
```{r , echo=FALSE}
# Use caret package to get classification table
confusion_matrix <- caret::confusionMatrix(data = df5$predict_prob, reference = df5$degree.perforation)
confusion_matrix

# Extract the confusion matrix table
cm_table <- as.data.frame(confusion_matrix$table)
colnames(cm_table) <- c("Predicted", "True", "Count")
cm_table <- cm_table %>%
  mutate(Percent = round(Count / sum(Count) * 100, 1))

# Plot confusion matrix as a heatmap
ggplot(data = cm_table, aes(x = Predicted, y = True, fill = Percent)) +
  geom_tile() +
  geom_text(aes(label = Percent), vjust = 0.5, color = "black") +
  scale_fill_gradient(low = "white", high = "blue") +
  theme_minimal() +
  labs(title = "Confusion Matrix", x = "Predicted", y = "True") +
  theme(axis.text.x = element_text(angle = 0, hjust = 1))

# Get cm metrics
cm_metrics_df <- data.frame(
  Metric = c("Accuracy", "PPV", "NPV", "Sensitivity", "Specificity", "Precision", "F1 Score"),
  Value = round(c(
    confusion_matrix$overall["Accuracy"],
    confusion_matrix$byClass["Pos Pred Value"],
    confusion_matrix$byClass["Neg Pred Value"],
    confusion_matrix$byClass["Sensitivity"],
    confusion_matrix$byClass["Specificity"],
    confusion_matrix$byClass["Precision"],
    confusion_matrix$byClass["F1"]
  ), 2)
)
cm_metrics_df

cm_metrics_df %>%
  gt()
```

## Hosmer-Lemeshow test for GOF
```{r , echo=FALSE}
# Use ResourceSelection to get the Hosmer–Lemeshow test result
hl_test <- hoslem.test(df5$degree.perforation, df5$predicted_prob, g = 10)
hl_test
```

## Create distribution plots
```{r predict, echo = TRUE}}
# Distribution of degree_perforation
flexplot(degree.perforation ~  pulp | hemoglobin + systolic, data = df5, alpha = 0.6, method = "lm")

flexplot(degree.perforation ~ 1, data = df5, alpha = 0.6)

# Distribution of hemoglobin vs degree_perforation
flexplot(hemoglobin ~ degree.perforation, data = df5, alpha = 0.6)

# Distribution of systolic vs degree_perforation
flexplot(systolic ~ degree.perforation, data = df5, alpha = 0.6)

# # Distribution of PULP vs degree_perforation (assuming numeric)
p <- flexplot(pulp ~ degree.perforation, data = df5, alpha = 0.6)
p <- p + ggtitle("PULP Score by Degree of Perforation")
p
```

### Create prediction plots
```{r predict, echo = TRUE}
# Predicted probabilities for pulp
preds_pulp <- ggpredict(ord_lr_perf, terms = "pulp")
# Build the plot with ggplot2 directly
ggplot(preds_pulp, aes(x = x, y = predicted, color = response.level)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = response.level), alpha = 0.2, color = NA) +
  labs(
    title = "Predicted Probability of Degree of Perforation by pulp",
    x = "pulp",
    y = "Predicted Probability",
    color = "Outcome",
    fill = "Outcome"
  ) +
  theme_minimal()

# Predicted probabilities for hemoglobin
preds_hemoglobin <- ggpredict(ord_lr_perf, terms = "hemoglobin")
# Build the plot with ggplot2 directly
ggplot(preds_hemoglobin, aes(x = x, y = predicted, color = response.level)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = response.level), alpha = 0.2, color = NA) +
  labs(
    title = "Predicted Probability of Degree of Perforation by hemoglobin",
    x = "pulp",
    y = "Predicted Probability",
    color = "Outcome",
    fill = "Outcome"
  ) +
  theme_minimal()


# Predicted probabilities for systolic
preds_systolic <- ggpredict(ord_lr_perf, terms = "systolic")
# Build the plot with ggplot2 directly
ggplot(preds_systolic, aes(x = x, y = predicted, color = response.level)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = response.level), alpha = 0.2, color = NA) +
  labs(
    title = "Predicted Probability of Degree of Perforation by systolic bp",
    x = "pulp",
    y = "Predicted Probability",
    color = "Outcome",
    fill = "Outcome"
  ) +
  theme_minimal()
```
