###
### name  : __________________
### ID  : [מספר תעודת זהות]
### course: אפידמיולוגיה (HIT.ac.il)
### lesson: רגרסיה אורדינלית 06   
### date  : 19/04/2026
###  
###  נלמד לעשות רגרסיה לוגיסטי מרובה על משתנים אורדינליים
###
### 


## Reset memory
```{r setup-packages, include = FALSE}
rm(list = ls(all.names = TRUE))          # 👈 Clear all data objects        👇 Clear packages
lapply(paste("package:", names(sessionInfo()$otherPkgs), sep = ""), detach, character.only = TRUE, unload = TRUE, force = TRUE)
gc()                                     # Clear unused data from memory ("garbage cllection")
dev.off()                                # Reset the display output of R
system("powershell -Command Clear-Host") # This will clear the terminal in VScode
```

## Load required packages
```{r load-packages, include = FALSE}  
options("install.lock" = FALSE)
if (!require("pacman")) install.packages("pacman") ; library(pacman)
pacman::p_load(
  Rcpp, devtools, gh, remotes,   # For installing packages/files online (like from GitHub)
  rio, here,              # File import and path support
  labelled, DT,           # deal with labels
  tidyverse,              # data management + ggplot2 graphics,
  broom,                  # For tidying model outputs
  dplyr,                  # For data manipulation
  gt,                     # Nice beuatiful tables
  gtsummary,              # summary statistics and tests
  grid, gridExtra,
  nortest,                # Anderson-Darling test for normality
  ggplot2,                # For creating plots
  ggpubr,                 # For ggplot2 extensions
  DescTools,              # For statistical functions like skewness and kurtosis
  summarytools,           # to get a table of all summary statistics
  car,                    # For Levene's test and ANOVA
  lmtest,                 # do linear regression tests
  parameters,             # get standardized parameters/coefficients
  GGally,                 # to make a correlation matrix
  skimr,                  # get overview of data
  pROC,                   # For ROC analysis
  cvms,                   # For confusion matrix and classification metrics
  ggeffects,              # To display the confusion matrix
  performance,            # For model performance metrics
  generalhoslem,          # For Hosmer-Lemeshow test for goodness of fit 
  rms,
  # MASS,                   # For statistical functions (polr for ordinal regression)ordinal
  ordinal                 # For ordinal regression statistical functions 
              )
# Install/load flexplot
if (!require("devtools")) devtools::install_github("dustinfife/flexplot", ref = "development")  ;  if (!require("pacman")) install.packages("pacman")  ;  library(flexplot)
```

## Access the data
```{r read-data, echo = TRUE}
## read the data in a file online
# file <- "./EpiData/05 peptic_ulcer.rds"     # use this if the file is on your computer
file <- "https://raw.githubusercontent.com/Model-Lab-Net/Courses/refs/heads/main/Epi/Data/05%20peptic_ulcer.rds"
df5 <- import(file, trust = TRUE)
# show variable translations to Hebrew in lables
labels <- data.frame(Variable = names(df5), Label = unlist(var_label(df5)))   ;   datatable(labels, options = list(pageLength = 100, autoWidth = TRUE, dom = 't'), caption = "Variable Labels for df5", rownames = FALSE)
glimpse(df5)               # get an overview of the data
```

## Perfrom an ordinal linear regression
```{r ord-logisitc-reg, echo = TRUE}
# Ordinal regression model
ord_model <- clm(perforation.ord ~ pulp + hemoglobin + systolic, data = df5, link = "logit")
summary(ord_model) 

ordinal_model <- tbl_regression(ord_model, exponentiate = TRUE) 
ordinal_model
```

## Vcov matrix
```{r ord-logisitc-reg, echo = TRUE}
# Variance-covariance matrix
vcov_matrix <- vcov(ord_model)
vcov_matrix

# Send the vcov matrix to a nice table
vcov_matrix %>%
  as.data.frame() %>%
  rownames_to_column(var = "Variable") %>%
  gt() %>%
  fmt_number(
    columns = -Variable, # Format all columns except the 'Variable' names
    decimals = 3
  )
```

## Model diagnostics
```{r ord-logisitc-reg, echo = TRUE}
dd <- datadist(df5)   ;   options(datadist = "dd")
ord_model_ <- orm(perforation.ord ~ pulp + hemoglobin + systolic, data = df5)     # need this for model diagnostics. sorry.
# NEW way
AIC_val  <- as.numeric(AIC(ord_model))                       # AIC
r2_val   <- as.numeric(r2_nagelkerke(ord_model))
chi_sq   <- as.numeric(ord_model_$stats["Model L.R."])       # Omnibus Chi-square (Likelihood Ratio Chi²)
p_val    <- as.numeric(ord_model_$stats["P"]  )  

# Add model diagnostics to table
ordinal_model %>%
  as_gt() %>% # Convert to gt engine first
  tab_source_note(
    source_note = md(paste0(
      "R² = ", round(r2_val, 3), "<br>",
      ",  AIC = ", round(AIC_val, 0), "<br>",
      ",  χ² = ", round(chi_sq, 0), "<br>",
      ",  p = ", round(p_val, 5)
    ))
  )
```

## Classification table
```{r predict, echo = TRUE}
# Calculate predicted probabilities and append to df
df5$predict.prob <- predict(ord_model_, type = "fitted.ind")
df5$predict.class <- as.integer(df5$predict.prob > 0.5)
glimpse(df5)

# Calculate the "confusion_matrix" using cvms
cfm <- confusion_matrix(targets = df5$perforation.ord, predictions = df5$predict.class)

# Plot the "confusion_matrix" using cvms
plot_confusion_matrix(cfm)

# Display the "classification metrics" using cvms
# Get the "classification metrics" using cvms
cfm_metrics <- select_metrics(cfm) %>%
  # convert all columns to numeric (if needed)
  #mutate(across(everything(), as.numeric)) %>%
  # pivot to long format directly
  pivot_longer(
    cols = everything()
    #names_to = "Metric",
    #values_to = "Value"
  )


cfm_metrics %>%
  gt() %>%
  fmt_number(
    decimals = 3
  )
```

## Nominal effects test
```{r , echo=FALSE}
# Test to make sure model assumptions are not violated
nomial <- nominal_test(ord_model)
          # Note: If p>0.05 for all variables then fit is good.
          ### ------------------------------------------------
          ### Important: Hosmer–Lemeshow only works for binary 
          ### this is a more genralizable logistic regression

nomial %>% 
  gt() %>%
    fmt_number(
      decimals = 1
    )       
```

### Create prediction plots
```{r predict, echo = TRUE}
# Predicted probabilities for pulp
preds_pulp <- ggpredict(ord_model, terms = "pulp")
plot(preds_pulp)

# Predicted probabilities for hemoglobin
preds_hemoglobin <- ggpredict(ord_model, terms = "hemoglobin")
plot(preds_hemoglobin)

# Predicted probabilities for systolic
preds_systolic <- ggpredict(ord_model, terms = "systolic")
plot(preds_systolic)
```

## Create distribution plots
```{r predict, echo = TRUE}
# Distribution of degree_perforation
flexplot(perforation.ord ~  pulp | hemoglobin + systolic, data = df5, alpha = 0.6, method = "lm")

flexplot(perforation.ord ~ 1, data = df5, alpha = 0.6)

# Distribution of hemoglobin vs degree_perforation
flexplot(hemoglobin ~ perforation.ord, data = df5, alpha = 0.6)

# Distribution of systolic vs degree_perforation
flexplot(systolic ~ perforation.ord, data = df5, alpha = 0.6)

# # Distribution of PULP vs degree_perforation (assuming numeric)
p <- flexplot(pulp ~ perforation.ord, data = df5, alpha = 0.6)
p <- p + ggtitle("PULP Score by Degree of Perforation")
p
```





### -------------------------------------------------
## תרגול-עצמי

## שאלה 1
```{r question-1, echo = TRUE}
# 

```

## שאלה 2
```{r question-2, echo = TRUE}
# 

```

## שאלה 3
```{r question-3, echo = TRUE}
# 

```









### source: https://www.scielosp.org/pdf/rsp/2009.v43n1/183-194/en

## Clean the data
```{r clean-data, echo = TRUE}
# Load the data
file <- "https://raw.githubusercontent.com/drkamarul/multivar_data_analysis/refs/heads/main/data/peptic_ulcer.xlsx"
df5 <- import(file, trust = TRUE)
#write.csv(df5, here("EpiData", "05 peptic_ulcer.csv"))

# Clean up variable names
df5 <- clean_names(df5)
names(df5) <- gsub("_", ".", names(df5))

# Swap all char variables to fct
df5 <- mutate_if(df5, is.character, as.factor)

# Create an ordinal variable from perforation
df5$perforation.ord <- factor(
  ifelse(df5$perforation <= 1.0, 0L, 1L),  # use integer values
  levels = c(0L, 1L),                     # explicitly set levels
  ordered = TRUE                          # make it ordinal
)

# Check the updated dataset structure
glimpse(df5)

# Save as R data file
saveRDS(df5, here("EpiData", "05 peptic_ulcer.rds"))
```

## Give variable explanations in labels
```{r labels, echo = TRUE}
attr(df5$age, "label") <- "גיל"
attr(df5$gender, "label") <- "מגדר"
attr(df5$epigastric.pain, "label") <- "כאבי מעיים"
attr(df5$vomiting, "label") <- "הקאות"
attr(df5$nausea , "label") <- "בחילות"
attr(df5$fever, "label") <- "חום"
attr(df5$diarrhea, "label") <- "שלשול"
attr(df5$malena, "label") <- "דימום במעיים"
attr(df5$onset.more.24.hrs, "label") <- "כיב הנמשך יותר מ-24 שעות"
attr(df5$nsaids, "label") <- "תרופות אנטי-דלקתיות"
attr(df5$septic.shock, "label") <- "הלם ספטי"
attr(df5$previous.ogds, "label") <- "בדיקות מעיים"
attr(df5$asa, "label") <- "נטילת אספירין"
attr(df5$systolic, "label") <- "לחץ דם סיסטולי"
attr(df5$diastolic, "label") <- "לחץ דם דיאסטולי"
attr(df5$inotropes, "label") <- "תרופות למחלות לב"
attr(df5$pulse, "label") <- "דופק"
attr(df5$tenderness, "label") <- "רגישות באיזור המעיים"
attr(df5$guarding, "label") <- "כיווץ שרירי בטן"
attr(df5$hemoglobin, "label") <- "המוגלובין בדם"
attr(df5$twc, "label") <- "מדד חומרת הכיב של Truelove-Witts"
attr(df5$platelet, "label") <- "כמות טסיות דם"
attr(df5$creatinine, "label") <- "מדד לפעיולת הכליות"
attr(df5$albumin, "label") <- "מדד לפעיולת הכבד"
attr(df5$pulp, "label") <- "ניקוב כיב פפטי"
attr(df5$admission.to.op.hrs , "label") <- "מספר שעות בין אישפוז ובין ניתוח"
attr(df5$perforation, "label") <- "גודל ניקוב במעיים"
attr(df5$degree.perforation, "label") <- "חומרת ניקוב במעיים"
attr(df5$side.perforation, "label") <- "ניקוב צדדי"
attr(df5$icu, "label") <- "אישפוז בטיפול נמרץ"
attr(df5$sssi, "label") <- "כיב מודלק"
attr(df5$anast.leak, "label") <- "כיב הנוצר ליד מקום כירוגי"
attr(df5$sepsis, "label") <- "אלח דם"
attr(df5$outcome, "label") <- "חי / נפטר"
attr(df5$perforation.ord, "label") <- "חומרת ניקוב במעיים (במשתנה אורדינלי)"


labels <- NULL
labels <- data.frame(Variable = names(df5), Label = unlist(var_label(df5)))
datatable(labels, options = list(pageLength = 100, autoWidth = TRUE, dom = 't'), caption = "Variable Labels for df5", rownames = FALSE)
```