###
### name:  专
### ID  : [住驻专 转注转 转]
### course: 驻 (HIT.ac.il)
### lesson: 专专住 住转 06   
### date  : 15/06/2025
###  
###   注砖转 专住 住转 专 砖转
###
### source: https://bookdown.org/drki_musa/dataanalysis/binary-logistic-regression.html
###         https://github.com/drkamarul/multivar_data_analysis/tree/main/data

## Reset memory
```{r setup-packages, include = FALSE}
rm(list = ls(all.names = TRUE))
lapply(paste('package:',names(sessionInfo()$otherPkgs),sep=""),detach,character.only=TRUE,unload=TRUE,force=TRUE)
gc()
dev.off()
```

## Load required packages
```{r load-packages, include = FALSE}  
options("install.lock"=FALSE)
if (!require("pacman")) install.packages("pacman")
library(pacman)
pacman::p_load(
  rio,          # File import
  tidyverse,    # data management + ggplot2 graphics, 
  broom,        # For tidying model outputs
  dplyr,        # For data manipulation
  gt,           # Nice beuatiful tables
  nortest,      # Anderson-Darling test for normality
  ggplot2,      # For creating plots
  DescTools,    # For statistical functions like skewness and kurtosis
  janitor,      # adding totals and percents to tables
  car,           # For Levene's test and ANOVA
  gtsummary,    # summary statistics and tests
  corrr,        # correlation analayis for numeric variables
  reshape2,     # For reshaping data
  skimr,        # get overview of data
  performance,           # For model performance metrics
  ResourceSelection,     # For goodness of fit tests
  broom,                 # For tidying model outputs
  pROC,                  # For ROC analysis
  caret,
  MASS,          # For statistical functions (polr for ordinal regression)ordinal
  ordinal,       # For statistical functions (ordinal regression)ordinal
  ggeffects,
  VGAM
            )
```

## Access the data
```{r read-data, echo = TRUE}
# file <- "./EpiData/metabolic_syndrome.dta"     #  use this if the file is on your computer
file <- "https://raw.githubusercontent.com/drkamarul/multivar_data_analysis/refs/heads/main/data/metabolic_syndrome.dta"
df4 <- import(file, trust  = TRUE)
summary(df4)
glimpse(df4)  # get an overview of the data
```

## Fix the data
```{r fix-data, echo=FALSE}
# change all chr variables to factor
df4 <- df4 %>% 
    mutate_if(is.character, as.factor)
glimpse(df4)

# Create new categorical variable from fbs (normal, impaired, diabetes)
df4 <- df4 %>%
  mutate(cat_fbs = cut(fbs, 
                       breaks = c(2.50 , 6.10, 28.01 ),
                       labels = c("Normal", "Diabetes")))
df4 %>% 
  count(cat_fbs)
glimpse(df4)

# Remove NAs from cat_fbs
df4 <- df4 %>%
  filter(!is.na(cat_fbs)) 
df4 %>%
  count(cat_fbs)

# Flip dirction of cat_fbs
#df4 <- df4 %>%
#  mutate(cat_fbs = fct_relevel(cat_fbs, 
#                               c("Diabetes", 'Impaired', 'Normal')))
# levels(df4$cat_fbs)


# Summary statistics
summary_stats <- df4 %>%
  dplyr::select(totchol, hpt, weight, cat_fbs) %>%
  tbl_summary(
    by = cat_fbs,  # Group by cat_fbs
    statistic = list(
      all_continuous() ~ "{mean} ({sd})"  # Mean and SD for continuous variables
      #all_continuous() ~ "{median} ({p25}, {p75})"  # Median and IQR
    ),
    digits = all_continuous() ~ 2,  # Round to 2 decimal places
    label = list(
      totchol ~ "Total Cholesterol (mg/dL)",
      hpt ~ "Hypertension Status",
      weight ~ "Weight (kg)"
    )
  ) %>%
  add_p() %>%  # Add p-values for group comparisons
  modify_caption("**Summary Statistics by Fasting Blood Sugar Category**")

summary_stats
```

## Multiple Logistic regression
```{r logistic_regression, echo=FALSE}
# Perform logistic regression
mlog_reg <- glm(cat_fbs ~ totchol + hpt + weight , data = df4, family = binomial(link = 'logit'))
#mlog_reg <- vglm(cat_fbs ~ totchol + hpt + weight, multinomial, data = df4, model = TRUE)
summary(mlog_reg)
```

## Model performance measures
```{r model-performance, echo=FALSE}
mlog_reg_perf <- vglm(cat_fbs ~ totchol + hpt + weight, multinomial, data = df4, model = TRUE)
# Get -2LL and AIC
minus_2LL_mlog_reg <- -2 * as.numeric(logLik(mlog_reg_perf))
AIC_valm <- AIC(mlog_reg_perf)

# Get Efron's R虏
efron <- PseudoR2(mlog_reg_perf, which = "Efron")
r2_valm <- as.numeric(efron)

# Disiplay the summary of the logistic regression model
summary(mlog_reg)
paste("R2= ",r2_valm)
```

## Calculate logOR and OR results with model performance 
```{r log_reg_tbl, echo = TRUE}
# Calculate log OR
mlog_reg_tbl <- tbl_regression(mlog_reg, exponentiate = FALSE)
# Calculate OR
mlog_reg_OR <- tbl_regression(mlog_reg, exponentiate = TRUE)

#Combine results
combined_tbl_mlog_reg <- tbl_merge(
  tbls = list(mlog_reg_tbl, mlog_reg_OR),
  tab_spanner = c("**log Odds**", "**Odds Ratio**")  # Optional: Add spanning headers
) %>%
 as_gt() %>%
  tab_source_note(

    source_note = paste0("R虏 = ", round(r2_valm,3),
                          ", -2LL = ", round(minus_2LL_mlog_reg, 0), 
                          ", AIC = ", round(AIC_valm, 0)
                        )
  )

# Display the combined table
 combined_tbl_mlog_reg
```



### --------------------------------------


## Multiple Logistic regression /w inteaction term
```{r logistic_regression, echo=FALSE}
# Perform logistic regression
mlog_reg_ia <- vglm(cat_fbs ~ totchol + hpt + weight + totchol*weight , multinomial, data = df4, model = TRUE)
summary(mlog_reg_ia)
```

## Model performance measures
```{r model-performance, echo=FALSE}
# Get -2LL and AIC
minus_2LL_mlog_reg_ia <- -2 * as.numeric(logLik(mlog_reg_ia))
AIC(mlog_reg_ia)

# Calculate Efron's R虏
r2_values_ia <- PseudoR2(mlog_reg_ia, which = "Efron")
r2_valm_ia <- r2_values_ia["Efron"]

# Disiplay the summary of the logistic regression model
summary(mlog_reg_ia)
r2_valm_ia
```

## Combine logOR and OR results with model performance 
```{r log_reg_tbl, echo = TRUE}
# Calculate log OR
mlog_reg_tbl_ia <- tbl_regression(mlog_reg_ia, exponentiate = FALSE)
# Calculate OR
mlog_reg_OR_ia <- tbl_regression(mlog_reg_ia, exponentiate = TRUE)

# combine results
combined_tbl_mlog_reg_ia <- tbl_merge(
  tbls = list(mlog_reg_tbl_ia, mlog_reg_OR_ia),
  tab_spanner = c("**log Odds**", "**Odds Ratio**")  # Optional: Add spanning headers
) %>%
  as_gt() %>%
  tab_source_note(

    source_note = paste0("R虏 = ", round(r2_valm_ia,3),
                          ", -2LL = ", round(minus_2LL_mlog_reg_ia, 0), 
                          ", AIC = ", round(AIC(mlog_reg_ia), 0)
                        )
  )

# Display the combined table
combined_tbl_mlog_reg
```

## Logistic multi-regression prediction
```{r log_reg-pred, echo = TRUE}
# Create a new data frame for prediction
predict_values <- data.frame(df4 = c(0, 1))

# Calculate predicted probabilities
predicted_probs <- predict(mlog_reg, type = "response")
df4$predicted_prob_multi <- predict(mlog_reg, type = "response")
predict_prob
df4$predicted_prob_multi

glimpse(df4)
```
