###
### name  :  专
### ID    : [住驻专 转注转 转]
### course: 驻 (HIT.ac.il)
### lesson: 住 住 07  RR 
### date  : 13/07/2025
###  
###  counts 专专住转 驻住 - 住驻专转


## Reset memory
```{r setup-packages, include = FALSE}
rm(list = ls(all.names = TRUE))       #  Clear all data objects         Clear packages
lapply(paste("package:", names(sessionInfo()$otherPkgs), sep = ""), detach, character.only = TRUE, unload = TRUE, force = TRUE)           # Clear packages from memory
gc()                                  # Clear unused data from memory ("garbage cllection")
dev.off()                             # Reset the display output of R
q("no")                               # This will completely shutdown R terminal
```

## Load required packages
```{r load-packages, include = FALSE}
options("install.lock"=FALSE)
if (!require("pacman")) install.packages("pacman") ; library(pacman)
pacman::p_load(
  Rcpp, devtools, gh, remotes,   # For installing packages/files online (like from GitHub)
  rio, here,              # File import and path support
  labelled, DT,           # deal with labels
  tidyverse,              # data management + ggplot2 graphics,
  broom,                  # For tidying model outputs
  dplyr,                  # For data manipulation
  gt,                     # Nice beuatiful tables
  gtsummary,              # summary statistics and tests
  nortest,                # Anderson-Darling test for normality
  ggplot2,                # For creating plots
  ggpubr,                 # For ggplot2 extensions
  DescTools,              # For statistical functions like skewness and kurtosis
  summarytools,           # to get a table of all summary statistics
  car,                    # For Levene's test and ANOVA
  lmtest,                 # do linear regression tests
  parameters,             # get standardized parameters/coefficients
  GGally,                 # to make a correlation matrix
  skimr,                  # get overview of data
  pROC,                   # For ROC analysis
  cvms,                   # For confusion matrix and classification metrics
  ggeffects,              # To display the confusion matrix
  performance,            # For model performance metrics
  generalhoslem,          # For Hosmer-Lemeshow test for goodness of fit 
  # MASS,                   # For statistical functions (polr for ordinal regression)ordinal
  ordinal,                # For ordinal regression statistical functions 
  generalhoslem,          # For Hosmer-Lemeshow test for goodness of fit 
  ResourceSelection,      # For Hosmer-Lemeshow test for goodness of fit 
  VGAM,                   # for multinomial logistic regression
  cardx,
  epiDisplay             # For epi data type handling
            )
# Install/load flexplot
if (!require("flexplot", quietly = TRUE)) devtools::install_github("dustinfife/flexplot", ref = "logistic_plots")  ;  library(flexplot)
```

## Access the data
```{r read-data, echo = TRUE}
## read the data in a file online
# file <- "./EpiData/07 asthma.rds"     #  use this if the file is on your computer
file <- "https://raw.githubusercontent.com/Model-Lab-Net/Courses/refs/heads/main/Epi/Data/07%20asthma.rds"
df6 <- import(file, trust  = TRUE)
# show variable translations to Hebrew in lables
labels <- data.frame(Variable = names(df6), Label = unlist(var_label(df6)))   ;   gt(labels)
glimpse(df6)               # get an overview of the data
```

## Get summary stats
```{r summary-stats, echo = TRUE}
# read the data in a file online
epiDisplay::codebook(df6)
```

## Plot the data
```{r simple-poisson-reg, echo = TRUE}
flexplot(attack ~ 1, data = df6)
```

## Simple Poisson regression univariate model
```{r simple-poisson-reg, echo = TRUE}
# Do a univrariate Poisson regression - attack vs gender
spr1 <- glm(ghq12 ~ attack, data = df6, family = "poisson")
summary(spr1)
flexplot(attack ~ ghq12, data = df6, method = "poisson", ghost.line = "red", jitter=c(0,0.05))
```

## Multivariate Poisson regression model  
```{r poisson-multi-reg, echo = TRUE}
# Do a multivariate Poisson regression - attack vs everything
mpr <- glm(ghq12 ~ attack + res.inf, data = df6, family = "poisson")
summary(spr3)
flexplot(attack ~ ghq12 | res.inf, data = df6, method = "poisson", ghost.line = "red", jitter=c(0,0.05))
```

## Model comparison
```{r model-comparison, echo = TRUE}
# Uaing just flexplot
model.comparison(spr1, mpr)
summary(mpr)

visualize(mpr)
flexplot(attack ~ ghq12 + res.inf, data = df6, method = "poisson", ghost.line = "red", jitter=c(0,0.05))
```

## Model fit and performance
```{r model-comparison, echo = TRUE}
estimates(mpr)

# Uaing just flexplot
visualize(mpr)

# Scaled Pearson chi-square statistic

# Chi-square goodness-of-fit
epiDisplay::poisgof(mpr)

# AIC
AIC(mpr)
```

## Show model summary
```{r show-summary, echo = TRUE}
# summary of Poisson model
tbl_regression(mpr, exponentiate = TRUE)
round(summary(mpr)$coefficients, 2)
```

## Model prediction
```{r predict-data, echo = TRUE}
# expected number of asthmatic attacks per year for a patient with recurrent respiratory infection and GHQ-12 score of 8
predict <- predict(mpr, list(res.inf = "yes", ghq12 = 8),type = "response")
```





### -------------------------------------------------
## 转专-注爪

## 砖 1
```{r question-1, echo = TRUE}
# 

```

## 砖 2
```{r question-2, echo = TRUE}
# 

```

## 砖 3
```{r question-3, echo = TRUE}
# 

```









### source:  https://bookdown.org/drki_musa/dataanalysis/poisson-regression.html
###          https://github.com/drkamarul/multivar_data_analysis/tree/main/data

## Clean data
```{r clean-data, echo = TRUE}
# Load the data
file <- "https://raw.githubusercontent.com/drkamarul/multivar_data_analysis/refs/heads/main/data/asthma.csv"
df6 <- import(file, trust  = TRUE)
#write.csv(df6, here("EpiData", "07 asthma.csv"))

# Clean up variable names
df6 <- clean_names(df6)
names(df6) <- gsub("_", ".", names(df6))

# Check the updated dataset structure
glimpse(df6)

# Save as R data file
saveRDS(df6, here("EpiData", "07 asthma.rds"))
```

## Give variable explanations in labels
```{r labels, echo = TRUE}
attr(df6$gender, "label") <- "专"
attr(df6$res.inf, "label") <- "拽转 专 砖"
attr(df6$ghq12, "label") <- " 驻住 (砖 拽砖专  爪拽 驻砖转 转拽驻 住转)"
attr(df6$attack, "label") <- "转拽祝"

labels <- NULL
labels <- data.frame(Variable = names(df6), Label = unlist(var_label(df6)))
gt(labels)
```










### ----------------------------------------------



### question 1
```{r read-data, echo = TRUE}
# Create 2x2 table

df <- na.omit(data[, c("DisObes", "stroke")])

# Now compute the table
tab <- table(df$DisObes, df$stroke)


# Assign cells
a <- tab[1,1]  # Not obese, no stroke
b <- tab[1,2]  # Not obese, stroke
c <- tab[2,1]  # Obese, no stroke
d <- tab[2,2]  # Obese, stroke

# Calculate risks
risk_obese <- d / (c + d)
risk_not_obese <- b / (a + b)

# Relative Risk
RR <- risk_obese / risk_not_obese
# Standard error of log(RR)
se_log_rr <- sqrt( (1/b) - (1/(a + b)) + (1/d) - (1/(c + d)) )

# Z for 95% CI
z <- qnorm(0.975)

# Confidence Interval (log scale)
log_rr <- log(RR)
ci_lower <- exp(log_rr - z * se_log_rr)
ci_upper <- exp(log_rr + z * se_log_rr)
z_stat <- log_rr / se_log_rr
p_value <- 2 * (1 - pnorm(abs(z_stat)))

RR
ci_lower
ci_upper
p_value



```



### question 2
```{r setup, include=FALSE}
death_freq <- table(data$Death)
death_percent <- prop.table(death_freq) * 100

death_percent

df <- na.omit(data[, c("Death", "stroke", "SexCode", "DisObes", "DisSmok", "DisDM")])


model1 <- glm(Death ~ SexCode + DisSmok , data = df, family = binomial())

OR<-exp(coef(model1))


model2 <- glm(Death ~ SexCode + DisSmok , data = df, family = poisson())

RR <- exp(coef(model2))

OR
RR


####stroke

stroke_freq <- table(data$stroke)
stroke_percent <- prop.table(stroke_freq) * 100

stroke_percent



model1 <- glm(stroke~ SexCode + DisSmok , data = df, family = binomial())

OR<-exp(coef(model1))


model2 <- glm(stroke ~ SexCode + DisSmok , data = df, family = poisson())

RR <- exp(coef(model2))

OR
RR



```

#### question 3
```{r setup, include=FALSE}
# Load required packages
install.packages("sandwich")
install.packages("lmtest")
library(sandwich)
library(lmtest)

# Clean data
df <- na.omit(data[, c("stroke", "SexCode", "DisObes", "DisSmok", "DisDM")])

# Fit Poisson model
model <- glm(stroke ~ SexCode + DisObes + DisSmok + DisDM, data = df, family = poisson())

# Get robust standard errors
robust_se <- sqrt(diag(vcovHC(model, type = "HC0")))
coeftest(model, vcov. = vcovHC(model, type = "HC0"))
# Calculate RR and 95% CI
RR <- exp(coef(model))
CI_lower <- exp(coef(model) - 1.96 * robust_se)
CI_upper <- exp(coef(model) + 1.96 * robust_se)

RR
CI_lower
CI_upper


# Deviance and Pearson chi-square
deviance <- deviance(model)
df_resid <- df.residual(model)

pearson <- sum(residuals(model, type = "pearson")^2)

cat("Deviance =", deviance, "on", df_resid, "df\n")
cat("Pearson Chi-square =", pearson, "on", df_resid, "df\n")
```



### id deviance/df>1.5 than:
# Fit quasi-Poisson model
```{r setup, include=FALSE}

model_qp <- glm(stroke ~ SexCode + DisObes + DisSmok + DisDM, data = df, family = quasipoisson())

# Coefficients
coefs <- coef(model_qp)

# Standard errors (already adjusted for dispersion)
se_qp <- summary(model_qp)$coefficients[, "Std. Error"]

# RR (exponentiated coefficients)
RR <- exp(coefs)

# 95% CI using Wald method
CI_lower <- exp(coefs - 1.96 * se_qp)
CI_upper <- exp(coefs + 1.96 * se_qp)

  RR = RR
  CI_lower
  CI_upper
 
```


