###
### name: דוד בורג
### ID  : [מספר תעודת זהות]
### course: אפידמיולוגיה (HIT.ac.il)
### lesson: סיכון יחסי 07  RR 
### date  : 15/06/2025
###  
###  rates רגרסיית פאוסון - היערות
###
### source:  https://bookdown.org/drki_musa/dataanalysis/poisson-regression.html
###          https://github.com/drkamarul/multivar_data_analysis/tree/main/data


## Reset memory
```{r setup-packages, include = FALSE}
rm(list = ls(all.names = TRUE))
lapply(paste('package:',names(sessionInfo()$otherPkgs),sep=""),detach,character.only=TRUE,unload=TRUE,force=TRUE)
gc()
dev.off()
```

## Load required packages
```{r load-packages, include = FALSE}
options("install.lock"=FALSE)
if (!require("pacman")) install.packages("pacman")
library(pacman)
pacman::p_load(
  rio,          # File import
  tidyverse,    # data management + ggplot2 graphics, 
  gt,           # Nice beuatiful tables
  nortest,      # Anderson-Darling test for normality
  ggplot2,      # For creating plots
  DescTools,    # For statistical functions like skewness and kurtosis
  janitor,      # adding totals and percents to tables
  car,           # For Levene's test and ANOVA
  gtsummary,    # summary statistics and tests
  corrr,        # correlation analayis for numeric variables
  dplyr,        # For data manipulation
  reshape2,     # For reshaping data
  skimr,        # get overview of data
  performance,           # For model performance metrics
  ResourceSelection,     # For goodness of fit tests
  broom,                 # For tidying model outputs
  pROC,                  # For ROC analysis
  caret,
  MASS,          # For statistical functions (polr for ordinal regression)
  ordinal,       # For statistical functions (ordinal regression)ordinal
  ggeffects,
  epiDisplay    # 

            )
```

## Access the data
```{r read-data, echo = TRUE}
## read the data in a file online
# file <- "./EpiData/smoke.csv"     # use this if the file is on your computer
file <- "https://raw.githubusercontent.com/drkamarul/multivar_data_analysis/refs/heads/main/data/smoke.csv"
df7 <- import(file, trust  = TRUE)
summary(df7)
skim(df7)
```


## Add observed rates
```{r radd-obs, echo = TRUE}
## read the data in a file online
df7 <- df7 %>% 
      mutate(rate = round(case/person_yrs, 4))
glimpse(df7)
```

## Do summary stats
```{r summary-stats, echo = TRUE}
## read the data in a file online
epiDisplay::codebook(df7)
```

## A simple Poisson regression univariate analysis for rates
```{r read-data, echo = TRUE}
## read the data in a file online
spr1 <- glm(case ~ cigar_day, data = df7, family = "poisson", offset = log(person_yrs))
summary(spr1)

spr2 <- glm(case ~ smoke_yrs, data = df7, family = "poisson", offset = log(person_yrs))
summary(spr2)
```

##  Poisson multivariate regression analysis for rates
```{r poisson-multi-reg, echo = TRUE}
## read the data in a file online
mpr_all <- glm(case ~ cigar_day + smoke_yrs, data = df7, family = "poisson", offset = log(person_yrs))    # the "." uses all the other variables
summary(mpr_all)
```

## Multiple Poisson regression with interaction for rates
```{r poisson-multi-reg-ia, echo = TRUE}
## read the data in a file online
mpr_ia <- glm(case ~ cigar_day:smoke_yrs, data = df7, family = "poisson", offset = log(person_yrs))
summary(mpr_ia)
```

## Model fit and perfoarmance
```{r read-data, echo = TRUE}
#Scaled Pearson chi-square statistic

# Chi-square goodness-of-fit
epiDisplay::poisgof(mpr_ia)
# AIC
AIC(mpr_ia)
```

## Show summary
```{r read-data, echo = TRUE}
tbl_regression(mpr_all, exponentiate = TRUE)
round(summary(mpr_all)$coefficients, 2)
```






###question 1

# Create 2x2 table

df <- na.omit(data[, c("DisObes", "stroke")])

# Now compute the table
tab <- table(df$DisObes, df$stroke)


# Assign cells
a <- tab[1,1]  # Not obese, no stroke
b <- tab[1,2]  # Not obese, stroke
c <- tab[2,1]  # Obese, no stroke
d <- tab[2,2]  # Obese, stroke

# Calculate risks
risk_obese <- d / (c + d)
risk_not_obese <- b / (a + b)

# Relative Risk
RR <- risk_obese / risk_not_obese
# Standard error of log(RR)
se_log_rr <- sqrt( (1/b) - (1/(a + b)) + (1/d) - (1/(c + d)) )

# Z for 95% CI
z <- qnorm(0.975)

# Confidence Interval (log scale)
log_rr <- log(RR)
ci_lower <- exp(log_rr - z * se_log_rr)
ci_upper <- exp(log_rr + z * se_log_rr)
z_stat <- log_rr / se_log_rr
p_value <- 2 * (1 - pnorm(abs(z_stat)))

RR
ci_lower
ci_upper
p_value
```



###question 2
```{r setup, include=FALSE}
death_freq <- table(data$Death)
death_percent <- prop.table(death_freq) * 100

death_percent

df <- na.omit(data[, c("Death", "stroke", "SexCode", "DisObes", "DisSmok", "DisDM")])


model1 <- glm(Death ~ SexCode + DisSmok , data = df, family = binomial())

OR<-exp(coef(model1))


model2 <- glm(Death ~ SexCode + DisSmok , data = df, family = poisson())

RR <- exp(coef(model2))

OR
RR


####stroke

stroke_freq <- table(data$stroke)
stroke_percent <- prop.table(stroke_freq) * 100

stroke_percent



model1 <- glm(stroke~ SexCode + DisSmok , data = df, family = binomial())

OR<-exp(coef(model1))


model2 <- glm(stroke ~ SexCode + DisSmok , data = df, family = poisson())

RR <- exp(coef(model2))

OR
RR



```
####question 3
```{r setup, include=FALSE}
# Load required packages
install.packages("sandwich")
install.packages("lmtest")
library(sandwich)
library(lmtest)

# Clean data
df <- na.omit(data[, c("stroke", "SexCode", "DisObes", "DisSmok", "DisDM")])

# Fit Poisson model
model <- glm(stroke ~ SexCode + DisObes + DisSmok + DisDM, data = df, family = poisson())

# Get robust standard errors
robust_se <- sqrt(diag(vcovHC(model, type = "HC0")))
coeftest(model, vcov. = vcovHC(model, type = "HC0"))
# Calculate RR and 95% CI
RR <- exp(coef(model))
CI_lower <- exp(coef(model) - 1.96 * robust_se)
CI_upper <- exp(coef(model) + 1.96 * robust_se)

RR
CI_lower
CI_upper


# Deviance and Pearson chi-square
deviance <- deviance(model)
df_resid <- df.residual(model)

pearson <- sum(residuals(model, type = "pearson")^2)

cat("Deviance =", deviance, "on", df_resid, "df\n")
cat("Pearson Chi-square =", pearson, "on", df_resid, "df\n")
```



###id deviance/df>1.5 than:
# Fit quasi-Poisson model
```{r setup, include=FALSE}

model_qp <- glm(stroke ~ SexCode + DisObes + DisSmok + DisDM, data = df, family = quasipoisson())

# Coefficients
coefs <- coef(model_qp)

# Standard errors (already adjusted for dispersion)
se_qp <- summary(model_qp)$coefficients[, "Std. Error"]

# RR (exponentiated coefficients)
RR <- exp(coefs)

# 95% CI using Wald method
CI_lower <- exp(coefs - 1.96 * se_qp)
CI_upper <- exp(coefs + 1.96 * se_qp)

  RR = RR
  CI_lower
  CI_upper
 



```


