###
### name  : __________________
### ID    : [מספר תעודת זהות]
### course: אפידמיולוגיה (HIT.ac.il)
### lesson: סיכון יחסי 07  Rates 
### date  : 19/04/2026
###  
###  rates רגרסיית פאוסון - היערות
###
### source: https://bookdown.org/drki_musa/dataanalysis/poisson-regression.html#poisson-regression-for-rate


## Reset memory
```{r setup-packages, include = FALSE}
rm(list = ls(all.names = TRUE))          # 👈 Clear all data objects        👇 Clear packages
lapply(paste("package:", names(sessionInfo()$otherPkgs), sep = ""), detach, character.only = TRUE, unload = TRUE, force = TRUE)
gc()                                     # Clear unused data from memory ("garbage cllection")
dev.off()                                # Reset the display output of R
system("powershell -Command Clear-Host") # This will clear the terminal in VScode
```

## Load required packages
```{r load-packages, include = FALSE}
options("install.lock"=FALSE)
if (!require("pacman")) install.packages("pacman") ; library(pacman)
pacman::p_load(
  Rcpp, devtools, gh, remotes,   # For installing packages/files online (like from GitHub)
  rio, here,              # File import and path support
  labelled, DT,           # deal with labels
  tidyverse,              # data management + ggplot2 graphics,
  broom,                  # For tidying model outputs
  dplyr,                  # For data manipulation
  gt,                     # Nice beuatiful tables
  gtsummary,              # summary statistics and tests
  nortest,                # Anderson-Darling test for normality
  ggplot2,                # For creating plots
  ggpubr,                 # For ggplot2 extensions
  DescTools,              # For statistical functions like skewness and kurtosis
  summarytools,           # to get a table of all summary statistics
  car,                    # For Levene's test and ANOVA
  lmtest,                 # do linear regression tests
  parameters,             # get standardized parameters/coefficients
  GGally,                 # to make a correlation matrix
  skimr,                  # get overview of data
  pROC,                   # For ROC analysis
  cvms,                   # For confusion matrix and classification metrics
  ggeffects,              # To display the confusion matrix
  performance,            # For model performance metrics
  generalhoslem,          # For Hosmer-Lemeshow test for goodness of fit 
  # MASS,                   # For statistical functions (polr for ordinal regression)ordinal
  ordinal,                # For ordinal regression statistical functions 
  generalhoslem,          # For Hosmer-Lemeshow test for goodness of fit 
  ResourceSelection,      # For Hosmer-Lemeshow test for goodness of fit 
  VGAM,                   # for multinomial logistic regression
  cardx,
  epiDisplay,             # For epi data type handling
  AER
            )
# Install/load flexplot
if (!require("flexplot", quietly = TRUE)) devtools::install_github("dustinfife/flexplot", ref = "logistic_plots")  ;  library(flexplot)
```

## Access the data
```{r read-data, echo = TRUE}
## read the data in a file online
# file <- "./EpiData/08 smoke.rds"     # 👈 use this if the file is on your computer
file <- "https://raw.githubusercontent.com/Model-Lab-Net/Courses/refs/heads/main/Epi/Data/08%20smoke.rds"
df7 <- import(file, trust  = TRUE)
# show variable translations to Hebrew in lables
labels <- data.frame(Variable = names(df7), Label = unlist(var_label(df7)))   ;   gt(labels)
glimpse(df7)               # get an overview of the data
```

## Get summary stats
```{r summary-stats, echo = TRUE}
# Get summary stastistics
epiDisplay::codebook(df7)
```


## Plot the data
```{r simple-poisson-reg, echo = TRUE}
flexplot(case ~ 1, data = df7)
```

## A simple Poisson regression univariate model for rates
```{r read-data, echo = TRUE}
# read the data in a file online
spr <- glm(case ~ cigar.day,
           data = df7,
           family = "poisson",
           offset = log(person.yrs))
summary(spr)

spr_model <- tbl_regression(spr, exponentiate = TRUE)
spr_model
```

##  Poisson multivariate regression model for rates
```{r poisson-multi-reg, echo = TRUE}
# read the data in a file online
mpr <- glm(case ~ cigar.day + smoke.yrs,
           data = df7,
           family = "poisson",
           offset = log(person.yrs))    # the "." uses all the other variables
summary(mpr)

mpr_model <- tbl_regression(mpr, exponentiate = TRUE)
mpr_model
```

## Multiple Poisson regression with interaction for rates
```{r poisson-multi-reg-ia, echo = TRUE}
# read the data in a file online
mpr_ia <- glm(case ~ cigar.day * smoke.yrs,
              data = df7,
              family = "poisson", 
              offset = log(person.yrs))
summary(mpr_ia)

mpr_ia_model <- tbl_regression(mpr_ia, exponentiate = TRUE)
mpr_ia_model
```

## Model fit and performance
```{r read-data, echo = TRUE}
#Scaled Pearson chi-square statistic

# Chi-square goodness-of-fit
epiDisplay::poisgof(mpr_ia)
# AIC
AIC(mpr_ia)
```


## Model comparison
```{r mdodel-comparison, echo = TRUE}
# Compare models
model.comparison(spr, mpr)
model.comparison(spr, mpr_ia)
model.comparison(mpr, mpr_ia)
```

## Model diagnostics
```{r mpr-logisitc-reg, echo = TRUE}
dd <- datadist(df7)   ;   options(datadist = "dd")
ord_model_ <- orm(perforation.ord ~ pulp + hemoglobin + systolic, data = df5)     # need this for model diagnostics. sorry.
# NEW way
AIC_val  <- as.numeric(AIC(mpr))                       # AIC
r2_val   <- as.numeric(r2_nagelkerke(mpr))
# Test dspersion
dt_p <- dispersiontest(mpr)$p.value %>%
                    format.pval(dt_p, eps = 0.001) #%>% 
                    #paste("p", ., sep = " ")
dt <- as.numeric(dispersiontest(mpr)$estimate) 
# Test dspersion
zi_check <- check_zeroinflation(mpr)

# Calculate Omnibus chi-square test for goodness of fit
chi_sq <- mpr$null.deviance - mpr$deviance
df_diff <- mpr$df.null - mpr$df.residual
p_val <- pchisq(chi_sq, df = df_diff, lower.tail = FALSE) %>%
                    format.pval(p_val, eps = 0.001) #%>% 
                    #paste("p", ., sep = " ")

# Add model diagnostics to table
mpr_model %>%
  as_gt() %>% # Convert to gt engine first
  tab_source_note(
    source_note = md(paste0(
      "R² = ", round(r2_val, 3), "<br>",
      "AIC = ", round(AIC_val, 0), "<br>",
      "Dispersion test = ", round(dt, 1), 
      ",     p  ", dt_p, "<br>",
      "Omnibus test: p ", p_val
    ))
  )
```





### -------------------------------------------------
## תרגול-עצמי

## שאלה 1
```{r question-1, echo = TRUE}
# 

```

## שאלה 2
```{r question-2, echo = TRUE}
# 

```

## שאלה 3
```{r question-3, echo = TRUE}
# 

```









### source:  https://bookdown.org/drki_musa/dataanalysis/poisson-regression.html#poisson-regression-for-rate
###          https://github.com/drkamarul/multivar_data_analysis/tree/main/data

## Clean data
```{r clean-data, echo = TRUE}
# Load the data
file <- "https://raw.githubusercontent.com/drkamarul/multivar_data_analysis/refs/heads/main/data/smoke.csv"
df7 <- import(file, trust  = TRUE)
#write.csv(df7, here("EpiData", "08 smoke.csv"))

# Clean up variable names
df7 <- clean_names(df7)
names(df7) <- gsub("_", ".", names(df7))

# Check the updated dataset structure
glimpse(df7)

# Save as R data file
saveRDS(df7, here("EpiData", "08 smoke.rds"))
```

## Give variable explanations in labels
```{r labels, echo = TRUE}
attr(df7$smoke.yrs, "label") <- "מספר שנים של עישון"
attr(df7$cigar.day, "label") <- "מספר סיגריות ביום"
attr(df7$person.yrs, "label") <- "יחידת מידה המשלבת את מספר הפרטים במחקר עם משך הזמן בו הם נצפו"
attr(df7$case, "label") <- "מספר מקרים של סרטן הריאות"

labels <- NULL
labels <- data.frame(Variable = names(df7), Label = unlist(var_label(df7)))
gt(labels)
```