###
### name  : __________________
### ID    : [מספר תעודת זהות]
### course: אפידמיולוגיה (HIT.ac.il)
### lesson: סיכון יחסי 07  RR 
### date  : 19/04/2026
###  
###  quasi רגרסיית פאוסון 
###
### source: https://bookdown.org/drki_musa/dataanalysis/poisson-regression.html#quasi-poisson-regression-for-overdispersed-data 


## Reset memory
```{r setup-packages, include = FALSE}
rm(list = ls(all.names = TRUE))          # 👈 Clear all data objects        👇 Clear packages
lapply(paste("package:", names(sessionInfo()$otherPkgs), sep = ""), detach, character.only = TRUE, unload = TRUE, force = TRUE)
gc()                                     # Clear unused data from memory ("garbage cllection")
dev.off()                                # Reset the display output of R
system("powershell -Command Clear-Host") # This will clear the terminal in VScode
```

## Load required packages
```{r load-packages, include false}
options("install.lock"=FALSE)
if (!require("pacman")) install.packages("pacman") ; library(pacman)
pacman::p_load(
  Rcpp, devtools, gh, remotes,   # For installing packages/files online (like from GitHub)
  rio, here,              # File import and path support
  labelled, DT,           # deal with labels
  tidyverse,              # data management + ggplot2 graphics,
  broom,                  # For tidying model outputs
  dplyr,                  # For data manipulation
  gt,                     # Nice beuatiful tables
  gtsummary,              # summary statistics and tests
  nortest,                # Anderson-Darling test for normality
  ggplot2,                # For creating plots
  ggpubr,                 # For ggplot2 extensions
  DescTools,              # For statistical functions like skewness and kurtosis
  summarytools,           # to get a table of all summary statistics
  car,                    # For Levene's test and ANOVA
  lmtest,                 # do linear regression tests
  parameters,             # get standardized parameters/coefficients
  GGally,                 # to make a correlation matrix
  skimr,                  # get overview of data
  pROC,                   # For ROC analysis
  cvms,                   # For confusion matrix and classification metrics
  ggeffects,              # To display the confusion matrix
  performance,            # For model performance metrics
  generalhoslem,          # For Hosmer-Lemeshow test for goodness of fit 
  # MASS,                   # For statistical functions (polr for ordinal regression)ordinal
  ordinal,                # For ordinal regression statistical functions 
  generalhoslem,          # For Hosmer-Lemeshow test for goodness of fit 
  ResourceSelection,      # For Hosmer-Lemeshow test for goodness of fit 
  VGAM,                   # for multinomial logistic regression
  cardx,
  epiDisplay,             # For epi data type handling
  AER
            )
# Install/load flexplot
if (!require("flexplot", quietly = TRUE)) devtools::install_github("dustinfife/flexplot", ref = "logistic_plots")  ;  library(flexplot)
```

## Access the data
```{r read-data, echo=FALSE}
## read the data in a file online
# file <- "./EpiData/07 asthma.rds"     # 👈 use this if the file is on your computer
file <- "https://raw.githubusercontent.com/Model-Lab-Net/Courses/refs/heads/main/Epi/Data/07%20asthma.rds"
df6 <- import(file, trust  = TRUE)
# show variable translations to Hebrew in lables
labels <- data.frame(Variable = names(df6), Label = unlist(var_label(df6)))   ;   gt(labels)
glimpse(df6)               # get an overview of the data
```

## Do summary stats
```{r read-data, echo=FALSE}
## read the data in a file online
epiDisplay::codebook(df6)
```

## Poisson regression multivariate model for quasi (when “mean equals variance” assumption is violated)
```{r read-data, echo=FALSE}
## Perform a "quasi"-poisson model
spr <- glm(attack ~ ., data = df6, family = "poisson")
summary(spr)
spr_model <- tbl_regression(spr, exponentiate = TRUE)
spr_model
```

## Poisson regression multivariate model for quasi (when “mean equals variance” assumption is violated)
```{r read-data, echo=FALSE}
## Perform a "quasi"-poisson model
qpr <- glm(attack ~ ., data = df6, family = "quasipoisson")
summary(qpr)
qpr_model <- tbl_regression(qpr, exponentiate = TRUE)
qpr_model
```

## Model diagnostics
```{r mpr-logisitc-reg, echo = TRUE}
dd <- datadist(df7)   ;   options(datadist = "dd")
ord_model_ <- orm(perforation.ord ~ pulp + hemoglobin + systolic, data = df5)     # need this for model diagnostics. sorry.
# NEW way
AIC_val  <- as.numeric(AIC(qpr))                       # AIC
r2_val   <- as.numeric(r2_nagelkerke(qpr))
# Calculate Omnibus chi-square test for goodness of fit
chi_sq <- qpr$null.deviance - qpr$deviance
df_diff <- qpr$df.null - qpr$df.residual
p_val <- pchisq(chi_sq, df = df_diff, lower.tail = FALSE) %>%
                    format.pval(p_val, eps = 0.001) #%>% 
                    #paste("p", ., sep = " ")

# Add model diagnostics to table
qpr_model %>%
  as_gt() %>% # Convert to gt engine first
  tab_source_note(
    source_note = md(paste0(
      "R² = ", round(r2_val, 3), "<br>",
      "AIC = ", round(AIC_val, 0), "<br>",
      "Omnibus test: p ", p_val
    ))
  )
```

## Compare to simple Poisson regression
```{r simple-poisson-reg, echo = TRUE}
# Do a univrariate Poisson regression - attack vs gender
tbl_merge(
  tbls = list(spr_model, qpr_model),
  tab_spanner  = c("**SPR**", "**QPR**")
) %>%
  as_gt()
combined_tbl
```

## Model comparison
```{r mdodel-comparison, echo = TRUE}
# Check all combinations to select "best" model
model.comparison(spr, qpr)
```







### -------------------------------------------------
## תרגול-עצמי

## שאלה 1
```{r question-1, echo = TRUE}
# 

```

## שאלה 2
```{r question-2, echo = TRUE}
# 

```

## שאלה 3
```{r question-3, echo = TRUE}
# 

```








### source:  https://bookdown.org/drki_musa/dataanalysis/poisson-regression.html#quasi-poisson-regression-for-overdispersed-data
###          https://github.com/drkamarul/multivar_data_analysis/tree/main/data