###
### name  : __________________
### ID    : [住驻专 转注转 转]
### course: 驻 (HIT.ac.il)
### lesson: 砖专转 08   Survival
### date  : 13/07/2025
###  
###   爪注 转 砖专转

## Reset memory
```{r setup-packages, include = FALSE}
rm(list = ls(all.names = TRUE))          #  Clear all data objects         Clear packages
lapply(paste("package:", names(sessionInfo()$otherPkgs), sep = ""), detach, character.only = TRUE, unload = TRUE, force = TRUE)
gc()                                     # Clear unused data from memory ("garbage cllection")
dev.off()                                # Reset the display output of R
system("powershell -Command Clear-Host") # This will clear the terminal in VScode
```

## Load required packages
```{r load-packages, include = FALSE}
options("install.lock"=FALSE)
if (!require("pacman")) install.packages("pacman") ; library(pacman)
pacman::p_load(
  Rcpp, devtools, gh, remotes,   # For installing packages/files online (like from GitHub)
  rio, here,              # File import and path support
  labelled, DT,           # deal with labels
  tidyverse,              # data management + ggplot2 graphics,
  broom,                  # For tidying model outputs
  dplyr,                  # For data manipulation
  gt,                     # Nice beuatiful tables
  gtsummary,              # summary statistics and tests
  nortest,                # Anderson-Darling test for normality
  ggplot2,                # For creating plots
  ggpubr,                 # For ggplot2 extensions
  DescTools,              # For statistical functions like skewness and kurtosis
  summarytools,           # to get a table of all summary statistics
  car,                    # For Levene's test and ANOVA
  lmtest,                 # do linear regression tests
  parameters,             # get standardized parameters/coefficients
  GGally,                 # to make a correlation matrix
  skimr,                  # get overview of data
  pROC,                   # For ROC analysis
  cvms,                   # For confusion matrix and classification metrics
  ggeffects,              # To display the confusion matrix
  performance,            # For model performance metrics
  generalhoslem,          # For Hosmer-Lemeshow test for goodness of fit 
  # MASS,                   # For statistical functions (polr for ordinal regression)ordinal
  ordinal,                # For ordinal regression statistical functions 
  generalhoslem,          # For Hosmer-Lemeshow test for goodness of fit 
  ResourceSelection,      # For Hosmer-Lemeshow test for goodness of fit 
  VGAM,                   # for multinomial logistic regression
  cardx,
  epiDisplay,             # For epi data type handling
  survival,               # For survival analysis
  survminer               # For survival analysis and visualization
             )
# Install/load flexplot
if (!require("flexplot", quietly = TRUE)) devtools::install_github("dustinfife/flexplot", ref = "logistic_plots")  ;  library(flexplot)
```

## Access the data
```{r read-data, echo = TRUE}
## read the data in a file online
# file <- "./EpiData/09 stroke_data.rds"     #  use this if the file is on your computer
file <- "https://raw.githubusercontent.com/Model-Lab-Net/Courses/refs/heads/main/Epi/Data/09%20stroke_data.rds"
df8 <- import(file, trust  = TRUE)
# show variable translations to Hebrew in lables
labels <- data.frame(Variable = names(df8), Label = unlist(var_label(df8)))   ;   gt(labels)
glimpse(df8)               # get an overview of the data
```

## Show the data
```{r read-data, echo = TRUE}
# in a nice table
df8 %>%
  tbl_summary(
    by = status,
    include = -c(doa, dod)  # exclude these two date variables
  ) %>%
  as_gt()


df8 %>%
  tbl_summary(
    by = stroke.type,
    include = -c(doa, dod)  # exclude these two date variables
  ) %>%
  as_gt()
```

## Survival time statistics
```{r surv-stats, echo = TRUE}
# Survival time
mean_survival_time <- mean(df8$time2, na.rm = TRUE)
sd_survival_time <- sd(df8$time2, na.rm = TRUE)
CI1 <- MeanCI(df8$time2, na.rm = TRUE)
  lower_meanci_1 <- as.numeric((CI1["lwr.ci"]))
  upper_meanci_1 <- as.numeric((CI1["upr.ci"]))
median_survival_time <- median(df8$time2, na.rm = TRUE)
IQR_survival_time <- IQR(df8$time2, na.rm = TRUE)
q25_survival_time <- as.numeric(Quantile(df8$time2, probs = 0.25, na.rm = TRUE))
q75_survival_time <- as.numeric(Quantile(df8$time2, probs = 0.75, na.rm = TRUE))
min_survival_time <- min(df8$time2, na.rm = TRUE)
max_survival_time <- max(df8$time2, na.rm = TRUE)

# Create a data frame to neatly display the statistics
stats_df <- data.frame(
    Statistic = c("Mean", "SD", "CI low", "CI high", "Median", "Q1", "Q3", "IQR", "Minimum", "Maximum")
  , Value = c(mean_survival_time, sd_survival_time, lower_meanci_1, upper_meanci_1, median_survival_time, q25_survival_time, q75_survival_time, IQR_survival_time, min_survival_time, max_survival_time)
  , 住住 = c("爪注", "住'转", "专 住", "专 住", "爪", "专注 1", "专注 3", "专  专注", "", "拽住")
)
glimpse(stats_df)

# Clean, simple table
summary_stats <- stats_df %>% 
  gt() %>%
  tab_header(title = "Survival Statistics 砖专转") %>%
  fmt_number(columns = Value, decimals = 1)
summary_stats              # Show the table
```

## Do a Kaplan-Meier survival analayis
```{r km-surv, echo = TRUE}
# estimate the survival probabilities for all patients
KM_all <- survfit(Surv(time = time2, 
                       event = status == "dead" ) ~ 1, 
                       data = df8)
summary(KM_all)
```


## Life table
```{r life-table, echo = TRUE}
KM_table <- with(summary(KM_all),
                 data.frame(
                   time = time,
                   n_risk = n.risk,
                   n_event = n.event,
                   n_censor = n.censor,
                   surv = surv,
                   std_err = std.err,
                   lower_ci = lower,
                   upper_ci = upper
                 ))
KM_table
KM_table <- KM_table %>%
  mutate(across(c(surv, std_err, lower_ci, upper_ci),
         ~ round(.x, 2))
         )

# Display with gt
KM_table %>%
  gt() %>%
  tab_header(title = "Survival analysis") %>%
    cols_label(
    time      = "Time",
    n_risk    = "At Risk",
    n_event   = "Deaths",
    surv      = "Survival %",
    std_err   = "Std. Error",
    lower_ci  = "Lower CI",
    upper_ci  = "Upper CI"
  ) %>%
  fmt_percent(columns = c(surv, lower_ci, upper_ci), decimals = 1) %>%
  fmt_number(columns = std_err, decimals = 3)
```

## Kaplan-Meier survival curve
```{r read-data, echo = TRUE}
# estimate the survival probabilities for all patients
KM_all_plot <- ggsurvplot(KM_all, 
           data = df8, 
           risk.table = TRUE, 
           linetype = c(1,4),
           tables.height = 0.3,
           pval = TRUE)

KM_all_plot
# *** You may need to zoom out ***
```


### ----------------------------------------------------------


## Do a Kaplan-Meier survival analayis - stratified
```{r km-surv-strat, echo = TRUE}
# estimate the survival probabilities for stroke type
KM_str <- survfit(Surv(time = time2, 
                             event = status == "dead" ) ~ stroke.type, 
                        data = df8)
summary(KM_str)
```

## Survival curve
```{r read-data, echo = TRUE}
# estimate the survival probabilities for all patients
KM_str_plot <- ggsurvplot(KM_str, 
           data = df8, 
           risk.table = TRUE, 
           linetype = c(1,4),
           tables.height = 0.3,
           #conf.int = 0.95,
           pval = TRUE)

KM_str_plot
# *** You may need to zoom out ***
```

## Compare strata
```{r read-data, echo = TRUE}
# Log-rank test
# Test if survival estimates are  different between levels/groups
log_rank_stroke <- survdiff(Surv(time = time2, 
              event = status == "dead") ~ stroke.type, 
         data = df8,
         rho = 0)
log_rank_stroke


log_rank_diabetes <- survdiff(Surv(time = time2, 
              event = status == "dead") ~ dm, 
         data = df8,
         rho = 0)

log_rank_diabetes
```

## Compare strata another way
```{r read-data, echo = TRUE}
# Peto's log-rank test
# Test if survival estimates are different between levels/groups
log_rank_stroke <- survdiff(Surv(time = time2, 
              event = status == "dead") ~ stroke.type, 
         data = df8,
         rho = 1)
log_rank_stroke

# Test if survival estimates are different between levels/groups
log_rank_dm <- survdiff(Surv(time = time2, 
              event = status == "dead") ~ dm, 
         data = df8,
         rho = 1)

log_rank_dm
```





### -------------------------------------------------
## 转专-注爪

## 砖 1
```{r question-1, echo = TRUE}
# 

```

## 砖 2
```{r question-2, echo = TRUE}
# 

```

## 砖 3
```{r question-3, echo = TRUE}
# 

```









### source:  https://bookdown.org/drki_musa/dataanalysis/survival-analysis-kaplan-meier-and-cox-proportional-hazard-ph-regression.html#kaplan-meier-survival-estimates

## Clean the data
```{r clean-data, echo = TRUE}
# Load the data
file <- "https://raw.githubusercontent.com/drkamarul/multivar_data_analysis/refs/heads/main/data/stroke_data.csv"
df8 <- import(file, trust  = TRUE)
#write.csv(df8, here("EpiData", "09 stroke_data.csv"))

# Clean up variable names
df8 <- clean_names(df8)
names(df8) <- gsub("_", ".", names(df8))

# convert the variable doa and variable doa to a more valid format (date format)
df8 <- 
  df8 %>%
  mutate(doa = dmy(doa),
         dod = dmy(dod)
        )

# Convert status to fct
df8 <- df8 %>%
  mutate(status = fct_relevel(status, "alive"))

# Check the updated dataset structure
glimpse(df8)

# Save as R data file
saveRDS(df8, here("EpiData", "09 stroke_data.rds"))
```

## Give variable explanations in labels
```{r labels, echo = TRUE}
attr(df8$doa, "label") <- "转专 砖驻"
attr(df8$dod, "label") <- "转专 砖专专"
attr(df8$status, "label") <- "爪 ( / 驻专)"
attr(df8$sex, "label") <- "专"
attr(df8$gcs, "label") <- "住 转专转 砖 "
attr(df8$dm , "label") <- "住旨旨侄专侄转"
attr(df8$sbp, "label") <- "抓  住住"
attr(df8$dbp, "label") <- "抓  住"
attr(df8$wbc, "label") <- "住驻专转 转  "
attr(df8$time2, "label") <- " 砖专"
attr(df8$stroke.type, "label") <- "住 砖抓  (HS / IS)"
attr(df8$referral.from, "label") <- "驻 -"

labels <- NULL
labels <- data.frame(Variable = names(df8), Label = unlist(var_label(df8)))
gt(labels)
```