###
### name: דוד בורג
### ID  : [מספר תעודת זהות]
### course: אפידמיולוגיה (HIT.ac.il)
### lesson: שרידות 08   Survival
### date  : 25/06/2025
###  
###  נלמד לבצע ניתוח שרידות 
###
### source:  https://bookdown.org/drki_musa/dataanalysis/survival-analysis-kaplan-meier-and-cox-proportional-hazard-ph-regression.html#kaplan-meier-survival-estimates


## Reset memory
```{r setup-packages, include = FALSE}
rm(list = ls(all.names = TRUE))
lapply(paste('package:',names(sessionInfo()$otherPkgs),sep=""),detach,character.only=TRUE,unload=TRUE,force=TRUE)
gc()
dev.off()
```

## Load required packages
```{r load-packages, include = FALSE}
options("install.lock"=FALSE)
if (!require("pacman")) install.packages("pacman")
library(pacman)
pacman::p_load(
  rio,          # File import
  tidyverse,    # data management + ggplot2 graphics, 
  gt,           # Nice beuatiful tables
  nortest,      # Anderson-Darling test for normality
  ggplot2,      # For creating plots
  DescTools,    # For statistical functions like skewness and kurtosis
  janitor,      # adding totals and percents to tables
  car,           # For Levene's test and ANOVA
  gtsummary,    # summary statistics and tests
  corrr,        # correlation analayis for numeric variables
  dplyr,        # For data manipulation
  reshape2,     # For reshaping data
  skimr,        # get overview of data
  performance,           # For model performance metrics
  ResourceSelection,     # For goodness of fit tests
  broom,                 # For tidying model outputs
  pROC,                  # For ROC analysis
  caret,
  MASS,          # For statistical functions (polr for ordinal regression)
  epiDisplay,    # 
  survival,
  survminer
            )
```

## Access the data
```{r read-data, echo = TRUE}
## read the data in a file online
# file <- "./EpiData/stroke_data.csv"     # use this if the file is on your computer
file <- "https://raw.githubusercontent.com/drkamarul/multivar_data_analysis/refs/heads/main/data/stroke_data.csv"
df7 <- import(file, trust  = TRUE)
summary(df7)
glimpse(df7)
```

## Clean the data
```{r read-data, echo = TRUE}
# convert the variable doa and variable doa to a more valid format (date format)
df7 <- 
  df7 %>%
  mutate(doa = dmy(doa),
         dod = dmy(dod)
        )
```

## Show the data
```{r read-data, echo = TRUE}
# in a nice table
df7 %>%
  tbl_summary(
    by = status,
    include = -c(doa, dod)  # exclude these two date variables
  ) %>%
  as_gt()


df7 %>%
  tbl_summary(
    by = stroke_type,
    include = -c(doa, dod)  # exclude these two date variables
  ) %>%
  as_gt()
```


## Survival time statistics
```{r surv-stats, echo = TRUE}
# Survival time
mean_survival_time <- mean(df7$time2, na.rm = TRUE)
sd_survival_time <- sd(df7$time2, na.rm = TRUE)
CI1 <- MeanCI(df7$time2, na.rm = TRUE)
  lower_meanci_1 <- as.numeric((CI1["lwr.ci"]))
  upper_meanci_1 <- as.numeric((CI1["upr.ci"]))
median_survival_time <- median(df7$time2, na.rm = TRUE)
IQR_survival_time <- IQR(df7$time2, na.rm = TRUE)
q25_survival_time <- as.numeric(Quantile(df7$time2, probs = 0.25, na.rm = TRUE))
q75_survival_time <- as.numeric(Quantile(df7$time2, probs = 0.75, na.rm = TRUE))
min_survival_time <- min(df7$time2, na.rm = TRUE)
max_survival_time <- max(df7$time2, na.rm = TRUE)

# Create a data frame to neatly display the statistics
stats_df <- data.frame(
    Statistic = c("Mean", "SD", "⠀⠀CI₉₅ low", "⠀⠀CI₉₅ high", "Median", "⠀⠀Q1", "⠀⠀Q3", "⠀⠀IQR", "Minimum", "Maximum")
  , Value = c(mean_survival_time, sd_survival_time, lower_meanci_1, upper_meanci_1, median_survival_time, q25_survival_time, q75_survival_time, IQR_survival_time, min_survival_time, max_survival_time)
  , סטטיסטי = c("ממוצע", "ס'ת", "⠀⠀רווח סמך", "⠀⠀רווח סמך", "חציון", "⠀⠀רבעון 1", "⠀⠀רבעון 3", "⠀⠀מרווח בין רבועני", "מינימום", "מקסימום")
)

# Clean, simple table
summary_stats <- stats_df %>% 
  gt() %>%
  tab_header(title = "Survival Statistics⠀⠀⠀⠀זמן הישרדות") %>%
  fmt_number(columns = Value, decimals = 1)
summary_stats              # Show the table
```


## Do a Kaplan-Meyer survival analayis
```{r km-surv, echo = TRUE}
# estimate the survival probabilities for all patients
KM_all <- survfit(Surv(time = time2, 
                       event = status == "dead" ) ~ 1, 
                       data = df7)
summary(KM_all)
```


## Life table
```{r life-table, echo = TRUE}
# Convert status to int
df7 <- df7 %>%
  mutate(event = ifelse(status == "dead", 1, 0))

# Fit Kaplan-Meier model stratified by stroke_type (example)
km_fit <- survfit(Surv(time2, event) ~ 1, data = df7)

# Extract survival summary at intervals of 10 units
km_summary <- summary(km_fit, times = seq(0, max(df7$time2), by = 10))

# Convert to data frame
km_df <- data.frame(
  time = km_summary$time,
  #strata = km_summary$strata,
  n_risk = km_summary$n.risk,
  n_event = km_summary$n.event,
  surv = km_summary$surv,
  std_err = km_summary$std.err,
  lower_ci = km_summary$lower,
  upper_ci = km_summary$upper
)

# Display with gt
km_df %>%
  gt() %>%
  tab_header(title = "Survival analysis") %>%
    cols_label(
    time      = "Time",
    n_risk    = "At Risk",
    n_event   = "Deaths",
    surv      = "Survival %",
    std_err   = "Std. Error",
    lower_ci  = "Lower CI",
    upper_ci  = "Upper CI"
  ) %>%
  fmt_percent(columns = c(surv, lower_ci, upper_ci), decimals = 1) %>%
  fmt_number(columns = std_err, decimals = 3) 
  
```

## Kaplan-Meyer survival analayis plot
```{r read-data, echo = TRUE}
# estimate the survival probabilities for all patients
KM_all_plot <- ggsurvplot(KM_all, 
           data = df7, 
           risk.table = TRUE, 
           linetype = c(1,4),
           tables.height = 0.3,
           pval = TRUE)

KM_all_plot
# *** You may need to zoom out ***
```





### ----------------------------------------------------------





## Do a Kaplan-Meyer survival analayis - stratified
```{r km-surv-strat, echo = TRUE}
# estimate the survival probabilities for stroke type
KM_str_type <- survfit(Surv(time = time2, 
                             event = status == "dead" ) ~ stroke_type, 
                        data = df7)
summary(KM_str_type)
```

## Survival curve
```{r read-data, echo = TRUE}
# estimate the survival probabilities for all patients
KM_str_plot <- ggsurvplot(KM_str_type, 
           data = df7, 
           risk.table = TRUE, 
           linetype = c(1,4),
           tables.height = 0.3,
           pval = TRUE)

KM_str_plot
# *** You may need to zoom out ***
```

## Life table
```{r life-table, echo = TRUE}
# Convert status to int
# Fit Kaplan-Meier model stratified by stroke_type (example)
km_fit_strat <- survfit(Surv(time2, event) ~ stroke_type, data = df7)

# Extract survival summary at intervals of 10 units
km_summary_strat <- summary(km_fit_strat, times = seq(0, max(df7$time2), by = 10))

# Convert to data frame
km_df <- data.frame(
  time = km_summary_strat$time,
  strata = km_summary_strat$strata,
  n_risk = km_summary_strat$n.risk,
  n_event = km_summary_strat$n.event,
  surv = km_summary_strat$surv,
  std_err = km_summary_strat$std.err,
  lower_ci = km_summary_strat$lower,
  upper_ci = km_summary_strat$upper
)

# Display with gt
km_df %>%
  gt() %>%
  tab_header(title = "Survival analysis") %>%
    cols_label(
    time      = "Time",
    strata    = "Group",
    n_risk    = "At Risk",
    n_event   = "Deaths",
    surv      = "Survival %",
    std_err   = "Std. Error",
    lower_ci  = "Lower CI",
    upper_ci  = "Upper CI"
  ) %>%
  fmt_percent(columns = c(surv, lower_ci, upper_ci), decimals = 1) %>%
  fmt_number(columns = std_err, decimals = 3) 
```


### ----------------------------------
### Compare strata

## Log-rank test
```{r read-data, echo = TRUE}
# Test if survival estimates are  different between levels/groups
log_rank_stroke <- survdiff(Surv(time = time2, 
              event = status == "dead") ~ stroke_type, 
         data = df7,
         rho = 0)

log_rank_stroke

# Test if survival estimates are  different between levels/groups
log_rank_diabetes <- survdiff(Surv(time = time2, 
              event = status == "dead") ~ dm, 
         data = df7,
         rho = 0)

log_rank_diabetes
```

## Peto's log-rank test
```{r read-data, echo = TRUE}
# Test if survival estimates are  different between levels/groups
log_rank_stroke <- survdiff(Surv(time = time2, 
              event = status == "dead") ~ stroke_type, 
         data = df7,
         rho = 1)

log_rank_stroke

# Test if survival estimates are  different between levels/groups
log_rank_dm <- survdiff(Surv(time = time2, 
              event = status == "dead") ~ dm, 
         data = df7,
         rho = 1)

log_rank_dm
```









## Q1
```{r setup, include=FALSE}
data<-birth_survivalR
```

## Q1
```{r until 15, echo=FALSE}
install.packages("survival")
install.packages("survminer")
install.packages("broom")

library(broom)
library(survival)
library(survminer)
library(dplyr)
```

## Q2
```{r until 15, echo=FALSE}
##question 2
# Creating the Surv object
surv_obj_1 <- Surv(time = data$time_to_anemia, event = data$anemia1)

# survival model
fit1<- survfit(surv_obj_1 ~ 1)

# Extract summary from the fit for intervals up to 15 days
fit_summary1 <- summary(fit1, times = seq(0, 15, by = 1))

# Create  life table
life_table <- data.frame(
  time = fit_summary1$time,
  n_risk = fit_summary1$n.risk,
  n_event = fit_summary1$n.event,
  n_censor = fit_summary1$n.censor
)



# Calculate the number exposed to risk accounting for half the censored individuals
life_table <- life_table %>%
  mutate(exposed_to_risk = n_risk - 0.5 * n_censor)

# Calculate the survival proportion at each interval
life_table <- life_table %>%
  mutate(survival_proportion = 1 - (n_event / exposed_to_risk))

# Calculate the cumulative survival proportion
life_table <- life_table %>%
  mutate(cumulative_survival = cumprod(survival_proportion))

# Print the life table
print(life_table)
```

```{r cars}
###question 3


#  Surv object
surv_obj <- Surv(time = data$time_to_anemia, event = data$anemia1)

# Fit the survival model
fit <- survfit(surv_obj ~ 1)
#for maximum time
fit_summary <- summary(fit, times = seq(0, max(data$time_to_anemia, na.rm = TRUE), by = 1))

# Create the life table
life_table <- data.frame(
  time = fit_summary$time,
  n_risk = fit_summary$n.risk,
  n_event = fit_summary$n.event,
  n_censor = fit_summary$n.censor
)

# Calculate the number exposed to risk accounting for half the censored individuals
life_table <- life_table %>%
  mutate(exposed_to_risk = n_risk - 0.5 * n_censor)

# Calculate the survival proportion at each interval
life_table <- life_table %>%
  mutate(survival_proportion = 1 - (n_event / exposed_to_risk))

# Calculate the cumulative survival proportion
life_table <- life_table %>%
  mutate(cumulative_survival = cumprod(survival_proportion))

# Print the life table
print(life_table)


# Create the survival plot
ggsurvplot(fit, data = data, conf.int = TRUE,
           ggtheme = theme_minimal(),
           risk.table = FALSE, # Add risk table
           pval = FALSE)       # Add p-value of the log-rank test


```

## Q4
```{r cars}
##question 4

surv_obj <- Surv(time = data$time_to_anemia, event = data$anemia1)

# Fit the survival model


# Fit the survival model stratified by race
fit <- survfit(surv_obj ~ race, data = data)

# Extract the summary from the fit with intervals of one unit for each race group
fit_summary <- summary(fit, times = seq(0, max(data$time_to_anemia, na.rm = TRUE), by = 1))

# Initialize an empty data frame for the life table
life_table <- data.frame()

# Loop through each race group to create life tables
for (i in seq_along(fit$strata)) {
  race_group <- names(fit$strata)[i]
  race_summary <- summary(fit[i], times = seq(0, max(data$time_to_anemia, na.rm = TRUE), by = 1))
  
  race_life_table <- data.frame(
    race = race_group,
    time = race_summary$time,
    n_risk = race_summary$n.risk,
    n_event = race_summary$n.event,
    n_censor = race_summary$n.censor
  )
  
  # Calculate the number exposed to risk accounting for half the censored individuals
  race_life_table <- race_life_table %>%
    mutate(exposed_to_risk = n_risk - 0.5 * n_censor)
  
  # Calculate the survival proportion at each interval
  race_life_table <- race_life_table %>%
    mutate(survival_proportion = 1 - (n_event / exposed_to_risk))
  
  # Calculate the cumulative survival proportion
  race_life_table <- race_life_table %>%
    mutate(cumulative_survival = cumprod(survival_proportion))
  
  # Append to the main life table data frame
  life_table <- rbind(life_table, race_life_table)
}


# Print the life table
print(life_table)
```

```{r cars}

#KM


# Plot Kaplan-Meier survival curves
ggsurvplot(
  fit,
  data = data,
  conf.int = TRUE,
  pval = TRUE,
  risk.table = FALSE,
  ggtheme = theme_minimal(),
  title = "Kaplan-Meier Survival Curves by Race Group",
  xlab = "Time to Anemia",
  ylab = "Survival Probability"
)



 print(fit,print.rmean = T)

 
 
 # Perform the log-rank test not mandatory 
log_rank_test <- survdiff(surv_obj ~ race, data = data)

# Print the results of the log-rank test
print(log_rank_test)

# Calculate the p-value
p_value <- 1 - pchisq(log_rank_test$chisq, length(log_rank_test$n) - 1)
p_value







###KAPLAN survival table


km_table <- data.frame()

for (i in seq_along(fit$strata)) {
  race_group <- names(fit$strata)[i]
  # TRUE KM: No "times = ..." so we get real event time points
  km_summary <- summary(fit[i])

  stratum_df <- data.frame(
    race = race_group,
    time = km_summary$time,
    n_risk = km_summary$n.risk,
    n_event = km_summary$n.event,
    n_censor = km_summary$n.censor,
    survival = km_summary$surv,
    std_err = km_summary$std.err,
    lower_CI = km_summary$lower,
    upper_CI = km_summary$upper
  )

  km_table <- rbind(km_table, stratum_df)
}

print(km_table)


```


