###
### name  :  专
### ID    : [住驻专 转注转 转]
### course: 驻 (HIT.ac.il)
### lesson: 09 专专住转 Cox regression
### date  : 13/07/2025
###  
###  Cox  爪注 转 砖专转 爪注转 专专住转 
###
### source:  https://bookdown.org/drki_musa/dataanalysis/survival-analysis-kaplan-meier-and-cox-proportional-hazard-ph-regression.html#semi-parametric-models-in-survival-analysis


## Reset memory 
```{r setup-packages, include = FALSE}
rm(list = ls(all.names = TRUE))       #  Remove all data objects         Remove packages
lapply(paste("package:", names(sessionInfo()$otherPkgs), sep = ""), detach, character.only = TRUE, unload = TRUE, force = TRUE)           # Remove packages from memory
gc()                                  # Remove unused data from memory ("garbage cllection")
dev.off()                             # Reset the display output of R
q("no")                               # This will completely shutdown R terminal
```

## Load required packages
```{r load-packages, include = FALSE}
options("install.lock"=FALSE)
if (!require("pacman")) install.packages("pacman")
library(pacman)
pacman::p_load(
  Rcpp, devtools, gh, remotes,   # For installing packages/files online (like from GitHub)
  rio, here,              # File import and path support
  tidyverse,              # data management + ggplot2 graphics,
  broom,                  # For tidying model outputs
  dplyr,                  # For data manipulation
  gt,                     # Nice beuatiful tables
  nortest,                # Anderson-Darling test for normality
  ggplot2,                # For creating plots
  DescTools,              # For statistical functions like skewness and kurtosis
  psych,                  # to get a table of all summary statistics
  janitor,                # adding totals and percents to tables
  car,                    # For Levene's test and ANOVA
  gtsummary,              # summary statistics and tests in beautiful tables
  corrr,                  # correlation
  lmtest,                 # linear regression tets
  parameters,             # calculate standardized parameters/coefficients
  GGally,                 # to make a correlation matrix
  skimr,                  # get overview of data
  pROC,                   # For ROC analysis
  performance,            # For model performance metrics
  #MASS,                   # For statistical functions (polr for ordinal regression)ordinal
  ordinal,                # For ordinal regression statistical functions
  caret,                  # For confusion matrix and classification metrics
  ggeffects,              # To display the confusion matrix
  generalhoslem,          # For Hosmer-Lemeshow test for goodness of fit
  ResourceSelection,      # For Hosmer-Lemeshow test for goodness of fit
  VGAM,                   # for multinomial logistic regression
  epiDisplay,             # For epi data type handling
  survival,               # For survival analysis
  survminer,              # For survival analysis and visualization
  muhaz                   #
            )
# Install/load flexplot 
if (!require("flexplot", quietly = TRUE)) devtools::install_github("dustinfife/flexplot", ref = "logistic_plots")  ;  library(flexplot)
```

## Access the data
```{r read-data, echo = TRUE}
# file <- "./EpiData/stroke_data.csv"     #  use this if the file is on your computer
file <- "https://raw.githubusercontent.com/drkamarul/multivar_data_analysis/refs/heads/main/data/stroke_data.csv"
      ## read the data in a file online
df7 <- import(file, trust  = TRUE)
summary(df7)
glimpse(df7)
```

## Clean the data
```{r read-data, echo = TRUE}
# Clean up variable names
df7 <- clean_names(df7)
names(df7) <- gsub("_", ".", names(df7))

# convert the variable doa and variable doa to a more valid format (date format)
df7 <- 
  df7 %>%
  mutate(doa = dmy(doa),
         dod = dmy(dod))

df7 <- df7 %>%
  mutate(status = fct_relevel(status, "alive"))
```

## Do a simple Cox regession
```{r cox-reg, echo = TRUE}
# Cox regression
cox_reg <- 
  coxph(Surv(time = time2, 
             event = status == 'dead') ~ stroke.type,
                     data = df7)
```

## Calculate Hazard ratios and display
```{r hz, echo = TRUE}
# Calculate log Hazard using gtsummary
log_hz <- cox_reg %>%
  tbl_regression(exp = FALSE)  # exp = FALSE shows log hazard ratios

# Calculate Crude Hazard using gtsummary
crude_hz <- cox_reg %>%
  tbl_regression(exp = TRUE)  # exp = TRUE shows hazard ratios

# Display results
hz_combined_tbl <- tbl_merge(
  tbls = list(log_hz, crude_hz),
  tab_spanner = c("**log Hazard**", "**Crude Hazard**")
)  

#%>% 
#  as_gt() %>%
#  tab_source_note(
#    source_note = paste0("R虏 = ", round(r2_valm,3),
#                         ",  -2LL = ", round(minus_2LL_mlog_reg, 0), 
#                         ",  AIC = ", round(AIC(mlog_reg), 0)
#                        )
#  )

hz_combined_tbl
```

## Cox hazard plot
```{r cox-plot, echo = TRUE}
# Fit the survival object
surv_obj <- with(df7, Surv(time2, stroke.type == "dead"))

# Estimate hazard function
hz <- muhaz(df7$time2, df7$status == "dead")

# Plot
plot(hz, main = "Smoothed Hazard Function", xlab = "Time", ylab = "Hazard")
```


### -----------------------------------------



### Cox multi-regression

## Do a multi-variate Cox regression
```{r read-data, echo = TRUE}
# Perform the Cos regression on multple variables
cox_mreg <-
  coxph(Surv(time = time2, 
             event = status == 'dead') ~ stroke.type +  gcs + referral.from, 
        data = df7)

# get results for Crude Hazard
cox_cr_table <- tbl_regression(cox_mreg)

# get for Hazard Ration (HR)
cox_hr_tbl <- tbl_regression(cox_mreg, exponentiate = TRUE)
# Combine them side by side
combined_tbl <- tbl_merge(
  tbls = list(cox_cr_table, cox_hr_tbl),
  tab_spanner = c("**Log Hazard**", "**Hazard Ratio**")
)  
combined_tbl
```

## Test with respect to stroke.type and refereal_form
```{r read-data, echo = TRUE}
# Perform the Cos regression on multple variables
cox_mreg_test <-
  coxph(Surv(time = time2, 
             event = status == 'dead') ~ stroke.type + referral.from , 
        data = df7)

# Do ANOVA
anova(cox_mreg_test, cox_mreg, test = 'Chisq')
```

### Model Perforamnce
```{r read-data, echo = TRUE}
# -2LL
minus_2LL_cox <- -2 * as.numeric(logLik(cox_mreg))
# AIC
AIC_cox <- extractAIC(cox_mreg)


# Assumptions - Schoenfeld_test
#             - test Ho: constant regression coefficients
#             - p>0.05 = risks are proportional over time
schoenfeld_test <- cox.zph(cox_mreg)
schoenfeld_test

plot(schoenfeld_test)
```

## Do a multi-variate Cox regression with interaction term
```{r read-data, echo = TRUE}
### -----------------------------------------
# Perform the Cos regression on multple variables
cox_mreg_ia <-
  coxph(Surv(time = time2, 
             event = status == 'dead') ~ gcs + stroke.type + stroke.type:gcs , 
        data = df7)

cox_mreg_ia_tbl <- tbl_regression(cox_mreg_ia)
cox_mreg_ia_tbl
```




### -------------------------------------------------
## 转专-注爪

## 砖 1
```{r question-1, echo = TRUE}
# 

```

## 砖 2
```{r question-2, echo = TRUE}
# 

```

## 砖 3
```{r question-3, echo = TRUE}
# 

```