###
### name  : [ שם משפחה + שם פרטי ]
### ID    : [ מספר תעודת זהות ]
### course: אפידמיולוגיה (HIT.ac.il)
### lesson: שיעורי בית 2
### date  : [ תאריך ]
###  
###  רגרסיה לינראית
###
### source: https://gtpb.github.io/PSLS20/pages/04-dataExploration/04-Data-exploration-FEV.html

# Reset memory
```{r setup-packages, include false}
rm(list = ls(all.names = TRUE))
lapply(paste('package:',names(sessionInfo()$otherPkgs),sep=""),detach,character.only=TRUE,unload=TRUE,force=TRUE)
gc()
```

# Load required packages
```{r setup-packages, include false}
install.packages("pacman")
library(pacman)
pacman::p_load(
  rio,          # File import
  here,         # File locator
  skimr,        # get overview of data
  tidyverse,    # data management + ggplot2 graphics, 
  gtsummary,    # summary statistics and tests
  rstatix,      # statistics
  nortest,      # Anderson-Darling test for normality
  moments,      # skewness and kurtosis
  corrr,        # correlation analayis for numeric variables
  janitor,      # adding totals and percents to tables
  flextable,    # converting tables to HTML
  dplyr,        # For data manipulation
  ggplot2,      # For creating plots
  DescTools,    # For statistical functions like skewness and kurtosis
  car,          # For Levene's test and ANOVA
  vctrs        # For data type handling
            )

## library(kableExtra)
#library(tibble)
#ibrary(readxl)
## library(lubridate)
## library(haven)
## library(gtsummary)
#ibrary(ggplot2)
## library(ggpubr)
## library(GGally)
## library(rsq)
## library(broom)
## library(here)
## library(patchwork)
## library(plotly)
```

# Get data
```{r load data, include false}
#file <- "HW1_fev.txt"     # use this if the file is on your computer
file <- "https://raw.githubusercontent.com/GTPB/PSLS20/master/data/fev.txt"
df_fev <- import(file, trust  = TRUE)
```

# Check out the variables
```{r load data, include false}
print(df_fev, n=20)
glimpse(df_fev)
summary(df_fev)
```

# summary statistics for numerical variables
```{r sort_data, include false}
# Show median and IQR for all numerical variables (5)
df_fev %>% 
  select(age, fev, height) %>% 
  tbl_summary()
```

# summary statistics for cat variables
```{r sort_data, include false}
# Show percent for all cat variables (5)
df_fev %>% 
  select(gender, smoking) %>% 
  tbl_summary()
```


# Display boxplot
```{r select_data, include false}
# filter() will select males 7+ years old and send
df_fev %>%
  ggplot(aes(x=smoking,y=fev,fill=smoking)) +
  #scale_fill_manual(values=c("dimgrey","firebrick")) +
  theme_bw() +
  geom_boxplot(outlier.shape=NA) + 
  geom_jitter(width = 0.2, size=0.1) +
  ggtitle("Boxplot of FEV versus smoking") +
  ylab("fev (l)") +
  xlab("smoking status")
```



# Display pairs
```{r select_data, include false}
# filter() will select males 7+ years old and send
pairs_plot <- df_fev %>% 
  select(-gender, -smoking) %>% 
  ggpairs()

pairs_plot
ggplotly(pairs_plot)
```



# Multi-variable regression with lm
```{r num-to-cat, include false}
# Do a simple linear regression (SLR) with chol as the dependent variable 
mlr_fev <- lm(fev ~ age + fev + height + gender + smoking, data = df_fev)
summary(mlr_fev)

tidy(mlr_fev) %>%
  mutate(p.value = case_when(
    p.value < 0.001 ~ "<0.001",
    p.value < 0.01  ~ "<0.01",
    p.value < 0.05  ~ "<0.05",
    TRUE            ~ "≥0.05"
  )) %>%
  kable("html", caption = "Muliple Linear Regression (p-values)") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    position = "left"
  )
# Assume dbp and race are significant

```



# Multi-variable regression with lm
```{r num-to-cat, include false}
# Do a simple linear regression (SLR) with chol as the dependent variable 
mlr_fev <- lm(fev ~ age + fev + height + gender + smoking, data = df_fev)
summary(mlr_fev)

tidy(mlr_fev) %>%
  mutate(p.value = case_when(
    p.value < 0.001 ~ "<0.001",
    p.value < 0.01  ~ "<0.01",
    p.value < 0.05  ~ "<0.05",
    TRUE            ~ "≥0.05"
  )) %>%
  kable("html", caption = "Muliple Linear Regression (p-values)") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    position = "left"
  )
# Assume dbp and race are significant

```




# Run a stepwise analysis
```{r num-to-cat, include false}
# Do a multiple regression (MLR) with chol as the dependent variable 
# both
mlr_fev_stepboth <- step(mlr_fev, direction = "both")

mlr_fev_sel <- lm(age + fev + height + gender + smoking, data = df_fev)
summary(mlr_fev)

model_table_stats <- tbl_regression(mlr_fev_sel,
                             #label = list(dbp = "FEV"),
                             conf.int = TRUE,
                             show_single_row = "dbp",
                             intercept = TRUE) %>%
  modify_header(label = "**coeff**",
                estimate = "**beta**",
                std.error = "**SE**",
                statistic = "**F**",
                p.value = "**p**",
                conf.low = "**CI<sub>95%**") %>%
  modify_table_body(~ .x %>% filter(label != "Abbreviations")) %>%
  add_glance_source_note(
    include = c(r.squared, adj.r.squared, statistic, p.value, logLik, AIC, sigma),
    label = list(
      r.squared = "R²",
      adj.r.squared = "Adj.R²",
      statistic = "F",
      p.value = "p",
      logLik = "log-L",
      sigma = "\U03C3",
      AIC = "AIC"
    )
  ) %>%
  modify_caption("**Table 2: Model Coefficients<br>(fit statistics)**")

model_table_stats
```