###
### name  : [ 砖 砖驻 + 砖 驻专]
### ID    : [ 住驻专 转注转 转 ]
### course: 驻 (HIT.ac.il)
### lesson: 砖注专 转 2
### date  : [ 转专 ]
###   
###  专专住 住转
###
### https://github.com/ThecoderPinar/Diabetes_Health_Prediction_and_Analysis


## Reset memory
```{r setup-packages, include false}
rm(list = ls(all.names = TRUE))
lapply(paste('package:',names(sessionInfo()$otherPkgs),sep=""),detach,character.only=TRUE,unload=TRUE,force=TRUE)
gc()
```

## Load required packages
```{r load-packages, include false}
options("install.lock"=FALSE)
if (!require("pacman")) install.packages("pacman")
library(pacman)
pacman::p_load(
  rio,          # File import
  tidyverse,    # data management + ggplot2 graphics, 
  broom,        # For tidying model outputs
  dplyr,        # For data manipulation
  gt,           # Nice beuatiful tables
  dplyr,        # For data manipulation
  nortest,      # Anderson-Darling test for normality
  ggplot2,      # For creating plots
  DescTools,    # For statistical functions like skewness and kurtosis
  janitor,      # adding totals and percents to tables
  car,           # For Levene's test and ANOVA
  gtsummary,    # summary statistics and tests
  corrr,        # correlation analayis for numeric variables
  reshape2,     # For reshaping data
  skimr,        # get overview of data
  performance,           # For model performance metrics
  ResourceSelection,     # For goodness of fit tests
  pROC,                  # For ROC analysis
  caret                  # For confusion matrix and classification metrics
            )
```

## Access the data
```{r read-data, echo=FALSE}
# file <- "./EpiData/HWdata_2.csv       #  use this if the file is on your computer
file <- "https://raw.githubusercontent.com/ThecoderPinar/Diabetes_Health_Prediction_and_Analysis/refs/heads/main/raw/diabetes_data.csv"
df <- import(file)
summary(df)
glimpse(df)  # get an overview of the data
```

## Show the data
```{r fix-data, echo=FALSE}
# Need to change status to be 1,0 
df %>%
  tbl_summary(by = Diagnosis) %>%
  as_gt()
```

## Multiple Logistic regression
```{r logistic_regression, echo=FALSE}
# Perform logistic regression
mlog_reg <- glm(Diagnosis ~ Smoking + AlcoholConsumption + AntihypertensiveMedications + Statins + SerumCreatinine + HbA1c + FastingBloodSugar + FamilyHistoryDiabetes , data = df, family = binomial(link = 'logit'))
tidy(mlog_reg, conf.int = TRUE)
```

## Model performance measures
```{r model-performance, echo = TRUE}
# Get -2LL and AIC
minus_2LL_mlog_reg <- -2 * as.numeric(logLik(mlog_reg))
AIC_val <- AIC(mlog_reg)

# Calculate Nagelkerke's R虏
r2_val <- PseudoR2(mlog_reg, which ="Nagelkerke")

# Chi-square statistic for the omnibus test
null_model <- glm(Diagnosis ~ 1, data = df, family = binomial(link = 'logit')) # null model (intercept-only model)
      model_perf <- anova(null_model, mlog_reg, test = "Chisq")
      chi_sq <- model_perf$Deviance[2]     # Chi-square statistic
      p_val <- model_perf$`Pr(>Chi)`[2]    # p-value

# Disiplay the summary of the logistic regression model
summary(mlog_reg)
paste("R虏= ",r2_val)
```

# Calculate log OR
```{r calc-log_OR, echo = TRUE}
mlog_reg_tbl <- tbl_regression(mlog_reg, exponentiate = FALSE)
mlog_reg_tbl
```

# Calculate OR
```{r calc-OR, include=FALSE}
mlog_reg_OR <- tbl_regression(mlog_reg, exponentiate = TRUE)
mlog_reg_OR
```

# Combine logOR and OR results with model performance 
```{r log_reg_tbl, echo = TRUE}
combined_tbl <- tbl_merge(
  tbls = list(mlog_reg_tbl, mlog_reg_OR),
  tab_spanner = c("**log Odds**", "**Odds Ratio**")  # Optional: Add spanning headers
) %>%
  as_gt() %>%
  tab_source_note(
    source_note = paste0("R虏 = ", round(r2_val,3),
                          ", -2LL = ", round(minus_2LL_mlog_reg, 0), 
                         ",  AIC = ", round(AIC_val, 0),
                         ",  虏 = ", round(chi_sq, 0),
                         ",  p = ", round(p_val, 3)
                        )
  )

# Display the combined table
combined_tbl
```


## Add inreaction term
```{r OR, include=FALSE}
# Recalculate logistic regression for OR
mlog_reg_ia <- glm(Diagnosis ~ Smoking + HbA1c + FastingBloodSugar + HbA1c:FastingBloodSugar, data = df, family = binomial(link = 'logit'))
mlog_reg_ia
tidy(mlog_reg_ia)

mlog_reg_ia_table <- tbl_regression(
  mlog_reg_ia,
  exponentiate = TRUE,     # show odds ratios instead of log-odds
)

mlog_reg_ia_table
```

## Logistic multi-regression prediction
```{r log_reg-pred, echo = TRUE}
# Create a new data frame for prediction
predict_values <- data.frame(df = c(0, 1))

# Calculate predicted probabilities
predicted_probs <- predict(mlog_reg, type = "response")
df$predicted_prob_multi <- predict(mlog_reg, type = "response")
predict_prob
df$predicted_prob_multi <- predict(mlog_reg, type = "response")

glimpse(df)
```

## ROC curve and AUC (area under curve)
```{r mlog_reg-ROC, echo = TRUE}
roc_obj <- roc(df$Diagnosis, df$predicted_prob_multi)

# Plot ROC curve
plot(roc_obj, col = "blue", main = "ROC Curve")

# Print AUC (AUC > 0.7)
auc(roc_obj)
```


## classification table
```{r , echo=FALSE}
# Use caret package to get classification table
confusion_matrix <- caret::confusionMatrix(data = as.factor(round(df$predicted_prob_multi,0)), reference = as.factor(df$Diagnosis))
confusion_matrix
```

## Hosmer-Lemeshow test for GOF
```{r , echo=FALSE}
# Use ResourceSelection to get the HosmerLemeshow test result
hl_test <- hoslem.test(df$Diagnosis, df$predicted_prob_multi, g = 10)  
hl_test
```