###
### name  : [ 砖 砖驻 + 砖 驻专]
### ID    : [ 住驻专 转注转 转 ]
### course: 驻 (HIT.ac.il)
### lesson: 注 砖 2
### date  : [ 转专 ]
###   
###  专专住 住转


## Reset memory
```{r setup-packages, include = FALSE}
rm(list = ls(all.names = TRUE))       #  Clear all data objects         Clear packages
lapply(paste("package:", names(sessionInfo()$otherPkgs), sep = ""), detach, character.only = TRUE, unload = TRUE, force = TRUE)           # Remove packages from memory
gc()                                  # Clear unused data from memory ("garbage cllection")
dev.off()                             # Reset the display output of R
q("no")                               # This will completely shutdown R terminal
```

## Load required packages
```{r load-packages, include false}
options("install.lock"=FALSE)
if (!require("pacman")) install.packages("pacman")
library(pacman)
pacman::p_load(
  Rcpp, devtools, gh, remotes,   # For installing packages/files online (like from GitHub)
  rio, here,              # File import and path support
  tidyverse,              # data management + ggplot2 graphics,
  broom,                  # For tidying model outputs
  dplyr,                  # For data manipulation
  gt,                     # Nice beuatiful tables
  gtsummary,              # summary statistics and tests in beautiful tables
  nortest,                # Anderson-Darling test for normality
  ggplot2,                # For creating plots
  DescTools,              # For statistical functions like skewness and kurtosis
  janitor,                # adding totals and percents to tables
  psych,                  # to get a table of all summary statistics
  car,                    # For Levene's test and ANOVA
  gtsummary,              # summary statistics and tests in beautiful tables
  corrr,                  # correlation
  lmtest,                 # linear regression tets
  parameters,             # calculate standardized parameters/coefficients
  GGally,                 # to make a correlation matrix
  skimr,                  # get overview of data
  pROC,                   # For ROC analysis
  performance,            # For model performance metrics
  caret,                  # For confusion matrix and classification metrics
  ggeffects,              # To display the confusion matrix
  ResourceSelection       # For Hosmer-Lemeshow test for goodness of fit 
            )
# Install/load flexplot
if (!require("devtools")) devtools::install_github("dustinfife/flexplot", ref = "development")  ;  if (!require("pacman")) install.packages("pacman")  ;  library(flexplot)
```

## Access the data
```{r read-data, echo=FALSE}
# file <- "./EpiData/HWdata_2.rds               #  use this if the file is on your computer
file <- "https://raw.githubusercontent.com/Model-Lab-Net/Courses/refs/heads/main/Epi/Data/HWdata_2.rds"
df <- import(file, trust  = TRUE)
# show variable translations to Hebrew in lables
labels <- data.frame(Variable = names(df), Label = unlist(var_label(df)))   ;   gt(labels)
glimpse(df)               # get an overview of the data
```

## Get summary statistics
```{r read-data, echo=FALSE}
# Need to change status to be 1,0 
df %>%
  tbl_summary(by = diagnosis) %>%
  as_gt()
```

## Multiple Logistic regression
```{r logistic_regression, echo=FALSE}
# Perform logistic regression
mlog_reg <- glm(diagnosis ~ smoking + alcohol.consumption + antihypertensive.medications + 
              statins + serum.creatinine + hb.a1c + fasting.blood.sugar + 
              family.history.diabetes , data = df, family = binomial(link = 'logit'))
tidy(mlog_reg, conf.int = TRUE)
summary(mlog_reg)
```

## Model performance measures
```{r model-performance, echo = TRUE}
# Get -2LL and AIC
minus_2LL_mlog_reg <- -2 * as.numeric(logLik(mlog_reg))
AIC_val <- AIC(mlog_reg)

# Calculate Nagelkerke's R虏
r2_val <- PseudoR2(mlog_reg, which ="Nagelkerke")

# Chi-square statistic for the omnibus test
null_model <- glm(diagnosis ~ 1, data = df, family = binomial(link = 'logit')) # null model (intercept-only model)
      model_perf <- anova(null_model, mlog_reg, test = "Chisq")
      chi_sq <- model_perf$Deviance[2]     # Chi-square statistic
      p_val <- model_perf$`Pr(>Chi)`[2]    # p-value

# Disiplay the summary of the logistic regression model
summary(mlog_reg)
paste("R虏= ",r2_val)
```

## Calculate log OR
```{r calc-log_OR, echo = TRUE}
mlog_reg_tbl <- tbl_regression(mlog_reg, exponentiate = FALSE)
mlog_reg_tbl
```

## Calculate OR
```{r calc-OR, include=FALSE}
mlog_reg_OR <- tbl_regression(mlog_reg, exponentiate = TRUE)
mlog_reg_OR
```

## Combine logOR and OR results with model performance 
```{r log_reg_tbl, echo = TRUE}
combined_tbl <- tbl_merge(
  tbls = list(mlog_reg_tbl, mlog_reg_OR),
  tab_spanner = c("**log Odds**", "**Odds Ratio**")  # Optional: Add spanning headers
) %>%
  as_gt() %>%
  tab_source_note(
    source_note = paste0("R虏 = ", round(r2_val,3),
                          ", -2LL = ", round(minus_2LL_mlog_reg, 0), 
                         ",  AIC = ", round(AIC_val, 0),
                         ",  虏 = ", round(chi_sq, 0),
                         ",  p = ", round(p_val, 3)
                        )
  )

# Display the combined table
combined_tbl
```


## Use interaction term on significant variables
```{r OR, include=FALSE}
# Recalculate logistic regression for OR
mlog_reg_ia <- glm(diagnosis ~ hb.a1c*fasting.blood.sugar,
                   data = df, family = binomial(link = 'logit'))
mlog_reg_ia
tidy(mlog_reg_ia)

mlog_reg_ia_table <- tbl_regression(
  mlog_reg_ia,
  exponentiate = TRUE,     # show odds ratios instead of log-odds
)

mlog_reg_ia_table
```


## Model comparison
```{r mdodel-comparison, echo = TRUE}
model.comparison(mlog_reg, mlog_reg_ia)
```


## Classification table
```{r log_reg-pred, echo = TRUE}
# Create a new data frame for prediction
predict.values <- data.frame(df = c(0, 1))

# Calculate predicted probabilities
df$predict.prob <- predict(mlog_reg_ia, type = "response")
glimpse(df)

# Use caret package to get classification table
confusion_matrix <- caret::confusionMatrix(data = as.factor(round(df$predict.prob,0)), reference = as.factor(df$diagnosis))
confusion_matrix
```

## Hosmer-Lemeshow test for GOF
```{r , echo=FALSE}
# Use ResourceSelection to get the HosmerLemeshow test result
hl_test <- hoslem.test(df$diagnosis, df$predict.prob, g = 10)  
hl_test
```


## ROC curve and AUC (area under curve)
```{r mlog_reg-ROC, echo = TRUE}
roc_obj <- roc(df$diagnosis, df$predict.prob)

# Plot ROC curve
plot(roc_obj, col = "blue", main = "ROC Curve")

# Print AUC (AUC > 0.7)
auc(roc_obj)
```

## Get standardizec estimate
```{r std-est, echo = TRUE}
estimates(mlog_reg_ia)
```









### source: https://github.com/ThecoderPinar/Diabetes_Health_Prediction_and_Analysis
## Clean the data
```{r clean-data, echo=FALSE}
# Load the data
file <- "https://raw.githubusercontent.com/ThecoderPinar/Diabetes_Health_Prediction_and_Analysis/refs/heads/main/raw/diabetes_data.csv"
df <- import(file)
#write.csv(df, here("EpiData", "HWdata_2.csv"))

# Clean up variable names
df <- clean_names(df)
names(df) <- gsub("_", ".", names(df))

# Check the updated dataset structure
glimpse(df)

# Save as R data file
saveRDS(df, here("EpiData", "HWdata_2.rds"))
```

## Give variable explanations in labels
```{r labels, echo = TRUE}
attr(df$patient.id, "label") <- " 驻"
attr(df$age, "label") <- ""
attr(df$gender, "label") <- "专"
attr(df$ethnicity, "label") <- "转转"
attr(df$socioeconomic.status, "label") <- "爪 专转-"
attr(df$education.level, "label") <- "专转 砖"
attr(df$bmi, "label") <- "BMI"
attr(df$smoking, "label") <- "注砖"
attr(df$alcohol.consumption, "label") <- "爪专转 "
attr(df$physical.activity, "label") <- "驻注转 驻转"
attr(df$diet.quality, "label") <- "转 转"
attr(df$sleep.quality, "label") <- "转 砖"
attr(df$family.history.diabetes, "label") <- "住专 砖驻转转 住专转"
attr(df$gestational.diabetes, "label") <- "住专转 专"
attr(df$polycystic.ovary.syndrome, "label") <- "转住转 砖转 驻爪住转"
attr(df$previous.pre.diabetes, "label") <- "专-住专转"
attr(df$hypertension, "label") <- "抓  "
attr(df$systolic.bp, "label") <- "抓  住住"
attr(df$diastolic.bp, "label") <- "抓  住"
attr(df$fasting.blood.sugar, "label") <- "专转 住专  专 爪"
attr(df$hb.a1c, "label") <- "专转 住专  专 爪 -2,3 砖 专"
attr(df$serum.creatinine , "label") <- "转驻拽 转"
attr(df$bun.levels, "label") <- "专转 拽 砖转"
attr(df$cholesterol.total, "label") <- "住专 "
attr(df$cholesterol.ldl, "label") <- "住专 专注"
attr(df$cholesterol.hdl, "label") <- "住专 "
attr(df$cholesterol.triglycerides, "label") <- "专爪专"
attr(df$antihypertensive.medications, "label") <- "转专驻转 抓  "
attr(df$statins, "label") <- "转 住"
attr(df$antidiabetic.medications, "label") <- "转专驻转  住专转"
attr(df$frequent.urination, "label") <- "转 砖转 转祝"
attr(df$excessive.thirst, "label") <- "爪 "
attr(df$unexplained.weight.loss, "label") <- "专 砖拽 转-住专转"
attr(df$fatigue.levels, "label") <- "专转 注驻转"
attr(df$blurred.vision, "label") <- "专 砖砖转"
attr(df$slow.healing.sores, "label") <- "专驻  砖 驻爪注" 
attr(df$tingling.hands.feet, "label") <- "注拽爪抓  / 专"
attr(df$quality.of.life.score, "label") <- " 转 "
attr(df$heavy.metals.exposure, "label") <- "砖驻 转转 转"
attr(df$occupational.exposure.chemicals, "label") <- "砖驻 转注住拽转转 拽"
attr(df$water.quality , "label") <- "转 "
attr(df$medical.checkups.frequency, "label") <- "转专转 拽转 专驻"
attr(df$medication.adherence, "label") <- "转 转专驻转 驻 拽注"
attr(df$health.literacy, "label") <- "专转 专转"
attr(df$diagnosis, "label") <- ""
attr(df$doctor.in.charge, "label") <- "专驻 驻"

labels <- NULL
labels <- data.frame(Variable = names(df), Label = unlist(var_label(df)))
gt(labels)
```