###
### name   : [ שם משפחה + שם פרטי]
### ID     : [ מספר תעודת זהות ]
### course: אפידמיולוגיה (HIT.ac.il)
### lesson: שיעורי בית 3
### date  : [ תאריך ]
###   
###  שרידות
###
### source: https://quantdev.ssri.psu.edu/sites/qdev/files/12_SurvivalAnalysis_2020_0408.html

# Reset memory
```{r setup-packages, include false}
rm(list = ls(all.names = TRUE))
lapply(paste('package:',names(sessionInfo()$otherPkgs),sep=""),detach,character.only=TRUE,unload=TRUE,force=TRUE)
gc()
```

# Load required packages
```{r setup-packages, include false}
install.packages("pacman")
library(pacman)
pacman::p_load(
  rio,          # File import
  tidyverse,    # data management + ggplot2 graphics, 
  gt,           # Nice beuatiful tables
  nortest,      # Anderson-Darling test for normality
  ggplot2,      # For creating plots
  DescTools,    # For statistical functions like skewness and kurtosis
  janitor,      # adding totals and percents to tables
  car,           # For Levene's test and ANOVA
  gtsummary,    # summary statistics and tests
  corrr,        # correlation analayis for numeric variables
  dplyr,        # For data manipulation
  reshape2,     # For reshaping data
  skimr,        # get overview of data
  performance,           # For model performance metrics
  ResourceSelection,     # For goodness of fit tests
  broom,                 # For tidying model outputs
  pROC,                  # For ROC analysis
  caret,
  MASS,          # For statistical functions (polr for ordinal regression)
  epiDisplay,    # 
  survival,
  survminer
            )
```

# Get data
```{r load data, include false}
#file <- "HWdata_3.csv"     # use this if the file is on your computer
file <- "https://raw.githubusercontent.com/vincentarelbundock/Rdatasets/refs/heads/master/csv/survival/cancer.csv"
df7 <- import(file, trust  = TRUE)
summary(df7)
glimpse(df7)
```

## Survival time statistics
```{r surv-stats, echo = TRUE}
# Survival time
mean_survival_time <- mean(df7$time, na.rm = TRUE)
sd_survival_time <- sd(df7$time, na.rm = TRUE)
CI1 <- MeanCI(df7$time, na.rm = TRUE)
  lower_meanci_1 <- as.numeric((CI1["lwr.ci"]))
  upper_meanci_1 <- as.numeric((CI1["upr.ci"]))
median_survival_time <- median(df7$time, na.rm = TRUE)
IQR_survival_time <- IQR(df7$time, na.rm = TRUE)
q25_survival_time <- as.numeric(Quantile(df7$time, probs = 0.25, na.rm = TRUE))
q75_survival_time <- as.numeric(Quantile(df7$time, probs = 0.75, na.rm = TRUE))
min_survival_time <- min(df7$time, na.rm = TRUE)
max_survival_time <- max(df7$time, na.rm = TRUE)

# Create a data frame to neatly display the statistics
stats_df <- data.frame(
    Statistic = c("Mean", "SD", "⠀⠀CI₉₅ low", "⠀⠀CI₉₅ high", "Median", "⠀⠀Q1", "⠀⠀Q3", "⠀⠀IQR", "Minimum", "Maximum")
  , Value = c(mean_survival_time, sd_survival_time, lower_meanci_1, upper_meanci_1, median_survival_time, q25_survival_time, q75_survival_time, IQR_survival_time, min_survival_time, max_survival_time)
  , סטטיסטי = c("ממוצע", "ס'ת", "⠀⠀רווח סמך", "⠀⠀רווח סמך", "חציון", "⠀⠀רבעון 1", "⠀⠀רבעון 3", "⠀⠀מרווח בין רבועני", "מינימום", "מקסימום")
)

# Clean, simple table
summary_stats <- stats_df %>% 
  gt() %>%
  tab_header(title = "Survival Statistics⠀⠀⠀⠀זמן הישרדות") %>%
  fmt_number(columns = Value, decimals = 1)
summary_stats              # Show the table
```

## Do a Kaplan-Meyer survival analayis
```{r km-surv, echo = TRUE}
# estimate the survival probabilities for all patients
KM_all <- survfit(Surv(time = time, 
                       event = status == "2" ) ~ 1, 
                       data = df7)
summary(KM_all)
```


## Life table
```{r life-table, echo = TRUE}
# Convert status to int
df7 <- df7 %>%
  mutate(event = ifelse(status == "2", 1, 0))

# Fit Kaplan-Meier model stratified by stroke_type (example)
km_fit <- survfit(Surv(time, event) ~ 1, data = df7)

# Extract survival summary at intervals of 10 units
km_summary <- summary(km_fit, times = seq(0, max(df7$time), by = 100))

# Convert to data frame
km_df <- data.frame(
  time = km_summary$time,
  #strata = km_summary$strata,
  n_risk = km_summary$n.risk,
  n_event = km_summary$n.event,
  surv = km_summary$surv,
  std_err = km_summary$std.err,
  lower_ci = km_summary$lower,
  upper_ci = km_summary$upper,
  n_censor = km_summary$n.censor
)

# Display with gt
km_df %>%
  gt() %>%
  tab_header(title = "Survival analysis") %>%
    cols_label(
    time      = "Time",
    n_risk    = "At Risk",
    n_event   = "Deaths",
    surv      = "Survival %",
    std_err   = "Std. Error",
    lower_ci  = "Lower CI",
    upper_ci  = "Upper CI",
    n_censor  = "Censoerd"
  ) %>%
  fmt_percent(columns = c(surv, lower_ci, upper_ci), decimals = 1) %>%
  fmt_number(columns = std_err, decimals = 3) 
```

## Kaplan-Meyer survival analayis plot
```{r read-data, echo = TRUE}
# estimate the survival probabilities for all patients
KM_all_plot <- ggsurvplot(KM_all, 
           data = df7, 
           risk.table = TRUE, 
           linetype = c(1,4),
           tables.height = 0.3,
           pval = TRUE)

KM_all_plot
# *** You may need to zoom out ***
```





# Calculate the number exposed to risk accounting for half the censored individuals
race_life_table <- race_life_table %>%
mutate(exposed_to_risk = n_risk - 0.5 * n_censor)
  
# Calculate the survival proportion at each interval
race_life_table <- race_life_table %>%
mutate(survival_proportion = 1 - (n_event / exposed_to_risk))
  
# Calculate the cumulative survival proportion
race_life_table <- race_life_table %>%
mutate(cumulative_survival = cumprod(survival_proportion))
 
# Append to the main life table data frame
life_table <- rbind(life_table, race_life_table)




