###
### name   : [ 砖 砖驻 + 砖 驻专]
### ID     : [ 住驻专 转注转 转 ]
### course: 驻 (HIT.ac.il)
### lesson: 注 砖 3
### date  : [ 转专 ]
###   
###  砖专转


# Reset memory
```{r setup-packages, include false}
rm(list = ls(all.names = TRUE))
lapply(paste('package:',names(sessionInfo()$otherPkgs),sep=""),detach,character.only=TRUE,unload=TRUE,force=TRUE)
gc()
```

# Load required packages
```{r setup-packages, include false}
install.packages("pacman")
library(pacman)
pacman::p_load(
  Rcpp, devtools, gh, remotes,   # For installing packages/files online (like from GitHub)
  rio, here,              # File import and path support
  tidyverse,              # data management + ggplot2 graphics,
  broom,                  # For tidying model outputs
  dplyr,                  # For data manipulation
  gt,                     # Nice beuatiful tables
  nortest,                # Anderson-Darling test for normality
  ggplot2,                # For creating plots
  DescTools,              # For statistical functions like skewness and kurtosis
  psych,                  # to get a table of all summary statistics
  janitor,                # adding totals and percents to tables
  car,                    # For Levene's test and ANOVA
  gtsummary,              # summary statistics and tests in beautiful tables
  corrr,                  # correlation
  lmtest,                 # linear regression tets
  parameters,             # calculate standardized parameters/coefficients
  GGally,                 # to make a correlation matrix
  skimr,                  # get overview of data
  pROC,                   # For ROC analysis
  performance,            # For model performance metrics
   #MASS,                   # For statistical functions (polr for ordinal regression)ordinal
  ordinal,                # For ordinal regression statistical functions 
  caret,                  # For confusion matrix and classification metrics
  ggeffects,              # To display the confusion matrix
  generalhoslem,          # For Hosmer-Lemeshow test for goodness of fit 
  ResourceSelection,      # For Hosmer-Lemeshow test for goodness of fit 
  VGAM,                   # for multinomial logistic regression
  epiDisplay,             # For epi data type handling    
  survival,               # For survival analysis
  survminer               # For survival analysis and visualization
            )
```

# Get data
```{r load data, include false}
# file <- "./EpiData/HWdata_3.rds"            #  use this if the file is on your computer
file <- "https://raw.githubusercontent.com/Model-Lab-Net/Courses/refs/heads/main/Epi/Data/HWdata_3.rds"
df <- import(file, trust  = TRUE)
summary(df)
glimpse(df)
```

## Survival time statistics
```{r surv-stats, echo = TRUE}
# Survival time
mean_survival_time <- mean(df$time, na.rm = TRUE)
sd_survival_time <- sd(df$time, na.rm = TRUE)
CI1 <- MeanCI(df$time, na.rm = TRUE)
  lower_meanci_1 <- as.numeric((CI1["lwr.ci"]))
  upper_meanci_1 <- as.numeric((CI1["upr.ci"]))
median_survival_time <- median(df$time, na.rm = TRUE)
IQR_survival_time <- IQR(df$time, na.rm = TRUE)
q25_survival_time <- as.numeric(Quantile(df$time, probs = 0.25, na.rm = TRUE))
q75_survival_time <- as.numeric(Quantile(df$time, probs = 0.75, na.rm = TRUE))
min_survival_time <- min(df$time, na.rm = TRUE)
max_survival_time <- max(df$time, na.rm = TRUE)

# Create a data frame to neatly display the statistics
stats_df <- data.frame(
    Statistic = c("Mean", "SD", "CI low", "CI high", "Median", "Q1", "Q3", "IQR", "Minimum", "Maximum")
  , Value = c(mean_survival_time, sd_survival_time, lower_meanci_1, upper_meanci_1, median_survival_time, q25_survival_time, q75_survival_time, IQR_survival_time, min_survival_time, max_survival_time)
  , 住住 = c("爪注", "住'转", "专 住", "专 住", "爪", "专注 1", "专注 3", "专  专注", "", "拽住")
)

# Clean, simple table
summary_stats <- stats_df %>% 
  gt() %>%
  tab_header(title = "Survival Statistics 砖专转") %>%
  fmt_number(columns = Value, decimals = 1)
summary_stats              # Show the table
```

## Do a Kaplan-Meyer survival analayis
```{r km-surv, echo = TRUE}
# estimate the survival probabilities for all patients
KM <- survfit(Surv(time = time, 
                       event = status == "2" ) ~ 1, 
                       data = df)
KM
```


## Life table
```{r life-table, echo = TRUE}
# Convert status to int
df <- df %>%
  mutate(event = ifelse(status == "2", 1, 0))

# Fit Kaplan-Meier model stratified by stroke_type (example)
km_fit <- survfit(Surv(time, event) ~ 1, data = df)

# Extract survival summary at intervals of 10 units
km_summary <- summary(km_fit, times = seq(0, max(df$time), by = 100))

# Convert to data frame
km_df <- data.frame(
  time = km_summary$time,
  #strata = km_summary$strata,
  n_risk = km_summary$n.risk,
  n_event = km_summary$n.event,
  surv = km_summary$surv,
  std_err = km_summary$std.err,
  lower_ci = km_summary$lower,
  upper_ci = km_summary$upper,
  n_censor = km_summary$n.censor
)

# Display with gt
km_df %>%
  gt() %>%
  tab_header(title = "Life Table") %>%
    cols_label(
    time      = "Time",
    n_risk    = "At Risk",
    n_event   = "Deaths",
    surv      = "Survival %",
    std_err   = "Std. Error",
    lower_ci  = "Lower CI",
    upper_ci  = "Upper CI",
    n_censor  = "Censoerd"
  ) %>%
  fmt_percent(columns = c(surv, lower_ci, upper_ci), decimals = 1) %>%
  fmt_number(columns = std_err, decimals = 3) 
```

## Kaplan-Meyer survival analayis plot
```{r read-data, echo = TRUE}
# estimate the survival probabilities for all patients
KM_plot <- ggsurvplot(KM, 
           data = df, 
           risk.table = TRUE, 
           linetype = c(1,4),
           tables.height = 0.3,
           pval = TRUE)

KM_plot
# *** You may need to zoom out ***
```



## Calculate survivial probablities and append the life table
```{r surv-prob, echo = TRUE}
# estimate the survival probabilities for all patients
# Calculate the number exposed to risk accounting for half the censored individuals
km_df <- km_df %>%
    mutate(exposed_to_risk = n_risk - 0.5 * n_censor)
  
# Calculate the survival proportion at each interval
km_df <- km_df %>%
    mutate(survival_proportion = 1 - (n_event / exposed_to_risk))
  
# Calculate the cumulative survival proportion
km_df <- km_df %>%
    mutate(cumulative_survival = cumprod(survival_proportion))

km_df
```


### ---------------------------------------------------



## Do a stratified Kaplan-Meyer survival analayis 
```{r km-surv-strat, echo = TRUE}
# estimate the survival probabilities for stroke type
 KM_str <- survfit(Surv(time = time, 
                              event = status == "2" ) ~ sex, 
                         data = df)
summary(KM_str)
```

## Survival curve
```{r read-data, echo = TRUE}
# estimate the survival probabilities for all patients
KM_str_plot <- ggsurvplot(KM_str, 
           data = df, 
           risk.table = TRUE, 
           linetype = c(1,4),
           tables.height = 0.3,
           pval = TRUE,
           conf.int = TRUE)

KM_str_plot
# *** You may need to zoom out ***
```

## Life table
```{r life-table, echo = TRUE}
# Convert status to int
# Fit Kaplan-Meier model stratified by stroke_type (example)
km_fit_strat <- survfit(Surv(time, status) ~ sex, data = df)

# Extract survival summary at intervals of 10 units
km_summary_strat <- summary(km_fit_strat, times = seq(0, max(df$time), by = 100))

# Convert to data frame
km_df <- data.frame(
  time = km_summary_strat$time,
  strata = km_summary_strat$strata,
  n_risk = km_summary_strat$n.risk,
  n_event = km_summary_strat$n.event,
  surv = km_summary_strat$surv,
  std_err = km_summary_strat$std.err,
  lower_ci = km_summary_strat$lower,
  upper_ci = km_summary_strat$upper,
  n_censor = km_summary_strat$n.censor
)

# Display with gt
km_df %>%
  gt() %>%
  tab_header(title = "Survival analysis") %>%
    cols_label(
    time      = "Time",
    strata    = "Group",
    n_risk    = "At Risk",
    n_event   = "Deaths",
    surv      = "Survival %",
    std_err   = "Std. Error",
    lower_ci  = "Lower CI",
    upper_ci  = "Upper CI",
    n_censor  = "Censoerd"
  ) %>%
  fmt_percent(columns = c(surv, lower_ci, upper_ci), decimals = 1) %>%
  fmt_number(columns = std_err, decimals = 3) 
```


### ----------------------------------
### Compare strata

## Log-rank test
```{r read-data, echo = TRUE}
# Test if survival estimates are  different between levels/groups
log_rank_sex <- survdiff(Surv(time = time, 
              event = status == "2") ~ sex, 
         data = df,
         rho = 0)

log_rank_sex
```

## Peto's log-rank test
```{r read-data, echo = TRUE}
# Test if survival estimates are different between levels/groups
log_rank_peto_sex <- survdiff(Surv(time = time, 
              event = status == "2") ~ sex, 
         data = df,
         rho = 1)

log_rank_peto_sex
```










### source: https://github.com/vincentarelbundock/Rdatasets
## Clean the data
```{r clean-data, echo=FALSE}
# Load the data
file <- "https://raw.githubusercontent.com/vincentarelbundock/Rdatasets/refs/heads/master/csv/survival/cancer.csv"
df <- import(file)

# Clean up variable names
df <- clean_names(df)
names(df) <- gsub("_", ".", names(df))

# Check the updated dataset structure
glimpse(df)

# Save as R data file
saveRDS(df, here("EpiData", "HWdata_3.rds"))
```