###
### name: דוד בורג
### course: אפידמיולוגיה (HIT.ac.il)
### lesson: 1
### date: 19/05/2025
###
###  קורונה בערי ישראל
###
### source: https://data.gov.il/dataset/covid-19
###         https://www.cbs.gov.il/he/publications/Pages/2019/הרשויות-המקומיות-בישראל-קובצי-נתונים-לעיבוד-1999-2017.aspx
###         https://data.gov.il/dataset/residents_in_israel_by_communities_and_age_groups


## Reset memory
```{r setup-packages, include false}
rm(list = ls(all.names = TRUE))
lapply(paste('package:',names(sessionInfo()$otherPkgs),sep=""),detach,character.only=TRUE,unload=TRUE,force=TRUE)
gc()
```

load("covid.RData")

## Load required packages
```{r load-packages, include false}
options("install.lock"=FALSE)
install.packages("pacman")
library(pacman)
pacman::p_load(
  rio,          # File import
  here,         # File locator
  skimr,        # get overview of data
  tidyverse,    # data management + ggplot2 graphics, 
  gtsummary,    # summary statistics and tests
  rstatix,      # statistics
  nortest,      # Anderson-Darling test for normality
  moments,      # skewness and kurtosis
  corrr,        # correlation analayis for numeric variables
  janitor,      # adding totals and percents to tables
  flextable,    # converting tables to HTML
  dplyr,        # For data manipulation
  ggplot2,      # For creating plots
  plotly,
  DescTools,    # For statistical functions like skewness and kurtosis
  car,          # For Levene's test and ANOVA
  vctrs,         # For data type handling
  rvest,
  tidyr,
  lmtest,
  vars,
  httr,
  jsonlite,
  ISOweek,
  lubridate
            )
```
 
## Get COVID main data for Israel cities
```{r read-data, echo=FALSE}
# Define the API endpoint
api_endpoint <- "https://data.gov.il/api/3/action/datastore_search?resource_id=8a21d39d-91e3-40db-aca1-f73f7ab1df69&limit=400000"

# Send GET request to the API
response <- GET(api_endpoint)

# Check if the request was successful (status code 200)
if (status_code(response) == 200) {
      # Extract content as text and parse JSON
      content <- content(response, as = "text", encoding = "UTF-8")
      data <- fromJSON(content)
      records <- data$result$records    # Extract the records from the result
      df <- as.data.frame(records)      # Convert records to a data frame
      } else {
       cat("Error: Failed to fetch data. Status code:", status_code(response), "\n")
     }
```

## Cleanup data
```{r load data, include false}
# Setting the seed based on the current time
df <- df[, -c(1)]                           # Remove columns
colnames(df) <- c("city", "cityid", "date", "cases", "recovered", "deaths", "tests", "positive", "color", "score")
df$cityid <- as.numeric(df$cityid)
df$date  <- as.Date(df$date, format = "%Y-%m-%d")
df$cases <- as.numeric(df$cases)
    df$cases[df$cases == 0] <- NA
df$recovered <- as.numeric(df$recovered)
    df$recovered[df$recovered == 0] <- NA
df$deaths <- as.numeric(df$deaths)
    df$deaths[df$deaths == 0] <- NA
df$tests <- gsub("<15", "5", df$tests)
    df$tests <- as.numeric(df$tests)
    df$tests[df$tests == 0] <- NA
df$positive <- gsub("<15", "5", df$positive)
df$positive <- as.numeric(df$positive)
    df$positive[df$positive == 0] <- NA
df$score <- as.numeric(df$score)
df$color <- gsub("NULL", "--", df$color)
df$color <- gsub("ירוק", "green", df$color)
df$color <- gsub("צהוב", "yellow", df$color)
df$color <- gsub("כתום", "orange", df$color)
df$color <- gsub("אדום", "red", df$color)

df$year <- year(df$date)
```

## Check out the variables
```{r load data, include false}
# Setting the seed based on the current time
glimpse(df)
```

## Get population data for Israel cities
```{r read-data, echo=FALSE}
# Get population for 2020 (this will be the main df)
file <- "https://www.cbs.gov.il/he/publications/DocLib/2019/hamakomiot1999_2017/2020.xlsx"
df_pop <- import(file, which = 2, trust  = TRUE)
df_pop <- df_pop[, c(1, 2, 13)]
df_pop <- df_pop[c(5:205), ]                              # keep rows
colnames(df_pop) <- c("city", "cityid", "pop2020")
df_pop$cityid <- as.numeric(df_pop$cityid)
df_pop$pop2020 <- as.numeric(df_pop$pop2020)
glimpse(df_pop)

file <- "https://www.cbs.gov.il/he/publications/DocLib/2019/hamakomiot1999_2017/2020.xlsx"
df_area2020 <- import(file, which = 3, trust  = TRUE)
df_area2020 <- df_area2020[, c(1, 2, 74)]
df_area2020 <- df_area2020[c(7:207), ]                              # keep rows
colnames(df_area2020) <- c("city", "cityid", "apt_area")
### estimate apt_area for 7300, 1247, 3608, 0520
df_area2020$apt_area[df_area2020$cityid == "7300"] <- "1545"
df_area2020$apt_area[df_area2020$cityid == "1247"] <- "439"
df_area2020$apt_area[df_area2020$cityid == "3608"] <- "36"
df_area2020$apt_area[df_area2020$cityid == "0520"] <- "205"
df_area2020$cityid <- as.numeric(df_area2020$cityid)
df_area2020$apt_area <- as.numeric(df_area2020$apt_area)
glimpse(df_area2020)

#combine 
df_pop <- df_pop %>%
  left_join(df_area2020 %>% dplyr::select(cityid, apt_area), by = "cityid")

# ---------------------------------

# Get population for 2021 and built area data (this will be appended to main df)
file <- "https://www.cbs.gov.il/he/publications/DocLib/2019/hamakomiot1999_2017/p_libud_2021.xlsx"
df_2021 <- import(file, which = 2, trust  = TRUE)
df_2021 <- df_2021[, c(1, 2, 13, 245)]
df_2021 <- df_2021[c(5:205), ]                              # keep rows
colnames(df_2021) <- c("city", "cityid", "pop2021", "built_area")
df_2021$cityid <- as.numeric(df_2021$cityid)
df_2021$pop2021 <- as.numeric(df_2021$pop2021)
df_2021$built_area <- as.numeric(df_2021$built_area)
glimpse(df_2021)
#combine 
df_pop <- df_pop %>%
  left_join(df_2021 %>% dplyr::select(cityid, built_area, pop2021), by = "cityid")

# ---------------------------------

# Get population for 2022 (this will be appended to main df)
file <- "https://www.cbs.gov.il/he/publications/DocLib/2019/hamakomiot1999_2017/p_libud_22.xlsx"
df_2022 <- import(file, which = 2, trust  = TRUE)
df_2022 <- df_2022[, c(1, 2, 13)]
df_2022 <- df_2022[c(9:210), ]                              # keep rows
colnames(df_2022) <- c("city", "cityid", "pop2022")
df_2022$cityid <- as.numeric(df_2022$cityid)
df_2022$pop2022 <- as.numeric(df_2022$pop2022)
df_2022$pop2022 <- round(as.numeric(df_2022$pop2022), 0)
#combine 
df_pop <- df_pop %>%
  left_join(df_2022 %>% dplyr::select(cityid, pop2022), by = "cityid")

# ---------------------------------

# Get population for 2023 (this will be appended to main df)
years <- c(2020, 2021, 2022)
df_pop$pop2023 <- apply(df_pop[, c("pop2020", "pop2021", "pop2022")], 1, function(pops) {
  model <- lm(pops ~ years)
  predict(model, newdata = data.frame(years = 2023))
})
df_pop$pop2023 <- round(as.numeric(df_pop$pop2023), 0)

colnames(df_pop) <- c("city", "cityid", "2020", "apt_area", "built_area", "2021", "2022", "2023")
df_pop$built_area <- df_pop$built_area * 1000000
df_pop$apt_area <- df_pop$apt_area * 1000

# Get english city names (this will be appended to main df)
file <- "https://www.cbs.gov.il/he/publications/LochutTlushim/2020/%D7%90%D7%95%D7%9B%D7%9C%D7%95%D7%A1%D7%99%D7%99%D7%94.xlsx"
df_names <- import(file, skip = 7, trust  = TRUE)
df_names <- df_names[ , c(1,9)]                           # Remove columns
colnames(df_names) <- c("cityid", "name")
df_names$cityid <- as.numeric(df_names$cityid)
glimpse(df_names)
write_csv(df_names, "Israel city names.csv")
```


## combine
```{r read-data, echo=FALSE}
df <- merge(df, 
            df_pop[, c("cityid", "built_area", "apt_area"), drop = FALSE], 
            by = "cityid", all.x = TRUE)
print(head(df, n = 30))

df_pop_long <- df_pop %>%
  pivot_longer(cols = c("2020", "2021", "2022", "2023"),
               names_to = "year",
               values_to = "pop") %>%
  mutate(year = as.integer(year))  # make year numeric to match df$year

# Step 2: Join the population to df based on cityid and year
df <- df %>%
  left_join(df_pop_long, by = c("cityid", "year"))

df <- df %>%
  arrange(cityid, date)

df <- df %>% dplyr::select(-city.y, -built_area.y, -apt_area.y)
df <- df %>% rename(city = city.x, built_area = built_area.x, apt_area = apt_area.x)


df <- merge(df, 
            df_names[, c("cityid", "name"), drop = FALSE], 
            by = "cityid", all.x = TRUE)
```

## calculate density
```{r read-data, echo=FALSE}
df <- df %>%
  mutate(pop_density = pop / (built_area/1000))
    df$pop_density[df$pop_density == 0] <- NA

df <- df %>%
  mutate(tests_density = tests / pop_density)
    df$tests_density[df$tests_density == 0] <- NA

df <- df %>%
  mutate(cases_density = cases / pop_density)
    df$cases_density[df$cases_density == 0] <- NA

df <- df %>%
  mutate(deaths_density = deaths / pop_density)
    df$deaths_density[df$deaths_density == 0] <- NA

df <- df %>%
  mutate(recovered_density = recovered / pop_density)
    df$recovered_density[df$recovered_density == 0] <- NA

glimpse(df)
print(head(df, n = 30))

#df1 <- df
```


## graph
```{r read-data, echo=FALSE}
city_name <-  "TEL AVIV - YAFO"

one_city_df <- df %>%
  filter(name == city_name)  # or filter(cityid == 123)


# Process cases
one_city_cases <- one_city_df %>%
  arrange(date) %>%
  mutate(
    daily_cases = cases_density - lag(cases_density),    #get the rate from the cum.
    week = isoweek(date),
    year = isoyear(date)
  ) %>%
  group_by(year, week) %>%
  summarise(
    avg_weekly_cases = mean(daily_cases, na.rm = TRUE),
    .groups = "drop"
  )

one_city_cases <- one_city_cases %>%
  mutate(
    week_str = sprintf("%d-W%02d", year, week),
    date = ISOweek2date(paste0(week_str, "-2"))  # "-2" = Tuesday
  )
one_city_cases <- one_city_cases %>%
  filter(avg_weekly_cases > 0 | is.na(avg_weekly_cases))

glimpse(one_city_deaths)


# Process deaths
one_city_deaths <- one_city_df %>%
  arrange(date) %>%
  mutate(
    daily_deaths = deaths_density - lag(deaths_density),    #get the rate from the cum.
    week = isoweek(date),
    year = isoyear(date)
  ) %>%
  group_by(year, week) %>%
  summarise(
    avg_weekly_deaths = mean(daily_deaths, na.rm = TRUE),
    .groups = "drop"
  )

one_city_deaths <- one_city_deaths %>%
  mutate(
    week_str = sprintf("%d-W%02d", year, week),
    date = ISOweek2date(paste0(week_str, "-2"))  # "-2" = Tuesday
  )
one_city_deaths <- one_city_deaths %>%
  filter(avg_weekly_deaths > 0 | is.na(avg_weekly_deaths))


# Process tests
one_city_tests <- one_city_df %>%
  arrange(date) %>%
  mutate(
    daily_tests = tests_density - lag(tests_density),    #get the rate from the cum.
    week = isoweek(date),
    year = isoyear(date)
  ) %>%
  group_by(year, week) %>%
  summarise(
    avg_weekly_tests = mean(daily_tests, na.rm = TRUE),
    .groups = "drop"
  )

one_city_tests <- one_city_tests %>%
  mutate(
    week_str = sprintf("%d-W%02d", year, week),
    date = ISOweek2date(paste0(week_str, "-2"))  # "-2" = Tuesday
  )
one_city_tests <- one_city_tests %>%
  filter(avg_weekly_cases > 0 | is.na(avg_weekly_cases))

glimpse(one_city_tests)

# Process recovered
one_city_recovered <- one_city_df %>%
  arrange(date) %>%
  mutate(
    daily_recovered = recovered_density - lag(recovered_density),    #get the rate from the cum.
    week = isoweek(date),
    year = isoyear(date)
  ) %>%
  group_by(year, week) %>%
  summarise(
    avg_weekly_recovered = mean(daily_recovered, na.rm = TRUE),
    .groups = "drop"
  )

one_city_recovered <- one_city_recovered %>%
  mutate(
    week_str = sprintf("%d-W%02d", year, week),
    date = ISOweek2date(paste0(week_str, "-2"))  # "-2" = Tuesday
  )
one_city_recovered <- one_city_recovered %>%
  filter(avg_weekly_cases > 0 | is.na(avg_weekly_cases))

glimpse(one_city_recovered)


# Append to one_city_data
one_city_data <- reduce(
  list(
    one_city_cases,
    one_city_deaths,
    one_city_tests,
    one_city_recovered
  ),
  full_join,
  by = c("year", "week", "date")
)

one_city_data <- one_city_data[, -c(4, 7, 9, 11)]                           # Remove columns

one_city_data <- one_city_data %>%
  mutate(active_cases =  avg_weekly_cases - avg_weekly_deaths - avg_weekly_recovered)
#    one_city_data$cases_density[one_city_data$cases_density == 0] <- NA


glimpse(one_city_data)

# move deaths two weeks ahead
#one_city_deaths <- one_city_deaths %>%
#  arrange(date) %>%
#  mutate(
#    avg_weekly_deaths = lead(avg_weekly_deaths, 2)
#  )

## graph
```{r read-data, echo=FALSE}
plot <- plot_ly(data = one_city_data, 
                x = ~ date, 
                y = ~ avg_weekly_tests, 
                type = "bar", 
                marker = list(color = "darkgrey")) %>%
  add_trace(y = ~ avg_weekly_cases, 
            type = "bar", 
            name = "Deaths",  # Name for legend
            marker = list(color = "darkorange")) %>%
  add_trace(y = ~ avg_weekly_deaths, 
            type = "bar", 
            name = "Deaths",  # Name for legend
            marker = list(color = "darkred")) %>%
#  add_trace(y = ~ active_cases, 
#            type = "scatter",
#            mode = "lines", 
#            name = "Active",  # Name for legend
#            line = list(color = "darkgreen"),
#            marker = list(color = "darkgreen"),
#            yaxis = "y2"
#           ) %>%             
  layout(
    title = list(
      text = paste0("COVID-19 Cases Over Time<br><sup>City: ", city_name, "</sup>"), 
      x = 0.5, 
      xanchor = "center"
    ),
    xaxis = list(
      title = "Date",
      range = c("2020-01-01", "2023-07-01")  # ❗️Start x-axis at Feb 1, 2020
    ),
    yaxis = list(
      title = "Cases per population density<br><sup>(Log10 Scale)</sup>", 
      type = "log", 
      range = c(log10(0.001), log10(10000))
    ),
    barmode = "overlay",  # Set bars to overlap
    bargap = 0,
    showlegend = FALSE,
    paper_bgcolor = "rgba(0,0,0,0)",
    plot_bgcolor = "rgba(0,0,0,0)"
  )
# Display the plot
plot
```

## urban scaling
```{r read-data, echo=FALSE}

df_date <- df %>% filter(date == as.Date("2021-09-15"))
glimpse(df_date)
write_csv(data.frame(anme = df_date$name), "1.csv")
plot <- plot_ly(
  data = df_date,
  x = ~ pop_density,
  y = ~ cases,
  type = "scatter",
  mode = "markers",
  text = ~paste("Name:", name),  # tooltip content
  marker = list(color = "darkblue", size = 6)
) %>%
  layout(
    title = "Urban Scaling<br><sup>Allometry",
    xaxis = list(
      title = "log Population Density<br><sup>people per dunam",
      type = "log",
      range = c(log10(1), NA)  # Start at 1 to avoid log(0)
    ),
    yaxis = list(
      title = "log Cases",
      type = "log",
      range = c(log10(1), NA)
    ),
    plot_bgcolor = "rgba(0,0,0,0)",
    paper_bgcolor = "rgba(0,0,0,0)"
  )

plot
```

## Power law fit
```{r read-data, echo=FALSE}

```

## 3D urban scaling
```{r read-data, echo=FALSE}
plot <- plot_ly(
  data = df_date,
  x = ~tests,
  y = ~cases,
  z = ~deaths,
  type = "scatter3d",
  mode = "markers",
  marker = list(size = 4, color = ~cases, colorscale = "Viridis"),
  text = ~paste("Name:", name, 
                "<br>Tests:", tests, 
                "<br>Cases:", cases, 
                "<br>Deaths:", deaths),
  hoverinfo = "text"
) %>%
  layout(
    title = "3D Scatter Plot: Tests vs. Cases vs. Deaths",
    scene = list(
      xaxis = list(title = "Tests", type = "log"),
      yaxis = list(title = "Cases", type = "log"),
      zaxis = list(title = "Deaths", type = "log")
    )
  )

plot
```







save.image(file = "covid.RData")
