###
### name: David Burg, Jesse Ausubel
### course: PHE, Rockefeller University
### 
### date: 04/07/2025
###
###  US immigration 2019-present
###
### source: https://ohss.dhs.gov/topics/immigration/yearbook
###         https://csslab.uc3m.es/dataviz/projects/2023/100407614/


## Reset memory
```{r setup-packages, include false}
rm(list = ls(all.names = TRUE))
lapply(paste('package:',names(sessionInfo()$otherPkgs),sep=""),detach,character.only=TRUE,unload=TRUE,force=TRUE)
gc()
```

## Load required packages
```{r load-packages, include false}
options("install.lock"=FALSE)
install.packages("pacman")
library(pacman)
pacman::p_load(
  rio,          # File import
  here,         # File locator
  skimr,        # get overview of data
  tidyverse,    # data management + ggplot2 graphics, 
  gtsummary,    # summary statistics and tests
  rstatix,      # statistics
  nortest,      # Anderson-Darling test for normality
  moments,      # skewness and kurtosis
  corrr,        # correlation analayis for numeric variables
  janitor,      # adding totals and percents to tables
  flextable,    # converting tables to HTML
  dplyr,        # For data manipulation
  ggplot2,      # For creating plots
  plotly,
  DescTools,    # For statistical functions like skewness and kurtosis
  car,          # For Levene's test and ANOVA
  vctrs,         # For data type handling
  rvest,
  tidyr,
  lmtest,
  vars
            )
```
 
## Get US measles
```{r read-data, echo=FALSE}
## read the data in a file online
file <- "https://ourworldindata.org/grapher/sales-of-cigarettes-per-adult-per-day.csv?v=1&csvType=full&useColumnShortNames=true"
file_cig  <- import(file, trust  = TRUE)
file <- "https://ourworldindata.org/grapher/lung-cancer-death-rates-males.csv?v=1&csvType=full&useColumnShortNames=true"
file_cancer  <- import(file, trust  = TRUE)


us_cig <- file_cig %>%
  filter(Entity == "United States")
colnames(us_cig) <- c("country", "code", "year", "cigarettes")

us_cancer <- file_cancer %>%
  filter(Entity == "United States")
colnames(us_cancer) <- c("country", "code", "year", "cancer")

# Merge datasets for plotting
df <- merge(us_cig, us_cancer, by = "year", all = TRUE)
```

## Check out the variables
```{r load data, include false}
# Setting the seed based on the current time
glimpse(df)
```

## Check correlation between cancer and cigarettes
```{r load data, include false}
# Perform a Simple Linear Regression
slr <- lm(cancer ~ cigarettes, data = df)

# use gtsummary to create a nice table
slr_summary <-
  slr %>% 
  tbl_regression() %>%
  add_glance_table(include = c(nobs,
                               r.squared,
                               statistic,
                               logLik,
                               AIC,
                               sigma,
                               p.value))

# Show the table
slr_summary
```


## Show the scatter plot with regression line
```{r slr-gtsummary, include=FALSE}
# use ggplot to show the scatter plot with regression line with linear model and confidence interval
slr_plot <- ggplot(df, aes_string(x = cigarettes, y = cancer)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, color = "blue") +  # Adding the trend line with linear model
  labs(title = paste("Scatter Plot of cigarettes vs cancer"),
       x = cigarettes,
       y = cancer) +
  theme_minimal()
ggplotly(slr_plot)
```


## plot the cancer data
```{r deaths, include false}
us_cancer_plot <- ggplot(df, aes(x = year, y = cancer)) +
  geom_line(color = "darkred", size = 1) +
  #geom_point(color = "darkred", size = 2) +
  labs(
    title = "Number of Measles cases in USA",
    x = "Year",
    y = "Cases"
  ) +
  scale_y_log10() +  # Add logarithmic y-axis
  theme_minimal()
ggplotly(us_cancer_plot)
```



## plot the cig data
```{r deaths, include false}
us_plot <- ggplot(df, aes(x = year, y = cigarattes)) +
  geom_line(color = "steelblue", size = 1) +
  #geom_point(color = "darkred", size = 2) +
  labs(
    title = "Number of cigarettes sold in USA",
    x = "Year",
    y = "Cigarettes per person"
  ) +
  scale_y_log10() +  # Add logarithmic y-axis
  theme_minimal()
ggplotly(us_plot)
```

## plot the cancer data
```{r deaths, include false}
# Assuming us_cig and us_cancer are data frames with year, cigarattes, and M


# Calculate scaling factor for secondary axis
# Map cancer data (M) to cigarette data range
cig_range <- range(data$cigarattes, na.rm = TRUE)
cancer_range <- range(data$M, na.rm = TRUE)
scale_factor <- (cig_range[2] - cig_range[1]) / (cancer_range[2] - cancer_range[1])
offset <- cig_range[1] - cancer_range[1] * scale_factor

# Create the plot
us_combined_plot <- ggplot(data, aes(x = year)) +
  geom_line(aes(y = cigarattes, color = "Cigarette Cases"), size = 1) +
  geom_line(aes(y = M * scale_factor + offset, color = "Cancer Cases"), size = 1) +
  scale_y_continuous(
    name = "Cigarette Cases",
    sec.axis = sec_axis(
      ~ (. - offset) / scale_factor,
      name = "Cancer Cases"
    )
  ) +
  scale_color_manual(
    values = c("Cigarette Cases" = "steelblue", "Cancer Cases" = "darkred"),
    name = "Legend"
  ) +
  labs(
    title = "Cigarette and Cancer Cases in USA",
    x = "Year"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.title.y.right = element_text(color = "darkred"),
    axis.title.y.left = element_text(color = "steelblue")
  )

# Convert to interactive plot
ggplotly(us_combined_plot)
```

## plot the cancer data
```{r deaths, include false}
df_combined <- inner_join(us_cig, us_cancer, by = "Year")

model <- lm(Lung_cancer_deaths ~ Cigarettes_per_adult_per_day, data = df)
summary(model)

library(ggplot2)

ggplot(df, aes(x = Cigarettes_per_adult_per_day, y = Lung_cancer_deaths)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, color = "blue") +
  labs(
    title = "Lung Cancer Deaths vs Cigarette Sales",
    x = "Cigarettes per Adult per Day",
    y = "Lung Cancer Deaths"
  ) +
  theme_minimal()
```

## Do a causaility test
```{r deaths, include false}
# Select optimal lag length
ts_data <- data.frame(df[, c("year", "cigarettes", "cancer")])
ts_data <- na.omit(ts_data)

lag_selection <- VARselect(ts_data, lag.max = 20, type = "const")
lag_selection$selection

grangertest(cancer ~ cigarettes, order = 20, data = df)


```