###
### name  : __________________
### ID    : [מספר תעודת זהות]
### course: אפידמיולוגיה (HIT.ac.il)
### lesson: רגרסיה לינארית 05   
### date  : 07/08/2026
###  
### multivariate נלמד לעשות רגרסיה לינארית מרובת משתנים
###
### source: https://bookdown.org/drki_musa/dataanalysis/linear-regression.html#simple-linear-regression


## Reset memory
```{r setup-packages, include = FALSE}
rm(list = ls(all.names = TRUE))          # 👈 Clear all data objects        👇 Clear packages
lapply(paste("package:", names(sessionInfo()$otherPkgs), sep = ""), detach, character.only = TRUE, unload = TRUE, force = TRUE)
gc()                                     # Clear unused data from memory ("garbage cllection")
dev.off()                                # Reset the display output of R
system("powershell -Command Clear-Host") # This will clear the terminal in VScode
```

## Load required packages
```{r load-packages, include = FALSE}
options("install.lock" = FALSE)
if (!require("pacman")) install.packages("pacman") ; library(pacman)
pacman::p_load(
  Rcpp, devtools, gh, remotes,   # For installing packages/files online (like from GitHub)
  rio, here,              # File import and path support
  labelled,               # deal with labels
  tidyverse,              # data management + ggplot2 graphics,
  broom,                  # For tidying model outputs
  dplyr,                  # For data manipulation
  gt,                     # Nice beuatiful tables
  gtsummary,              # summary statistics and tests
  nortest,                # Anderson-Darling test for normality
  ggplot2,                # For creating plots
  ggpubr,                 # For ggplot2 extensions
  DescTools,              # For statistical functions like skewness and kurtosis
  summarytools,           # to get a table of all summary statistics
  car,                    # For Levene's test and ANOVA
  lmtest,                 # do linear regression tests
  parameters,             # get standardized parameters/coefficients
  performance,            # get diagnosticsa and model performance
  GGally                  # to make a correlation matrix
              )
# Install/load flexplot
if (!require("flexplot", quietly = TRUE)) devtools::install_github("dustinfife/flexplot", ref = "logistic_plots")  ;  library(flexplot)
```

## Access the data
```{r read-data, echo = TRUE}
# file <- "./EpiData/03 coronary.rds" # 👈 use this if the file is on your computer
file <- "https://raw.githubusercontent.com/Model-Lab-Net/Courses/refs/heads/main/Epi/Data/03%20coronary.rds"
df2 <- import(file, trust = TRUE)
# show variable translations to Hebrew in lables
labels <- data.frame(Variable = names(df2), Label = unlist(var_label(df2)))   ;   gt(labels)
glimpse(df2)               # get an overview of the data
```

## Simple linear regression model
```{r slr, echo = TRUE}
# Perform a Simple Linear Regression
slr <- lm(chol ~ sbp, data = df2)
# print the summary of the simple linear regression
summary(slr)

# use gtsummary to create a nice table
slr_model <- tbl_regression(slr, intercept = TRUE) %>%
  add_glance_table(include = c(
    nobs,
    r.squared,
    adj.r.squared,
    statistic,
    sigma,
    logLik,
    AIC,
    p.value
  ))
# Show the model results table
slr_model
```

## Flexplot scatter plot and analysis
```{r polt-regression, echo = TRUE}
# Plot the regression
flexplot(chol ~ sbp, data = df2, method = "lm", se = TRUE)

# Analyze the residuals
visualize(slr)
```

## Correlation matrix
```{r correlation-matrix, echo = TRUE}
ggcorr(df2, label = TRUE)
```

## Multi-variate linear regression model
```{r slr, echo = TRUE}
# Perform a Simple Linear Regression
mlr <- lm(chol ~ sbp + race + bmi + age + dbp, data = df2)
# print the summary of the simple linear regression
summary(mlr)

# use gtsummary to create a nice table
mlr_model <- tbl_regression(mlr, intercept = TRUE) %>%
  add_glance_table(include = c(
    nobs,
    r.squared,
    adj.r.squared,
    #statistic,
    #sigma,
    #logLik,
    AIC,
    p.value
  ))

# Show the model results table
mlr_model
```

## Flexplot scatter plot and analysis
```{r polt-regression, echo = TRUE}
# Plot the regression
flexplot(chol ~ sbp + race + bmi + age + dbp, data = df2, method = "lm", se = TRUE)

# Analyze the residuals
visualize(mlr)
```

## Diagnostics  (Ho: residuals are normal)
```{r polt-regression, echo = TRUE}
sw <- check_normality(mlr)                # Shapiro-Wilk for normality
bp <- check_heteroscedasticity(mlr)       # Breusch-Pagan for heteroscedacity
dw <- check_autocorrelation(mlr)          # Durbin-Watson test for autocorelation
vif <- check_collinearity(mlr)            # mullticolinearity (<5 low, >10 high)

# Add diagnostics to model table
mlr_model %>%
  as_gt() %>% # Convert to gt engine first
  tab_source_note(
    source_note = html(paste0(
      "Shapiro-Wilk (normality):⠀⠀⠀⠀⠀⠀⠀p = ", round(sw, 3), "<br>",
      "Breusch-Pagan (heteroscedasticity): p = ", round(bp, 3),"<br>",
      "Durbin-Watson (autocorrelation): ⠀ p = ", round(dw, 3),"<br>"
    )))
  
# Multicolinearity test
vif %>%
  as.data.frame() %>%
  select(Term, VIF) %>%
  mutate(
    Correlation = case_when(
      VIF < 5 ~ "Low ",
      VIF < 10 ~ "Moderate",
      TRUE ~ "High"
    )
  ) %>%
  gt() %>%
  fmt_number(
    columns = where(is.numeric),
    decimals = 1
  )
```

## Get the standardized parameter values
```{r std-parms, echo = TRUE}
model_std <- model_parameters(mlr, standardize = "refit")

model_std %>%
  as.data.frame() %>%
  dplyr::select(Parameter, beta = Coefficient, CI_low, CI_high, p) %>%
  gt() %>%
  fmt_number(
    #columns = c(Coefficient, SE, CI_low, CI_high),
    decimals = 3
  )
```




### -------------------------------------------------
## תרגול-עצמי

## שאלה 1
```{r question-1, echo = TRUE}
# 

```

## שאלה 2
```{r question-2, echo = TRUE}
# 

```

## שאלה 3
```{r question-3, echo = TRUE}
# 

```









### source: https://bookdown.org/drki_musa/dataanalysis/linear-regression.html#multiple-linear-regression
###         https://github.com/drkamarul/multivar_data_analysis/tree/main/data

## Clean data
```{r clean-data, echo = TRUE}
# Load the data
file <- "https://raw.githubusercontent.com/drkamarul/multivar_data_analysis/refs/heads/main/data/coronary.dta"
df2 <- import(file, trust = TRUE)
#write.csv(df2, here("EpiData", "03 coronary.csv"))

# Change id to factor
df2 <- df2 %>%
  mutate(id = as.factor(id))

# Change cad to factor
df2 <- df2 %>%
  mutate(cad = as.factor(cad))

# Change gender column from 1, 2 to "F", "M"
df2 <- df2 %>%
  mutate(gender = case_when(
    gender == 1 ~ "F",
    gender == 2 ~ "M",
    TRUE ~ as.factor(gender) # Preserve other values (e.g., NA)
  ))

# Change race column from 1, 2, 3  to "W", "B", "J"
df2 <- df2 %>%
  mutate(race = case_when(
    race == 1 ~ "W",
    race == 2 ~ "B",
    race == 3 ~ "J",
    TRUE ~ as.factor(race) # Preserve other values (e.g., NA)
  ))

# Convert all chr variables to fct
df2 <- mutate_if(df2, is.character, as.factor)

# Check the updated dataset structure
glimpse(df2)

# Save as R data file
saveRDS(df2, here("EpiData", "03 coronary.rds"))
```

