###
### name: דוד בורג
### course: אפידמיולוגיה (HIT.ac.il)
### lesson: 0   
### date: 19/06/2025
###  
###  זה קוד שיוריד את כל החבילות והקבצים שנצטרך בקורס הזה
###


# Reset memory
```{r setup-packages, include false}
rm(list = ls(all.names = TRUE))
lapply(paste('package:',names(sessionInfo()$otherPkgs),sep=""),detach,character.only=TRUE,unload=TRUE,force=TRUE)
gc()
```


### ------------------------------------------------


# Load required packages
```{r setup-packages, include false}
options("install.lock"=FALSE)
# List of packages to install and load
packages <- c("readr", "tidyverse", "tibble", "gt", "plotly", "dplyr",
              "skimr", "kableExtra", "rmarkdown", "descr", "knitr",
              "lubridate", "rio", "here", "patchwork", "readxl",
              "forcats", "haven", "haven", "ggpubr", "GGally", "rsq",
              "broom",
              "languageserver"
             )

# Loop to check, install, and load packages
for (pkg in packages) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    install.packages(pkg, dependencies = TRUE)
    cat("Installed package:", pkg, "\n")
  }
  library(pkg, character.only = TRUE)
}

# This will install the LogisticDx package from the archive
# Note: The package may not be available on CRAN anymore,
# so the URL is used
install.packages(c('rms', 'statmod', 'speedglm', 'pROC', 'aod'))
install.packages(
  "https://cran.r-project.org/src/contrib/Archive/LogisticDx/LogisticDx_0.3.tar.gz",
dependencies = TRUE
)
```


### ------------------------------------------------


# Download code examples for Course
```{r download-data, include false}
# List of files to download
files <- c("# 1st R code.Rmd",
           "# Histogram.Rmd",
           "# simple boxplot.Rmd",
           "# simple regression",
           "01 --------------",
           "01 Intro.R",
           "02-04 Univariate analysis.Rmd"
           "05a Linear Regression.Rmd",
           "05b Multi-variate Regression.Rmd",
           "06a Logistic Regression (OR).Rmd",
           "06b Multi-variate Logistic Regression.Rmd",
           "06c Ordinal Regression.Rmd",
           "07a Poisson Regression (RR).Rmd",
           "07b Poisson Regression (rates).Rmd",
           "07c Poisson Regression (quasi).Rmd",
           "08a Survival analysis (KM).Rmd",
           "08b Survival analysis (life tables).Rmd",
           "09 Cox regression.Rmd",
           "100 ------------",
           "HW 1.Rmd",
           "HW 2.Rmd",
           "HW 3.Rmd",
           "z --------------",
           "z John Snow Cholera London_ggplot.Rmd",
           "z John Snow Cholera London_plotly.Rmd",
          )
# Base URL
base_url <- "https://github.com/Model-Lab-Net/Courses/tree/main/Epi/Code"

# Destination folder
dest_folder <- "C:/RVSCode/Course/EpiCode"
# Create destination folder if it doesn't exist
if (!dir.exists(dest_folder)) {
  dir.create(dest_folder, recursive = TRUE)
}

# Loop through each file and download
for (file in files)
    {
       file_url <- paste0(base_url, file)                            # Construct full URL
       dest_path <- file.path(dest_folder, file)                     # Construct destination path
       download.file(file_url, destfile = dest_path, mode = "wb")    # Download file to specified folder
       cat("Downloaded:", file, "to", dest_path, "\n")               # Print status 
    }
```


### ------------------------------------------------


# Download HW data files (just in case)
```{r download-data, include false}
# file url and destination information
download.file("https://raw.githubusercontent.com/GTPB/PSLS20/master/data/fev.txt", destfile = "C:/RVSCode/Course/EpiCode/HWdata_1.txt")
download.file("https://raw.githubusercontent.com/ThecoderPinar/Diabetes_Health_Prediction_and_Analysis/refs/heads/main/raw/diabetes_data.csv", destfile = "C:/RVSCode/Course/EpiCode/HWdata_2.csv")
download.file("https://raw.githubusercontent.com/ThecoderPinar/Diabetes_Health_Prediction_and_Analysis/refs/heads/main/raw/diabetes_data.csv", destfile = "C:/RVSCode/Course/EpiCode/HWdata_3.csv")
```


### ------------------------------------------------


# Download exercise data files (just in case)
```{r download-data, include false}
# List of files to download
files <- c("asthma.csv", "coronary.dta", "coronaryNA.csv", "metabolic_syndrome.csv",
           "metabolic_syndrome.dta", "omega3.sav","peptic_ulcer.xlsx","smoke.csv",
           "smoking.dta","stroke.dta","stroke_data.csv","survivaloa.csv"
           )
# Base URL
base_url <- "https://github.com/drkamarul/multivar_data_analysis/raw/refs/heads/main/data/"

# Destination folder
dest_folder <- "C:/RVSCode/Course/EpiData"
# Create destination folder if it doesn't exist
if (!dir.exists(dest_folder)) {
  dir.create(dest_folder, recursive = TRUE)
}

# Loop through each file and download
for (file in files)
    {
       file_url <- paste0(base_url, file)                            # Construct full URL
       dest_path <- file.path(dest_folder, file)                     # Construct destination path
       download.file(file_url, destfile = dest_path, mode = "wb")    # Download file to specified folder
       cat("Downloaded:", file, "to", dest_path, "\n")               # Print status 
    }
```


### ------------------------------------------------


# Download nhanes data
```{r download-data, include false}
# download file
download.file("https://github.com/Model-Lab-Net/Courses/tree/main/Epi/Data/nhanes.csv",
              destfile = "C:/RVSCode/Course/Data",
              mode = "wb"
              )
```


### ------------------------------------------------


# Download nhanes data /w variables
```{r download-data, include false}
# file url and destination information
base_url <- "https://github.com/Model-Lab-Net/Courses/tree/main/Epi/Data"
file <- "nhanes.csv"
dest_folder <- "C:/RVSCode/CourseCP/Data"
file_url <- paste0(base_url, file)
# download file
download.file(file_url, destfile = dest_path, mode = "wb")
```





# Download some course data
```{r download-data, include false}
# file url and destination information
download.file("https://raw.githubusercontent.com/GTPB/PSLS20/master/data/fev.txt", destfile = "C:/RVSCode/Course/Data/fev.txt")
```


# Load data packages
```{r setup-packages, include false}
# List of packages to install and load
packages <- c("readr", "epiR")

# Loop to check, install, and load packages
for (pkg in packages) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    install.packages(pkg, dependencies = TRUE)
    cat("Installed package:", pkg, "\n")
  }
  library(pkg, character.only = TRUE)
}
```


# Download more course data
```{r download-data, include false}
# List of files to download
files <- c("____", "____", "____", "____",
           "____", "____", "____", "____",
           "____", "____", "____", "____"
           )
# Base URL
base_url <- "https://github.com/Model-Lab-Net/Courses/tree/main/Epi/Data"

# Destination folder
dest_folder <- "C:/RVSCode/Course/Data"
# Create destination folder if it doesn't exist
if (!dir.exists(dest_folder)) {
  dir.create(dest_folder, recursive = TRUE)
}

# Loop through each file and download
for (file in files)
    {
       file_url <- paste0(base_url, file)                            # Construct full URL
       dest_path <- file.path(dest_folder, file)                     # Construct destination path
       download.file(file_url, destfile = dest_path, mode = "wb")    # Download file to specified folder
       cat("Downloaded:", file, "to", dest_path, "\n")               # Print status 
    }
```


# Download more datasets
```{r download-more_data, include false}
# List of files to download
files <- c("https://_____,",
           "https://_____,",
           "https://_____,",
           )
dest_folder <- "C:/RVSCode/Course/Data"

# Loop through each file and download
for (file in files)
    {
       download.file(file, destfile = dest_folder, mode = "wb")    # Download file to specified folder
       cat("Downloaded:", file, "to", dest_folder, "\n")           # Print status 
    }
```